### 2021-12-11 03-post-process/compare-sim-to-obs.Rmd
Store the ratio summaries for all estimated quantities

---
title: "Observed vs. Simulated Year Comparison"
subtitle: "`r paste0('Model Scenario: ', params$scenario)`"
output: html_document
params:
  scenario:
    label: "Scenario"
    value: "base-with-sim-medium"
    input: text
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center", fig.width = 8, fig.height = 4)
knitr::opts_knit$set(root.dir = "../")
```

```{r load}
# load packages
source("00-packages.R")

# load all necessary functions
invisible(sapply(list.files(path = "01-functions", pattern = "\\.R$", full.names = T), source))

# set the input directory
in_dir = "02-model/model-output"

# read information from this model
model_info = readRDS(file.path(in_dir, paste0("output-", params$scenario, ".rds")))

# extract the two key structures from this object
# post = model_info$post                               # full posterior takes a long time to run on this script
post = post_thin(model_info$post, keep_percent = 0.05)  # uncomment this to run document with 10% of the retained MCMC samples
jags_data = model_info$jags_data
```

```{r select-years}
# all of the years included in the model
all_yrs = as.numeric(rownames(jags_data$phi_SL))

# which is the first year of observation period?
# may want to use 2001 to cut off years without hatchery supplementation
fy_obs = which(all_yrs == 2001)

# for non sar values, which years correspond to observed and simulated years?
obs_y = with(jags_data, fy_obs:ny_obs)
sim_y = with(jags_data, (ny_obs+1):ny)

# same but for sar values. A bit different because not all adults have returned from the last brood year.
obs_sar_y = with(jags_data, fy_obs:(ny_obs-kmax))
sim_sar_y = with(jags_data, (ny_obs-kmax+1):(ny-kmax))
```

```{r add-p-hos}
# extract spawners returning to tributary
Ra = post_subset(post, "^Ra[", matrix = TRUE)

# for each posterior sample, calculate the fraction that are hatchery origin
out = NULL
for (i in 1:nrow(Ra)) {
  Ra_i = array_format(Ra[i,])
  nor = Ra_i[,1,1,] + Ra_i[,2,1,] + Ra_i[,3,1,]
  hor = Ra_i[,1,2,] + Ra_i[,2,2,] + Ra_i[,3,2,]
  out = rbind(out, as.numeric(hor/(nor + hor)))
}

# build the row indices for each element: the year
r_ind = rep(1:nrow(nor), ncol(nor))

# build the column indices for each element: the population
c_ind = rep(1:ncol(nor), each = nrow(nor))

# make element names
colnames(out) = paste0("p_hor[", r_ind, ",", c_ind, "]")

# drop out those that have NA values (first k_max years)
na_ind = which(!is.na(out[1,]))
out = out[,na_ind]

# combine this with the rest of the posterior samples
post = post_bind(post, out)
```

```{r add-Rb-tot}
# extract spawners returning to tributary
Rb = post_subset(post, "^Rb[", matrix = TRUE)

# for each posterior sample, calculate the fraction that are hatchery origin
out = NULL
for (i in 1:nrow(Rb)) {
  Rb_i = array_format(Rb[i,])
  nor = Rb_i[,1,1,] + Rb_i[,2,1,] + Rb_i[,3,1,]
  hor = Rb_i[,1,2,] + Rb_i[,2,2,] + Rb_i[,3,2,]
  out = rbind(out, as.numeric(nor + hor))
}

# build the row indices for each element: the year
r_ind = rep(1:nrow(nor), ncol(nor))

# build the column indices for each element: the population
c_ind = rep(1:ncol(nor), each = nrow(nor))

# make element names
colnames(out) = paste0("Rb_tot[", r_ind, ",", c_ind, "]")

# drop out those that have NA values (first k_max years)
na_ind = which(!is.na(out[1,]))
out = out[,na_ind]

# combine this with the rest of the posterior samples
post = post_bind(post, out)
```

```{r add-Ma-tot}
# extract smolt at LGR
Ma = post_subset(post, "^Ma[", matrix = TRUE)

# for each posterior sample, calculate the number of NOR and HOR smolt at LGR
out_nor = NULL
out_hor = NULL
for (i in 1:nrow(Ma)) {
  Ma_i = array_format(Ma[i,])
  nor = Ma_i[,1,1,] + Ma_i[,2,1,]
  hor = Ma_i[,2,2,]
  out_nor = rbind(out_nor, as.numeric(nor))
  out_hor = rbind(out_hor, as.numeric(hor))
}

# build the row indices for each element: the year
r_ind = rep(1:nrow(nor), ncol(nor))

# build the column indices for each element: the population
c_ind = rep(1:ncol(nor), each = nrow(nor))

# make element names
colnames(out_nor) = paste0("Ma_tot[", r_ind, ",", "1", ",", c_ind, "]")
colnames(out_hor) = paste0("Ma_tot[", r_ind, ",", "2", ",", c_ind, "]")
out = cbind(out_nor, out_hor)

# drop out those that have NA values (first k_max years)
na_ind = which(!is.na(out[1,]))
out = out[,na_ind]

# combine this with the rest of the posterior samples
post = post_bind(post, out)
```

```{r add-age-comp}
Ra = post_subset(post, "^Ra[", matrix = TRUE)

# for each posterior sample, calculate the fraction that of each age by origin
p3_nor_out = p4_nor_out = p5_nor_out = p3_hor_out = p4_hor_out = p5_hor_out = NULL
for (i in 1:nrow(Ra)) {
  Ra_i = array_format(Ra[i,])
  nor_tot = Ra_i[,1,1,] + Ra_i[,2,1,] + Ra_i[,3,1,]
  hor_tot = Ra_i[,1,2,] + Ra_i[,2,2,] + Ra_i[,3,2,]

  p3_nor = Ra_i[,1,1,]/nor_tot
  p4_nor = Ra_i[,2,1,]/nor_tot
  p5_nor = Ra_i[,3,1,]/nor_tot
  p3_hor = Ra_i[,1,2,]/hor_tot
  p4_hor = Ra_i[,2,2,]/hor_tot
  p5_hor = Ra_i[,3,2,]/hor_tot

  p3_nor_out = rbind(p3_nor_out, as.numeric(p3_nor))
  p4_nor_out = rbind(p4_nor_out, as.numeric(p4_nor))
  p5_nor_out = rbind(p5_nor_out, as.numeric(p5_nor))
  p3_hor_out = rbind(p3_hor_out, as.numeric(p3_hor))
  p4_hor_out = rbind(p4_hor_out, as.numeric(p4_hor))
  p5_hor_out = rbind(p5_hor_out, as.numeric(p5_hor))
}

# build the row indices for each element: the year
r_ind = rep(1:nrow(p3_nor), ncol(p3_nor))

# build the column indices for each element: the population
c_ind = rep(1:ncol(p3_nor), each = nrow(p3_nor))

# make element names
colnames(p3_nor_out) = paste0("Ra_comp[", r_ind, ",", "1", ",", "1", ",", c_ind, "]")
colnames(p4_nor_out) = paste0("Ra_comp[", r_ind, ",", "2", ",", "1", ",", c_ind, "]")
colnames(p5_nor_out) = paste0("Ra_comp[", r_ind, ",", "3", ",", "1", ",", c_ind, "]")
colnames(p3_hor_out) = paste0("Ra_comp[", r_ind, ",", "1", ",", "2", ",", c_ind, "]")
colnames(p4_hor_out) = paste0("Ra_comp[", r_ind, ",", "2", ",", "2", ",", c_ind, "]")
colnames(p5_hor_out) = paste0("Ra_comp[", r_ind, ",", "3", ",", "2", ",", c_ind, "]")

out = cbind(p3_nor_out, p4_nor_out, p5_nor_out, p3_hor_out, p4_hor_out, p5_hor_out)

# drop out those that have NA values (first k_max years)
out[out == "NaN"] = 0
na_ind = which(!is.na(out[1,]))
out = out[,na_ind]

# combine this with the rest of the posterior samples
post = post_bind(post, out)
```

```{r add-trib-smolt-per-parr}
Mb = post_subset(post, "^Mb[", matrix = TRUE)
Pa = post_subset(post, "^Pa[", matrix = TRUE)

# for each posterior sample, calculate survival from parr to lgr smolt by LH type
out_fall = out_spring = NULL

for (i in 1:nrow(Mb)) {
  Mb_i = array_format(Mb[i,])
  Pa_i = array_format(Pa[i,])

  Mb_per_Pa = Mb_i[,,1,]/Pa_i

  out_fall = rbind(out_fall, as.numeric(Mb_per_Pa[,1,]))
  out_spring = rbind(out_spring, as.numeric(Mb_per_Pa[,2,]))
}

# build the row indices for each element: the year
r_ind = rep(1:dim(Mb_per_Pa)[1], dim(Mb_per_Pa)[3])

# build the column indices for each element: the population
c_ind = rep(1:dim(Mb_per_Pa)[3], each = dim(Mb_per_Pa)[1])

# make element names
colnames(out_fall) = paste0("Mb_per_Pa[", r_ind, ",", "1", ",", c_ind, "]")
colnames(out_spring) = paste0("Mb_per_Pa[", r_ind, ",", "2", ",", c_ind, "]")

out = cbind(out_fall, out_spring)

# drop out those that have NA values (first k_max years)
na_ind = which(!is.na(out[1,]))
out = out[,na_ind]

# combine this with the rest of the posterior samples
post = post_bind(post, out)
```

```{r add-lgr-smolt-per-parr}
Ma = post_subset(post, "^Ma[", matrix = TRUE)
Pa = post_subset(post, "^Pa[", matrix = TRUE)

# for each posterior sample, calculate survival from parr to lgr smolt by LH type
out_fall = out_spring = NULL

for (i in 1:nrow(Ma)) {
  Ma_i = array_format(Ma[i,])
  Pa_i = array_format(Pa[i,])

  Ma_per_Pa = Ma_i[,,1,]/Pa_i

  out_fall = rbind(out_fall, as.numeric(Ma_per_Pa[,1,]))
  out_spring = rbind(out_spring, as.numeric(Ma_per_Pa[,2,]))
}

# build the row indices for each element: the year
r_ind = rep(1:dim(Ma_per_Pa)[1], dim(Ma_per_Pa)[3])

# build the column indices for each element: the population
c_ind = rep(1:dim(Ma_per_Pa)[3], each = dim(Ma_per_Pa)[1])

# make element names
colnames(out_fall) = paste0("Ma_per_Pa[", r_ind, ",", "1", ",", c_ind, "]")
colnames(out_spring) = paste0("Ma_per_Pa[", r_ind, ",", "2", ",", c_ind, "]")

out = cbind(out_fall, out_spring)

# drop out those that have NA values (first k_max years)
na_ind = which(!is.na(out[1,]))
out = out[,na_ind]

# combine this with the rest of the posterior samples
post = post_bind(post, out)

```

```{r add-nor-lgr-smolt-aggregate}
# extract smolt at LGR
Ma = post_subset(post, "^Ma[", matrix = TRUE)

# for each posterior sample, calculate the total aggregate return
out_nor = out_hor = NULL
for (i in 1:nrow(Ma)) {
  Ma_i = array_format(Ma[i,])
  nor = Ma_i[,1,1,] + Ra_i[,2,1,]
  hor = Ra_i[,2,2,]

  out_nor = rbind(out_nor, rowSums(nor))
  out_hor = rbind(out_hor, rowSums(hor))
}

# build the indices for each element: the year
r_ind = 1:nrow(nor)

# make element names
colnames(out_nor) = paste0("Ma_aggregate[", r_ind, ",", "1", "]")
colnames(out_hor) = paste0("Ma_aggregate[", r_ind, ",", "2", "]")
out = cbind(out_nor, out_hor)

# drop out those that have NA values (first k_max years)
na_ind = which(!is.na(out[1,]))
out = out[,na_ind]

# combine this with the rest of the posterior samples
post = post_bind(post, out)
```

```{r add-Ra-aggregate}
# extract spawners returning to tributary
Ra = post_subset(post, "^Ra[", matrix = TRUE)

# for each posterior sample, calculate the total aggregate return
out_nor = out_hor = NULL
for (i in 1:nrow(Ra)) {
  Ra_i = array_format(Ra[i,])
  nor = Ra_i[,1,1,] + Ra_i[,2,1,] + Ra_i[,3,1,]
  hor = Ra_i[,1,2,] + Ra_i[,2,2,] + Ra_i[,3,2,]
  tot = nor + hor

  out_nor = rbind(out_nor, rowSums(nor))
  out_hor = rbind(out_hor, rowSums(hor))
}

# build the indices for each element: the year
r_ind = 1:nrow(nor)

# make element names
colnames(out_nor) = paste0("Ra_aggregate[", r_ind, ",", "1", "]")
colnames(out_hor) = paste0("Ra_aggregate[", r_ind, ",", "2", "]")
out = cbind(out_nor, out_hor)

# drop out those that have NA values (first k_max years)
na_ind = which(!is.na(out[1,]))
out = out[,na_ind]

# combine this with the rest of the posterior samples
post = post_bind(post, out)
```

```{r dim_names fn}
# function to create dimension names from numeric indicators
dim_names = function(LH_type = NULL, age = NULL, origin = NULL, pop = NULL) {

  # empty list
  out = list()

  # combine the dimention names for each type supplied
  if (!is.null(LH_type)) out = append(out, list(LH_type = c("fall", "spring")[LH_type]))
  if (!is.null(age)) out = append(out, list(age = c("age3", "age4", "age5")[age]))
  if (!is.null(origin)) out = append(out, list(origin = c("NOR", "HOR")[origin]))
  if (!is.null(pop)) out = append(out, list(pop = c("CAT", "LOS", "MIN", "UGR")[pop]))

  return(out)
}
```

```{r get_ratio fn}
# function to calculate the ratio of mean and CV between simulated and observed years
get_ratio = function(param, obs_y, sim_y, ...) {

  # get node names for the observed years
  obs_param = sub_index(param, year = obs_y, ...)

  # get node names for the simulated years
  sim_param = sub_index(param, year = sim_y, ...)

  # extract posterior samples of these
  obs_x = post_subset(post, obs_param, matrix = TRUE)
  sim_x = post_subset(post, sim_param, matrix = TRUE)

  # obtain a nicely formatted parameter identifier
  named_param = do.call(sub_index, append(list(x = param), dim_names(...)))
  named_param = stringr::str_remove(named_param, "year,")
  named_param = stringr::str_extract(named_param, "\\[.+\\]")
  named_param = stringr::str_remove(named_param, "age[:digit:],")
  named_param = stringr::str_remove(named_param, "\\[")
  named_param = stringr::str_remove(named_param, "\\]")

  # function calculate lag-1 autocorrelation coefficient
  get_1_acf = function(x) acf(x, lag.max = 1, plot = FALSE)$acf[2,1,1]

  # calculate and return the summary statistics for each posterior sample and period
  vals_out = array(NA, dim = c(nrow(obs_x), 3, 3, 1))
  dimnames(vals_out) = list(iter = 1:nrow(obs_x), period = c("obs", "sim", "sim-v-obs"), stat = c("cv", "mean", "acf"), dims = named_param)
  vals_out[,"obs","mean",] = sapply(1:nrow(obs_x), function(i) mean(obs_x[i,]))
  vals_out[,"sim","mean",] = sapply(1:nrow(sim_x), function(i) mean(sim_x[i,]))
  vals_out[,"obs","cv",] = sapply(1:nrow(obs_x), function(i) sd(obs_x[i,])/mean(obs_x[i,]))
  vals_out[,"sim","cv",] = sapply(1:nrow(sim_x), function(i) sd(sim_x[i,])/mean(sim_x[i,]))
  vals_out[,"obs","acf",] = sapply(1:nrow(obs_x), function(i) get_1_acf(obs_x[i,]))
  vals_out[,"sim","acf",] = sapply(1:nrow(sim_x), function(i) get_1_acf(sim_x[i,]))
  vals_out[,"sim-v-obs","cv",] = vals_out[,"sim","cv",]/vals_out[,"obs","cv",]
  vals_out[,"sim-v-obs","mean",] = vals_out[,"sim","mean",]/vals_out[,"obs","mean",]
  vals_out[,"sim-v-obs","acf",] = vals_out[,"sim","acf",] - vals_out[,"obs","acf",]

  return(vals_out)
}

```

```{r ratio_plot fn}
ratio_plot = function(ratios, line_y, ...) {

  # calculate summaries of the ratios
  summs = apply(ratios, 2, function(x) quantile(x, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm = TRUE))

  # set graphics parameters
  par(mar = c(7,1,1,1), mgp = c(2,0.35,0), tcl = -0.15, lend = "square", ljoin = "mitre")

  # make an empty plot with the right dimensions
  mp = barplot(summs["50%",], col = "white", ylab = "", border = "white", las = 2,
               ylim = c(min(0, min(summs, na.rm = TRUE)), max(summs, na.rm = TRUE)) * 1.1, ...)

  # draw tick labels
  axis(side = 1, at = mp, labels = F)

  # create an "acceptable region"
  usr = par("usr"); rect(usr[1], line_y - 0.25, usr[2], line_y + 0.25, col = "grey90", border = NA)
  abline(h = line_y, lty = 2, col = "grey")

  # draw the ratio summaries
  segments(mp, summs["2.5%",], mp, summs["97.5%",])
  segments(mp, summs["25%",], mp, summs["75%",], lwd = 6)
  points(summs["50%",] ~ mp, pch = 3, cex = 1.5)
  box()
}
```

```{r ratio_plots fn}
ratio_plots = function(x, main) {
  par(mfrow = c(1,3), oma = c(0,2,2,0))
  ratio_plot(x[,"sim-v-obs","cv",], main = "CV", line_y = 1)
  ratio_plot(x[,"sim-v-obs","mean",], main = "Mean", line_y = 1)
  ratio_plot(x[,"sim-v-obs","acf",], main = "ACF (Sim - Obs)", line_y = 0)
  mtext(side = 2, outer = TRUE, adj = 0.75, line = 1, "Ratio (Sim:Obs)")
  mtext(side = 3, outer = TRUE, line = 0.5, main, cex = 1.25, font = 2)
}
```

```{r vals_plot fn}
vals_plot = function(vals, ...) {
  f = function(x) quantile(x, c(0.025, 0.25, 0.5, 0.75, 0.975), na.rm = TRUE)

  summs = array(NA, dim = c(2,5,dim(vals)[3]))
  dimnames(summs) = list(c("obs", "sim"), names(f(rnorm(10))), dimnames(x)[[4]])

  summs["obs",,] = apply(vals[,"obs",], 2, f)
  summs["sim",,] = apply(vals[,"sim",], 2, f)

  par(mar = c(7,1,1,1), mgp = c(2,0.35,0), tcl = -0.15, lend = "square", ljoin = "mitre")
  mp = barplot(summs[,"50%",], beside = TRUE, col = "white", border = "white", ylim = range(summs, na.rm = TRUE) + diff(range(summs, na.rm = TRUE)) * c(-0.05,0.05), xaxt = "n", ...)
  segments(mp, summs[,"2.5%",], mp, summs[,"97.5%",], col = c("red", "blue"))
  segments(mp, summs[,"25%",], mp, summs[,"75%",], col = c("red", "blue"), lwd = 6)
  points(y = summs[,"50%",], x = mp, col = c("red", "blue"), pch = 3, cex = 1.5)
  axis(side = 1, at = colSums(mp)/2, labels = dimnames(summs)[[3]], las = 2)
  box()
}
```

```{r vals_plots fn}
vals_plots = function(x, main) {
  par(mfrow = c(1,3), oma = c(0,2,2,0))
  vals_plot(x[,c("obs", "sim"),"cv",], main = "CV")
  vals_plot(x[,c("obs", "sim"),"mean",], main = "Mean")
  vals_plot(x[,c("obs", "sim"),"acf",], main = "ACF")
  mtext(side = 2, outer = TRUE, adj = 0.75, line = 1, "Value")
  mtext(side = 3, outer = TRUE, line = 0.5, main, cex = 1.25, font = 2)
}
```

```{r summarize_ratios fn}
summarize_ratios = function(x, Class, Type) {
  xsumm = as.data.frame(t(apply(x[,"sim-v-obs",,], 3, colMeans)))
  xsumm = cbind(Class = Class, Type = Type, ID = rownames(xsumm), xsumm); rownames(xsumm) = NULL
  return(xsumm)
}
```

```{r assign_color fns}
# functions to assign a color based on a numeric value

# values close to 1 will be closer to white; values closer to 0 will be blue; values closer to 2 will be red
assign_color_1 = function(x) {
  red_pal = colorRampPalette(c("white", "red"))
  blue_pal = colorRampPalette(c("blue", "white"))
  red_cols = red_pal(5)
  blue_cols = blue_pal(5)
  cols = c(blue_cols, red_cols)

  breaks = c(0, 0.25, 0.5, 0.75, 0.9, 1.1, 1.25, 1.5, 1.75, 3)
  bin_names = levels(cut(runif(1e6, 0, 2), breaks = breaks, include.lowest = TRUE))
  names(cols) = c(bin_names[1:5], bin_names[5:9])
  bins = cut(x, breaks = breaks, include.lowest = TRUE)
  cols_use = cols[as.character(bins)]
  cols_use[is.na(cols_use)] = cols[5]
  alpha(unname(cols_use), 0.5)
}

# values close to 0 will be closer to white; values closer to -1 will be blue; values closer to 1 will be red
assign_color_0 = function(x) {
  red_pal = colorRampPalette(c("white", "red"))
  blue_pal = colorRampPalette(c("blue", "white"))
  red_cols = red_pal(5)
  blue_cols = blue_pal(5)
  cols = c(blue_cols, red_cols)

  breaks = c(-1, -0.75, -0.5, -0.25, -0.1, 0.1, 0.25, 0.5, 0.75, 1)
  bin_names = levels(cut(runif(1e6, -1, 1), breaks = breaks, include.lowest = TRUE))
  names(cols) = c(bin_names[1:5], bin_names[5:9])
  bins = cut(x, breaks = breaks, include.lowest = TRUE)
  cols_use = cols[as.character(bins)]
  cols_use[is.na(cols_use)] = cols[5]
  alpha(unname(cols_use), 0.5)
}
```

```{r prep_est_ratios fn}
# function to prepare ratios for table-making
prep_est_ratios = function(x, q) {
  # add/format additional column IDs
  x$origin = stringr::str_extract(x$ID, "NOR|HOR")
  x$pop = stringr::str_extract(x$ID, "CAT|LOS|MIN|UGR")
  x$LH_type = stringr::str_extract(x$ID, "fall|spring")
  x$LH_type[x$LH_type == "spring"] = "Spring"
  x$LH_type[x$LH_type == "fall"] = "Fall"
  x$pop = ifelse(is.na(x$pop), "Combined", x$pop)
  x$origin = ifelse(is.na(x$origin), "Combined", x$origin)
  x$LH_type = ifelse(is.na(x$LH_type), "Combined", x$LH_type)

  # make populations separate columns
  x = dcast(x, Class + Type + origin + LH_type ~ pop, value.var = q)

  # reorder columns
  x = x[,c("Class", "Type", "LH_type", "origin", "CAT", "LOS", "MIN", "UGR", "Combined")]

  # reorder rows
  x$Class = factor(x$Class, levels = c("FW Juvenile", "SW Juvenile", "FW Adult"))
  x$Type = factor(x$Type, levels = c("Egg to Parr Survival", "Proportion Fall Migrants", "Overwinter Survival", "Migration to LGR Survival", "LGR to Ocean Survival", "First Year Ocean Survival", "Age-3 Maturity", "Age-4 Maturity", "BON to LGR Survival", "Pre-Spawn Survival"))
  x$LH_type = factor(x$LH_type, levels = c("Combined", "Fall", "Spring"))
  x$origin = factor(x$origin, levels = c("Combined", "NOR", "HOR"))
  x = x[order(x$Class, x$Type, x$LH_type, x$origin),]

  colnames(x)[which(colnames(x) == "LH_type")] = "Mig. Type"
  colnames(x)[which(colnames(x) == "Origin")] = "Origin"

  return(x)
}
```

```{r prep_derived_ratios fn}
# function to prepare ratios for table-making
prep_derived_ratios = function(x, q) {
  # add/format additional column IDs
  x$origin = stringr::str_extract(x$ID, "NOR|HOR")
  x$pop = stringr::str_extract(x$ID, "CAT|LOS|MIN|UGR")
  x$LH_type = stringr::str_extract(x$ID, "fall|spring")
  x$LH_type[x$LH_type == "spring"] = "Spring"
  x$LH_type[x$LH_type == "fall"] = "Fall"
  x$pop = ifelse(is.na(x$pop), "Combined", x$pop)
  x$origin = ifelse(is.na(x$origin), "Combined", x$origin)
  x$LH_type = ifelse(is.na(x$LH_type), "Combined", x$LH_type)

  # make populations separate columns
  x = dcast(x, Class + Type + origin + LH_type ~ pop, value.var = q)

  # reorder columns
  x = x[,c("Class", "Type", "LH_type", "origin", "CAT", "LOS", "MIN", "UGR", "Combined")]

  # reorder rows
  x$Class = factor(x$Class, levels = c("Abundance", "Production Rates", "SAR", "Adult Composition"))
  x$Type = factor(x$Type, levels = c("Total Adults (Estuary)", "Total Adults (To Tributary)", "Total Spawners", "Total Egg Production", "Total Parr Recruitment", "Total LGR Smolt", "Aggregate Total Adult Return (To Tributary)", "Aggregate Total Smolt (At LGR)", "Parr per Spawner", "Smolt in Tributary per Spawner", "Spawners per Spawner", "Smolt in Tributary per Parr", "Smolt at LGR per Parr", "BON to BON", "LGR to LGR", "Proportion Hatchery Return", "Proportion Age-3 Return", "Proportion Age-4 Return", "Proportion Age-5 Return"))
  x$LH_type = factor(x$LH_type, levels = c("Combined", "Fall", "Spring"))
  x$origin = factor(x$origin, levels = c("Combined", "NOR", "HOR"))
  x = x[order(x$Class, x$Type, x$LH_type, x$origin),]

  colnames(x)[which(colnames(x) == "LH_type")] = "Mig. Type"
  colnames(x)[which(colnames(x) == "Origin")] = "Origin"

  return(x)
}
```

```{r make_ratios_table fn}
# function to make the table
make_ratios_table = function(ratios, q) {

  x = ratios
  table_x = x
  table_x[,5:9] = round(table_x[,5:9],2)
  table_x[is.na(table_x)] = ""

  # select the correct color rule
  if (q %in% c("mean", "cv")) {
    assign_color = assign_color_1
  } else {
    assign_color = assign_color_0
  }

  kable(table_x, "html", digits = 2, row.names = FALSE, align = "llllccccc") %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("condensed")) %>%
    column_spec(5, background = assign_color(x[,5])) %>%
    column_spec(6, background = assign_color(x[,6])) %>%
    column_spec(7, background = assign_color(x[,7])) %>%
    column_spec(8, background = assign_color(x[,8])) %>%
    column_spec(9, background = assign_color(x[,9])) %>%
    collapse_rows(1:4)
}
```

For us to use the state-space model as a simulation model, it would be desirable if several characteristics of the distribution of stochastic and derived quantities were similar between the time periods with and without observations.

This document is intended to identify whether simulated outcomes are similar to observed outcomes. It looks at two primary types of quantities: stochastic time series parameters (e.g., egg to parr survival fluctuations not attributable to density dependence) and emergent outcomes (e.g., total adult return abundance, BON to BON SAR rates, etc.). Because all emergent outcomes are derived from the stochastic elements, issues with emergent outcomes should be traceable to specific stochastic parents.

The JAGS model was altered to carry out the life cycle processes through brood year 2050. This means for each posterior sample we have an observed and simulated time series. The simulated time series picks up where the observed on left off, and assumes the same random processes for all components as used in the observed time series. Several quantities are supplied to the model as known without error (weir removals, hatchery smolt releases, harvest rates below BON, etc.). These time series were extended to the future years external to the model using basic averages or resampling from historical outcomes as appropriate. This exercise is purely for model validation.

Observed vs. simulated consistency is visualized in two ways: by comparing summary statistics of the inter-annual variation and central tendency and by directly comparing the historical time series with 5 random simulated futures.

# Distribution Comparison Plots

These plots show summaries of a quantity in the model:

1. **Inter-annual CV**: measures how variable the quantity is among years
2. **Inter-annual mean**: measures the central tendency of the quantity across years
3. **Lag-1 autocorrelation**: measures whether consecutive years are independent

If these summary statistics are inconsistent among the observed and simulated time periods, that indicates that the state-space model has not sufficiently captured the random processes that have generated the observed outcomes and that we may have reason to question the reliability of its ability to simulate future outcomes, at least as a direct extension of the historical population dynamics.

In all figures in this section, these time periods were used:

* **Observed**: `r paste0(all_yrs[min(obs_y)], " -- ", all_yrs[max(obs_y)])`
* **Simulated**: `r paste0(all_yrs[min(sim_y)], " -- ", all_yrs[max(sim_y)])`

The entire 1991 - 2019 period isn't used for the observed period because the early years did not have a hatchery program but the simulation has hatchery input in all years. Thus, including the non-hatchery years may throw off the inference.

For each quantity, two types of plots are shown:

* **Estimates**: this shows the CV, mean, and autocorrelation coefficient for a quantity based on the time period (red = observed, blue = simulated)
* **Ratios**: this shows a ratio between the summary statistic between simulated and observed periods. For CV and mean, this is simulated/observed, and for autocorrelation this is simulated - observed.

These summary statistics were calculated for each posterior sample. In all plots, the cross mark represents the posterior median, the thick bar represents the central 50% of posterior samples, and the thin bar represents the central 95% of posterior samples.

```{r}
est_ratios = NULL
```

## Time Series Parameters {.tabset .tabset-pills}

### Freshwater Juvenile Phase {.tabset .tabset-pills}

#### Egg to Parr Survival

```{r}
x = lapply(1:4, function(j) get_ratio("^Pb_per_f_tot[year,pop]", obs_y, sim_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Egg to Parr Survival: Estimates")
ratio_plots(x, "Egg to Parr Survival: Ratios")
est_ratios = rbind(est_ratios, summarize_ratios(x, Class = "FW Juvenile", Type = "Egg to Parr Survival"))
```

#### Proportion Fall Migrants

```{r}
x = lapply(1:4, function(j) get_ratio("^pi[year,LH_type,pop]", obs_y, sim_y, pop = j, LH_type = 1))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Proportion Fall Migrants: Estimates")
ratio_plots(x, "Proportion Fall Migrants: Ratios")
est_ratios = rbind(est_ratios, summarize_ratios(x, Class = "FW Juvenile", Type = "Proportion Fall Migrants"))
```

#### Overwinter Survival

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("^phi_Pa_Mb[year,LH_type,pop]", obs_y, sim_y, pop = j, LH_type = 1)),
  lapply(1:4, function(j) get_ratio("^phi_Pa_Mb[year,LH_type,pop]", obs_y, sim_y, pop = j, LH_type = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Overwinter Survival: Estimates")
ratio_plots(x, "Overwinter Survival: Ratios")
est_ratios = rbind(est_ratios, summarize_ratios(x, Class = "FW Juvenile", Type = "Overwinter Survival"))
```

#### Migration to LGR Survival

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("^phi_Mb_Ma[year,LH_type,origin,pop]", obs_y, sim_y, pop = j, LH_type = 2, origin = 1)),
  lapply(1:4, function(j) get_ratio("^phi_Mb_Ma[year,LH_type,origin,pop]", obs_y, sim_y, pop = j, LH_type = 2, origin = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Migration to LGR Survival: Estimates")
ratio_plots(x, "Migration to LGR Survival: Ratios")
est_ratios = rbind(est_ratios, summarize_ratios(x, Class = "FW Juvenile", Type = "Migration to LGR Survival"))
```

#### LGR to Ocean Survival

```{r}
x = append(
  list(get_ratio("^phi_Ma_O0[year,origin]", obs_y, sim_y, origin = 1)),
  list(get_ratio("^phi_Ma_O0[year,origin]", obs_y, sim_y, origin = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "LGR to Ocean Survival: Estimates")
ratio_plots(x, "LGR to Ocean Survival: Ratios")
est_ratios = rbind(est_ratios, summarize_ratios(x, Class = "FW Juvenile", Type = "LGR to Ocean Survival"))
```

### Marine Juvenile Phase {.tabset .tabset-pills}

#### Year 1 Ocean Survival

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("phi_O0_O1[year,origin,pop]", obs_y, sim_y, origin = 1, pop = j)),
  lapply(1:4, function(j) get_ratio("phi_O0_O1[year,origin,pop]", obs_y, sim_y, origin = 2, pop = j))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "First Year Ocean Survival: Estimates")
ratio_plots(x, "First Year Ocean Survival: Ratios")
est_ratios = rbind(est_ratios, summarize_ratios(x, Class = "SW Juvenile", Type = "First Year Ocean Survival"))
```

#### Age-3 Maturity

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("^psi_O1_Rb[year,origin,pop]", obs_y, sim_y, origin = 1, pop = j)),
  lapply(1:4, function(j) get_ratio("^psi_O1_Rb[year,origin,pop]", obs_y, sim_y, origin = 2, pop = j))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Age-3 Maturity: Estimates")
ratio_plots(x, "Age-3 Maturity: Ratios")
est_ratios = rbind(est_ratios, summarize_ratios(x, Class = "SW Juvenile", Type = "Age-3 Maturity"))
```

#### Age-4 Maturity

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("^psi_O2_Rb[year,origin,pop]", obs_y, sim_y, origin = 1, pop = j)),
  lapply(1:4, function(j) get_ratio("^psi_O2_Rb[year,origin,pop]", obs_y, sim_y, origin = 2, pop = j))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Age-4 Maturity: Estimates")
ratio_plots(x, "Age-4 Maturity: Ratios")
est_ratios = rbind(est_ratios, summarize_ratios(x, Class = "SW Juvenile", Type = "Age-4 Maturity"))
```

### Freshwater Adult Phase {.tabset .tabset-pills}

#### BON to LGR Survival

```{r}
x = append(
  list(get_ratio("^phi_Rb_Ra[year,origin]", obs_y, sim_y, origin = 1)),
  list(get_ratio("^phi_Rb_Ra[year,origin]", obs_y, sim_y, origin = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "BON to LGR Survival: Estimates")
ratio_plots(x, "BON to LGR Survival: Ratios")
est_ratios = rbind(est_ratios, summarize_ratios(x, Class = "FW Adult", Type = "BON to LGR Survival"))
```

#### Pre-Spawn Survival

```{r}
x = lapply(1:4, function(j) get_ratio("^phi_Sb_Sa[year,pop]", obs_y, sim_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Pre-Spawn Survival: Estimates")
ratio_plots(x, "Pre-Spawn Survival: Ratios")
est_ratios = rbind(est_ratios, summarize_ratios(x, Class = "FW Adult", Type = "Pre-Spawn Survival"))
```

## Emergent Properties {.tabset .tabset-pills}

### Abundance Quantities {.tabset .tabset-pills}

#### Total Adult Return (To Estuary)

```{r}
x = lapply(1:4, function(j) get_ratio("^Rb_tot[year,pop]", obs_y, sim_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Total Adults (Estuary): Estimates")
ratio_plots(x, "Total Adults (Estuary): Ratios")
```

#### Total Adult Return (To Tributary)

```{r}
x = lapply(1:4, function(j) get_ratio("^Ra_tot[year,pop]", obs_y, sim_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Total Adults (Tributary): Estimates")
ratio_plots(x, "Total Adults (Tributary): Ratios")
```

#### Total Spawners

```{r}
x = lapply(1:4, function(j) get_ratio("^Sa_tot[year,pop]", obs_y, sim_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Total Spawners: Estimates")
ratio_plots(x, "Total Spawners: Ratios")
```

#### Total Egg Production

```{r}
x = lapply(1:4, function(j) get_ratio("^f_tot[year,pop]", obs_y, sim_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Total Egg Production: Estimates")
ratio_plots(x, "Total Egg Production: Ratios")
```

#### Total Parr Recruitment

```{r}
x = lapply(1:4, function(j) get_ratio("^Pb[year,pop]", obs_y, sim_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Total Parr Recruitment: Estimates")
ratio_plots(x, "Total Parr Recruitment: Ratios")
```

#### Total LGR Smolt

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("Ma_tot[year,origin,pop]", obs_y, sim_y, pop = j, origin = 1)),
  lapply(1:4, function(j) get_ratio("Ma_tot[year,origin,pop]", obs_y, sim_y, pop = j, origin = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Total LGR Smolt: Estimates")
ratio_plots(x, "Total LGR Smolt: Ratios")
#### Aggregate Total Adult Return (To Tributary)

```{r}
x = append(
  list(get_ratio("Ra_aggregate[year,origin]", obs_y, sim_y, origin = 1)),
  list(get_ratio("Ra_aggregate[year,origin]", obs_y, sim_y, origin = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Aggregate Total Adult Return (To Tributary): Estimates")
ratio_plots(x, "Aggregate Total Adult Return (To Tributary): Ratios")
derived_ratios = rbind(derived_ratios, summarize_ratios(x, Class = "Abundance", Type = "Aggregate Total Adult Return (To Tributary)"))
```

#### Aggregate Total Smolt (At LGR)

```{r}
x = append(
  list(get_ratio("Ma_aggregate[year,origin]", obs_y, sim_y, origin = 1)),
  list(get_ratio("Ma_aggregate[year,origin]", obs_y, sim_y, origin = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Aggregate Total Smolt (At LGR): Estimates")
ratio_plots(x, "Aggregate Total Smolt (At LGR): Ratios")
derived_ratios = rbind(derived_ratios, summarize_ratios(x, Class = "Abundance", Type = "Aggregate Total Smolt (At LGR)"))
```

### SAR Rates {.tabset .tabset-pills}

#### BON to BON SAR

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("phi_O0_Rb_BON[year,origin,pop]", obs_sar_y, sim_sar_y, pop = j, origin = 1)),
  lapply(1:4, function(j) get_ratio("phi_O0_Rb_BON[year,origin,pop]", obs_sar_y, sim_sar_y, pop = j, origin = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "BON to BON SAR: Estimates")
ratio_plots(x, "BON to BON SAR: Ratios")
```

#### LGR to LGR SAR

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("Ra_per_Ma[year,origin,pop]", obs_sar_y, sim_sar_y, pop = j, origin = 1)),
  lapply(1:4, function(j) get_ratio("Ra_per_Ma[year,origin,pop]", obs_sar_y, sim_sar_y, pop = j, origin = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "LGR to LGR SAR: Estimates")
ratio_plots(x, "LGR to LGR SAR: Ratios")
```

### Production Rates {.tabset .tabset-pills}

#### Parr per Spawner

```{r}
x = lapply(1:4, function(j) get_ratio("Pb_per_Sa_tot[year,pop]", obs_sar_y, sim_sar_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Parr per Spawner: Estimates")
ratio_plots(x, "Parr per Spawner: Ratios")
```

#### Smolt per Spawner

```{r}
x = lapply(1:4, function(j) get_ratio("Mb_per_Sa_tot[year,pop]", obs_sar_y, sim_sar_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Smolt per Spawner: Estimates")
ratio_plots(x, "Smolt per Spawner: Ratios")
```

#### Smolt in Tributary per Parr

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("Mb_per_Pa[year,LH_type,pop]", obs_y, sim_y, pop = j, LH_type = 1)),
  lapply(1:4, function(j) get_ratio("Mb_per_Pa[year,LH_type,pop]", obs_y, sim_y, pop = j, LH_type = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Smolt in Tributary per Parr: Estimates")
ratio_plots(x, "Smolt in Tributary per Parr: Ratios")
```

#### Smolt at LGR per Parr

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("Ma_per_Pa[year,LH_type,pop]", obs_y, sim_y, pop = j, LH_type = 1)),
  lapply(1:4, function(j) get_ratio("Ma_per_Pa[year,LH_type,pop]", obs_y, sim_y, pop = j, LH_type = 2))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Smolt at LGR per Parr: Estimates")
ratio_plots(x, "Smolt at LGR per Parr: Ratios")
```

#### Spawners per Spawner

```{r}
x = lapply(1:4, function(j) get_ratio("Sa_tot_per_Sa_tot[year,pop]", obs_sar_y, sim_sar_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Spawners per Spawner: Estimates")
ratio_plots(x, "Spawners per Spawner: Ratios")
```

### Adult Composition {.tabset .tabset-pills}

#### Proportion Hatchery Return

```{r}
x = lapply(1:4, function(j) get_ratio("p_hor[year,pop]", obs_y, sim_y, pop = j))
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Proportion Hatchery Return: Estimates")
ratio_plots(x, "Proportion Hatchery Return: Ratios")
```

#### Proportion Age-3 Return

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("Ra_comp[year,age,origin,pop]", obs_y, sim_y, age = 1, origin = 1, pop = j)),
  lapply(1:4, function(j) get_ratio("Ra_comp[year,age,origin,pop]", obs_y, sim_y, age = 1, origin = 2, pop = j))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Proportion Age-3 Return: Estimates")
ratio_plots(x, "Proportion Age-3 Return: Ratios")
```

#### Proportion Age-4 Return

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("Ra_comp[year,age,origin,pop]", obs_y, sim_y, age = 2, origin = 1, pop = j)),
  lapply(1:4, function(j) get_ratio("Ra_comp[year,age,origin,pop]", obs_y, sim_y, age = 2, origin = 2, pop = j))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Proportion Age-4 Return: Estimates")
ratio_plots(x, "Proportion Age-4 Return: Ratios")
```

#### Proportion Age-5 Return

```{r}
x = append(
  lapply(1:4, function(j) get_ratio("Ra_comp[year,age,origin,pop]", obs_y, sim_y, age = 3, origin = 1, pop = j)),
  lapply(1:4, function(j) get_ratio("Ra_comp[year,age,origin,pop]", obs_y, sim_y, age = 3, origin = 2, pop = j))
)
x = do.call(abind, append(x, list(along = 4)))
vals_plots(x, "Proportion Age-5 Return: Estimates")
ratio_plots(x, "Proportion Age-5 Return: Ratios")
```

# Time Series Plots

```{r}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center", fig.width = 6, fig.height = 8)
```

```{r ts_plot fn}
n_sim_draw = 5
isim = sample(1:post_dim(post, "saved"), n_sim_draw, replace = FALSE)
obs_y = 6:jags_data$ny_obs
obs_sar_y = 6:(jags_data$ny_obs-jags_data$kmax)
sim_sar_y = (jags_data$ny_obs-jags_data$kmax+1):(jags_data$ny-jags_data$kmax)

# param = "phi_O0_Rb_BON[year,origin,pop]"
# obs_y_use = obs_sar_y
# sim_y_use = sim_sar_y
# origin = 1
# main = "main"

ts_plot = function(param, main, obs_y_use = obs_y, sim_y_use = sim_y, ...) {

  obs_vals = post_summ(post, sub_index(param, year = obs_y_use, ...))
  sim_vals = post_subset(post, sub_index(param, year = sim_y_use, ...), matrix = TRUE)
  # obs_vals = post_summ(post, sub_index(param, year = obs_y_use, pop = 1, origin = 1))
  # sim_vals = post_subset(post, sub_index(param, year = sim_y_use, pop = 1, origin = 1), matrix = TRUE)

  par(mar = c(1,2,2,1), mgp = c(2,0.35,0), tcl = -0.15)
  plot(obs_vals["mean",] ~ obs_y_use, type = "n", xlim = range(6, sim_y_use),
       ylim = range(obs_vals[c("2.5%", "97.5%"),], sim_vals[isim,]), xaxt = "n", xlab = "", ylab = "", main = main)
  usr = par("usr"); at_x = axisTicks(usr[1:2], log = FALSE)
  at_x = c(6, at_x[!(at_x %in% c(0, 6, max(sim_y_use)))], max(sim_y_use))
  axis(side = 1, at = at_x, labels = all_yrs[at_x])

  polygon(c(obs_y_use, rev(obs_y_use)), c(obs_vals["2.5%",], rev(obs_vals["97.5%",])), border = NA, col = alpha("salmon", 0.5))
  lines(obs_vals["mean",] ~ obs_y_use, col = "red", lwd = 2)

  junk = sapply(isim, function(i) lines(sim_vals[i,] ~ sim_y_use, col = alpha("blue", 0.5)))
}

pop_ts_plots = function(param, obs_y_use = obs_y, sim_y_use = sim_y, ...) {
  par(mfrow = c(4,1), oma = c(2,2,0,0))
  ts_plot(param, main = "CAT", obs_y_use = obs_y_use, sim_y_use = sim_y_use, pop = 1, ...)
  ts_plot(param, main = "LOS", obs_y_use = obs_y_use, sim_y_use = sim_y_use, pop = 2, ...)
  ts_plot(param, main = "MIN", obs_y_use = obs_y_use, sim_y_use = sim_y_use, pop = 3, ...)
  ts_plot(param, main = "UGR", obs_y_use = obs_y_use, sim_y_use = sim_y_use, pop = 4, ...)
  mtext(side = 1, outer = TRUE, line = 0.5, "Year")
  mtext(side = 2, outer = TRUE, line = 0.5, "Value")
}
```

## Time Series Parameters {.tabset .tabset-pills}

### Freshwater Juvenile Phase {.tabset .tabset-pills}

#### Egg to Parr Survival

```{r}
pop_ts_plots("^Pb_per_f_tot[year,pop]")
```

#### Proportion Fall Migrants

```{r}
pop_ts_plots("^pi[year,LH_type,pop]", LH_type = 1)
```

#### Overwinter Survival {.tabset .tabset-pills}

##### Fall Migrants

```{r}
pop_ts_plots("^phi_Pa_Mb[year,LH_type,pop]", LH_type = 1)
```

##### Spring Migrants

```{r}
pop_ts_plots("^phi_Pa_Mb[year,LH_type,pop]", LH_type = 2)
```

#### Migration to LGR Survival {.tabset .tabset-pills}

##### NOR

```{r}
pop_ts_plots("^phi_Mb_Ma[year,LH_type,origin,pop]", LH_type = 2, origin = 1)
```

##### HOR

```{r}
pop_ts_plots("^phi_Mb_Ma[year,LH_type,origin,pop]", LH_type = 2, origin = 2)
```

#### LGR to Ocean Survival

```{r}
par(mfrow = c(2,1), oma = c(2,2,0,0))
ts_plot("^phi_Ma_O0[year,origin]", main = "NOR", origin = 1)
ts_plot("^phi_Ma_O0[year,origin]", main = "HOR", origin = 2)
mtext(side = 1, outer = TRUE, line = 0.5, "Year")
mtext(side = 2, outer = TRUE, line = 0.5, "Value")
```

### Marine Juvenile Phase {.tabset .tabset-pills}

#### Year 1 Ocean Survival {.tabset .tabset-pills}

##### NOR

```{r}
pop_ts_plots("phi_O0_O1[year,origin,pop]", origin = 1)
```

##### HOR

```{r}
pop_ts_plots("phi_O0_O1[year,origin,pop]", origin = 1)
```

#### Age-3 Maturity {.tabset .tabset-pills}

##### NOR

```{r}
pop_ts_plots("^psi_O1_Rb[year,origin,pop]", origin = 1)
```

##### HOR

```{r}
pop_ts_plots("^psi_O1_Rb[year,origin,pop]", origin = 1)
```

#### Age-4 Maturity {.tabset .tabset-pills}

##### NOR

```{r}
pop_ts_plots("^psi_O2_Rb[year,origin,pop]", origin = 1)
```

##### HOR

```{r}
pop_ts_plots("^psi_O2_Rb[year,origin,pop]", origin = 1)
```

### Freshwater Adult Phase {.tabset .tabset-pills}

#### BON to LGR Survival

```{r}
par(mfrow = c(2,1), oma = c(2,2,0,0))
ts_plot("^phi_Rb_Ra[year,origin]", main = "NOR", origin = 1)
ts_plot("^phi_Rb_Ra[year,origin]", main = "HOR", origin = 2)
mtext(side = 1, outer = TRUE, line = 0.5, "Year")
mtext(side = 2, outer = TRUE, line = 0.5, "Value")
```

#### Pre-Spawn Survival

```{r}
pop_ts_plots("^phi_Sb_Sa[year,pop]")
```

## Emergent Properties {.tabset .tabset-pills}

### Abundance Quantities {.tabset .tabset-pills}

#### Total Adult Return (To Estuary)

```{r}
pop_ts_plots("^Rb_tot[year,pop]")
```

#### Total Adult Return (To Tributary)

```{r}
pop_ts_plots("^Ra_tot[year,pop]")
```

#### Total Spawners

```{r}
pop_ts_plots("^Sa_tot[year,pop]")
```

#### Total Egg Production

```{r}
pop_ts_plots("^f_tot[year,pop]")
```

#### Total Parr Recruitment

```{r}
pop_ts_plots("^Pb[year,pop]")
```

#### Total LGR Smolt {.tabset .tabset-pills}

##### NOR

```{r}
pop_ts_plots("Ma_tot[year,origin,pop]", origin = 1)
```

##### HOR

```{r}
pop_ts_plots("Ma_tot[year,origin,pop]", origin = 2)
```

#### Aggregate Total Adult Return (To Tributary) {.tabset .tabset-pills}

##### NOR

```{r, fig.width = 7, fig.height = 4}
ts_plot("Ra_aggregate[year,origin]", origin = 1)
```

##### HOR

```{r, fig.width = 7, fig.height = 4}
ts_plot("Ra_aggregate[year,origin]", origin = 2)
```

#### Aggregate Total Smolt (At LGR) {.tabset .tabset-pills}

##### NOR

```{r, fig.width = 7, fig.height = 4}
ts_plot("Ma_aggregate[year,origin]", origin = 1)
```

##### HOR

```{r, fig.width = 7, fig.height = 4}
ts_plot("Ma_aggregate[year,origin]", origin = 2)
```

### SAR Rates {.tabset .tabset-pills}

#### BON to BON SAR {.tabset .tabset-pills}

##### NOR

```{r}
pop_ts_plots("phi_O0_Rb_BON[year,origin,pop]", obs_y_use = obs_sar_y, sim_y_use = sim_sar_y, origin = 1)
```

##### HOR

```{r}
pop_ts_plots("phi_O0_Rb_BON[year,origin,pop]", obs_y_use = obs_sar_y, sim_y_use = sim_sar_y, origin = 2)
```

#### LGR to LGR SAR {.tabset .tabset-pills}

##### NOR

```{r}
pop_ts_plots("Ra_per_Ma[year,origin,pop]", obs_y_use = obs_sar_y, sim_y_use = sim_sar_y, origin = 1)
```

##### HOR

```{r}
pop_ts_plots("Ra_per_Ma[year,origin,pop]", obs_y_use = obs_sar_y, sim_y_use = sim_sar_y, origin = 2)
```

### Production Rates {.tabset .tabset-pills}

#### Parr per Spawner

```{r}
pop_ts_plots("Pb_per_Sa_tot[year,pop]")
```

#### Smolt per Spawner

```{r}
pop_ts_plots("Mb_per_Sa_tot[year,pop]")
```

#### Smolt in Tributary per Parr {.tabset .tabset-pills}

##### Fall Migrants

```{r}
pop_ts_plots("Mb_per_Pa[year,LH_type,pop]", LH_type = 1)
```

##### Fall Migrants

```{r}
pop_ts_plots("Mb_per_Pa[year,LH_type,pop]", LH_type = 2)
```

#### Smolt at LGR per Parr {.tabset .tabset-pills}

##### Fall Migrants

```{r}
pop_ts_plots("Ma_per_Pa[year,LH_type,pop]", LH_type = 1)
```

##### Fall Migrants

```{r}
pop_ts_plots("Ma_per_Pa[year,LH_type,pop]", LH_type = 2)
```

#### Spawners per Spawner

```{r}
pop_ts_plots("Sa_tot_per_Sa_tot[year,pop]", obs_y_use = obs_sar_y, sim_y_use = sim_sar_y)
```

### Adult Composition {.tabset .tabset-pills}

#### Proportion Hatchery Return

```{r}
pop_ts_plots("p_hor[year,pop]")
```

#### Proportion Age-3 Return {.tabset .tabset-pills}

##### NOR

```{r}
pop_ts_plots("Ra_comp[year,age,origin,pop]", age = 1, origin = 1)
```

##### HOR

```{r}
pop_ts_plots("Ra_comp[year,age,origin,pop]", age = 1, origin = 2)
```

#### Proportion Age-4 Return {.tabset .tabset-pills}

##### NOR

```{r}
pop_ts_plots("Ra_comp[year,age,origin,pop]", age = 2, origin = 1)
```

##### HOR

```{r}
pop_ts_plots("Ra_comp[year,age,origin,pop]", age = 2, origin = 2)
```

#### Proportion Age-5 Return {.tabset .tabset-pills}

##### NOR

```{r}
pop_ts_plots("Ra_comp[year,age,origin,pop]", age = 3, origin = 1)
```

##### HOR

```{r}
pop_ts_plots("Ra_comp[year,age,origin,pop]", age = 3, origin = 2)
```