### 2022-04-25 03-post-process/output-plots-children/05-length-based-values.Rmd

# Length-Based Quantities {.tabset .tabset-dropdown}

## Mean Parr Length vs. Egg Density

```{r}
plot_f = function(j) {

  # obtain posterior samples of total egg abundance scaled to WUL
  E = post_subset(post, sub_index("^E[year,pop]", pop = j, year = ts_yrs), matrix = TRUE)
  E_scaled = (E/jags_data$E_scale)/jags_data$wul[j]
  colnames(E_scaled) = gsub("E", "E_scaled", x = colnames(E_scaled))
  post_E = post_convert(cbind(postpack:::id_mat(post), E_scaled))

  # summarize scaled egg abundance
  E_scaled_mean = post_summ(post_E, "")["mean",]
  E_scaled_lwr = post_summ(post_E, "")["2.5%",]
  E_scaled_upr = post_summ(post_E, "")["97.5%",]

  # create a vector of scaled egg abundances to calculate at
  E_scaled_seq = seq(min(E_scaled_lwr), max(E_scaled_upr), length = 30)

  # extract posteriors of coefficients
  omega0 = post_subset(post, sub_index("omega0[pop]", pop = j), matrix = TRUE)
  omega1 = post_subset(post, sub_index("omega1[pop]", pop = j), matrix = TRUE)

  # extract posterior summaries of summer length outcomes
  L_Pb_mean = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["mean",]
  L_Pb_lwr = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["2.5%",]
  L_Pb_upr = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["97.5%",]

  # function to create predicted survival curves for one posterior sample
  pred_fn = function(i) {
    pred_L_Pb = exp(omega0[i] + omega1[i] * E_scaled_seq)
    names(pred_L_Pb) = paste0("pred_L_Pb[", 1:30, "1]")
    pred_L_Pb
  }

  # calculate predicted survivals and add to posterior samples
  pred_L_Pb = t(sapply(1:post_dim(post, "saved"), pred_fn))
  post_pred_L_Pb = post_convert(cbind(postpack:::id_mat(post), pred_L_Pb))

  # extract summarize predicted length
  pred_L_Pb_mean = post_summ(post_pred_L_Pb, "")["mean",]
  pred_L_Pb_lwr = post_summ(post_pred_L_Pb, "")["2.5%",]
  pred_L_Pb_upr = post_summ(post_pred_L_Pb, "")["97.5%",]

  # make a blank plot with the correct dimensions
  par(mgp = c(2,0.35, 0), tcl = -0.15, mar = c(1,1,2,1))

  plot(1,1, type = "n", 
       ylim = range(pred_L_Pb_lwr, pred_L_Pb_upr, L_Pb_lwr, L_Pb_upr, na.rm = T),
       xlim = range(0, E_scaled_seq), 
       main = pops[j]
  )

  # expected relationship
  polygon(c(E_scaled_seq, rev(E_scaled_seq)), c(pred_L_Pb_lwr, rev(pred_L_Pb_upr)), border = NA, col = alpha("salmon", 0.5))
  lines(pred_L_Pb_lwr ~ E_scaled_seq, col = "red", lty = 2)
  lines(pred_L_Pb_upr ~ E_scaled_seq, col = "red", lty = 2)
  lines(pred_L_Pb_mean ~ E_scaled_seq, col = "red", lwd = 2)

  # realized pairs
  segments(E_scaled_lwr, L_Pb_mean, E_scaled_upr, L_Pb_mean, col = alpha("red", 0.5))
  segments(E_scaled_mean, L_Pb_lwr, E_scaled_mean, L_Pb_upr, col = alpha("red", 0.5))
  points(L_Pb_mean ~ E_scaled_mean, pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)

}

xlab = paste0("Eggs/WUL (", prettyNum(jags_data$E_scale, big.mark = ","), "s per km)")
par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, xlab)
mtext(side = 2, outer = T, line = 0.25, TeX("$Mean\\,Length\\,at\\,Summer\\,Tagging\\,(mm)$"))
```

## Smolt Mean Length {.tabset .tabset-pills}

### "Growth Factor" vs. Parr Mean Length

```{r}
plot_f = function(j) {

  # extract posterior summaries of summer length outcomes
  L_Pb_mean = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["mean",]
  L_Pb_lwr = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["2.5%",]
  L_Pb_upr = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["97.5%",]

  # extract posterior summaries of growth outcomes
  Delta_L_Pb_Mb_mean = post_summ(post, sub_index("^Delta_L_Pb_Mb[year,pop]", year = ts_yrs, pop = j))["mean",]
  Delta_L_Pb_Mb_lwr = post_summ(post, sub_index("^Delta_L_Pb_Mb[year,pop]", year = ts_yrs, pop = j))["2.5%",]
  Delta_L_Pb_Mb_upr = post_summ(post, sub_index("^Delta_L_Pb_Mb[year,pop]", year = ts_yrs, pop = j))["97.5%",]

  # create a vector of scaled summer lengths to calculate at
  L_Pb_seq = seq(min(L_Pb_lwr), max(L_Pb_upr), length = 30)
  L_Pb_star_seq = (L_Pb_seq - jags_data$L_Pb_center[j])/jags_data$L_Pb_scale[j]

  # extract posteriors of coefficients
  theta0 = post_subset(post, sub_index("theta0[pop]", pop = j), matrix = TRUE)
  theta1 = post_subset(post, sub_index("theta1[pop]", pop = j), matrix = TRUE)

  # function to create predicted curves for one posterior sample
  pred_fn = function(i) {
    pred_Delta_L_Pb_Mb = exp(theta0[i] + theta1[i] * L_Pb_star_seq)
    names(pred_Delta_L_Pb_Mb) = paste0("pred_Delta_L_Pb_Mb[", 1:30, "1]")
    pred_Delta_L_Pb_Mb
  }

  # calculate predicted growth values and add to posterior samples
  pred_Delta_L_Pb_Mb = t(sapply(1:post_dim(post, "saved"), pred_fn))
  post_pred_Delta_L_Pb_Mb = post_convert(cbind(postpack:::id_mat(post), pred_Delta_L_Pb_Mb))

  # extract summarize predicted growth
  pred_Delta_L_Pb_Mb_mean = post_summ(post_pred_Delta_L_Pb_Mb, "")["mean",]
  pred_Delta_L_Pb_Mb_lwr = post_summ(post_pred_Delta_L_Pb_Mb, "")["2.5%",]
  pred_Delta_L_Pb_Mb_upr = post_summ(post_pred_Delta_L_Pb_Mb, "")["97.5%",]

  # make a blank plot with the correct dimensions
  par(mgp = c(2,0.35, 0), tcl = -0.15, mar = c(1,1,2,1))
  plot(1,1, type = "n", 
       ylim = range(pred_Delta_L_Pb_Mb_lwr, pred_Delta_L_Pb_Mb_upr, Delta_L_Pb_Mb_lwr, Delta_L_Pb_Mb_upr, na.rm = TRUE),
       xlim = range(L_Pb_seq), 
       main = pops[j]
  )

  # expected relationship
  polygon(c(L_Pb_seq, rev(L_Pb_seq)), c(pred_Delta_L_Pb_Mb_lwr, rev(pred_Delta_L_Pb_Mb_upr)), border = NA, col = alpha("salmon", 0.5))
  lines(pred_Delta_L_Pb_Mb_lwr ~ L_Pb_seq, col = "red", lty = 2)
  lines(pred_Delta_L_Pb_Mb_upr ~ L_Pb_seq, col = "red", lty = 2)
  lines(pred_Delta_L_Pb_Mb_mean ~ L_Pb_seq, col = "red", lwd = 2)

  # realized pairs
  segments(L_Pb_lwr, Delta_L_Pb_Mb_mean, L_Pb_upr, Delta_L_Pb_Mb_mean, col = alpha("red", 0.5))
  segments(L_Pb_mean, Delta_L_Pb_Mb_lwr, L_Pb_mean, Delta_L_Pb_Mb_upr, col = alpha("red", 0.5))
  points(Delta_L_Pb_Mb_mean ~ L_Pb_mean, pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)

}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, TeX("$Mean\\,Length\\,at\\,Summer\\,Tagging\\,(mm)$"))
mtext(side = 2, outer = T, line = 0.25, TeX("$Summer\\,Parr \\rightarrow Spring\\,Smolt\\,Growth\\, Factor$"))
```

### "Growth Factor" Time Series

```{r}
plot_f = function(j) {
  Delta_L_Pb_Mb_obs = cbind(mean = jags_data$L_Mb_obs[observable,j]/jags_data$L_Pb_obs[observable,j], lwr95 = NA, upr95 = NA)
  plot_tseries(post_summ(post, sub_index("^Delta_L_Pb_Mb[year,pop]$", year = ts_yrs, pop = j)),
               Delta_L_Pb_Mb_obs,
               main = pops[j],
               yrs = all_yrs[ts_yrs])
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, TeX("$Summer\\,Parr \\rightarrow Spring\\,Smolt\\,Growth\\, Factor$"))
```

## Parr $\rightarrow$ Smolt Overwinter Survival vs. Parr Mean Length

```{r overwinter-survival-plots, fig.width = 8, fig.height = 5}
plot_fn = function(j, i) {

  # extract posterior summaries of spring length outcomes
  L_Pb_mean = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["mean",]
  L_Pb_lwr = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["2.5%",]
  L_Pb_upr = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["97.5%",]

  # obtain scaled/centered versions
  L_Pb_star_mean = (L_Pb_mean - jags_data$L_Pb_center[j])/jags_data$L_Pb_scale[j]
  L_Pb_star_lwr = (L_Pb_lwr - jags_data$L_Pb_center[j])/jags_data$L_Pb_scale[j]
  L_Pb_star_upr = (L_Pb_upr - jags_data$L_Pb_center[j])/jags_data$L_Pb_scale[j]

  # summarize posterior of overwinter survival outcomes
  phi_Pa_Mb_mean = array_format(post_summ(post, sub_index("^phi_Pa_Mb[year,LH_type,pop]", pop = j, year = ts_yrs, LH_type = i))["mean",])[-1,i,j]
  phi_Pa_Mb_lwr = array_format(post_summ(post, sub_index("^phi_Pa_Mb[year,LH_type,pop]", pop = j, year = ts_yrs, LH_type = i))["2.5%",])[-1,i,j]
  phi_Pa_Mb_upr = array_format(post_summ(post, sub_index("^phi_Pa_Mb[year,LH_type,pop]", pop = j, year = ts_yrs, LH_type = i))["97.5%",])[-1,i,j]

  # create vectors to predict survival at: for credible regions and mean curve
  L_Pb_seq = seq(min(L_Pb_lwr), max(L_Pb_upr), length = 30)
  L_Pb_star_seq = seq(min(L_Pb_star_lwr), max(L_Pb_star_upr), length = 30)

  # extract posteriors of coefficients
  gamma0 = post_subset(post, sub_index("gamma0[LH_type,pop]", pop = j, LH_type = i), matrix = TRUE)
  gamma1 = post_subset(post, sub_index("gamma1[LH_type,pop]", pop = j, LH_type = i), matrix = TRUE)

  # function to create predicted survival curve for one posterior sample
  pred_fn = function(i) {
    pred_phi_Pa_Mb = plogis(gamma0[i] + gamma1[i] * L_Pb_star_seq)
    names(pred_phi_Pa_Mb) = paste0("pred_phi_Pa_Mb[", 1:30, "]")
    pred_phi_Pa_Mb
  }

  # calculate predicted survivals and add to posterior samples
  pred_phi_Pa_Mb = t(sapply(1:post_dim(post, "saved"), pred_fn))
  post_pred_phi_Pa_Mb = post_convert(cbind(postpack:::id_mat(post), pred_phi_Pa_Mb))

  # extract summarize predicted length
  pred_phi_Pa_Mb_mean = post_summ(post_pred_phi_Pa_Mb, "")["mean",]
  pred_phi_Pa_Mb_lwr = post_summ(post_pred_phi_Pa_Mb, "")["2.5%",]
  pred_phi_Pa_Mb_upr = post_summ(post_pred_phi_Pa_Mb, "")["97.5%",]

  # make a blank plot with the correct dimensions
  par(mgp = c(2,0.35, 0), tcl = -0.15, mar = c(1,1,2,1))
  plot(1,1, type = "n", 
       ylim = range(pred_phi_Pa_Mb_lwr, pred_phi_Pa_Mb_upr, phi_Pa_Mb_lwr, phi_Pa_Mb_upr, na.rm = T),
       xlim = range(L_Pb_seq), 
       main = ifelse(i == 1, pops[j], "")
  )

  # expected relationship
  polygon(c(L_Pb_seq, rev(L_Pb_seq)), c(pred_phi_Pa_Mb_lwr, rev(pred_phi_Pa_Mb_upr)), border = NA, col = alpha("salmon", 0.5))
  lines(pred_phi_Pa_Mb_lwr ~ L_Pb_seq, col = "red", lty = 2)
  lines(pred_phi_Pa_Mb_upr ~ L_Pb_seq, col = "red", lty = 2)
  lines(pred_phi_Pa_Mb_mean ~ L_Pb_seq, col = "red", lwd = 2)

  # realized pairs
  segments(L_Pb_lwr, phi_Pa_Mb_mean, L_Pb_upr, phi_Pa_Mb_mean, col = alpha("red", 0.5))
  segments(L_Pb_mean, phi_Pa_Mb_lwr, L_Pb_mean, phi_Pa_Mb_upr, col = alpha("red", 0.5))
  points(phi_Pa_Mb_mean ~ L_Pb_mean, pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)
}

plot_fn2 = function(j) {
    # create the plots for both migratory strategies
  par(mgp = c(2, 0.35, 0), tcl = -0.15, mar = c(1,1,2,1))
  plot_fn(j = j, i = 1); if (j == 1) mtext(side = 2, line = 1.5, "Fall Migrants")
  plot_fn(j = j, i = 2); if (j == 1) mtext(side = 2, line = 1.5, "Spring Migrants")
}

par(mfcol = c(jags_data$ni,jags_data$nj), oma = c(2.5,3.5,0,0))
junk = sapply(1:jags_data$nj, plot_fn2)
mtext(side = 1, line = 1.5, outer = T, "Mean Length at Summer Tagging (mm)", cex = 1.2)
mtext(side = 2, line = 2, outer = T, "Overwinter Survival", cex = 1.2)
```

## Smolt Survival to LGR vs. Smolt Mean Length

```{r toLGR-survival-plots, fig.width = 6, fig.height = 5}
plot_fn = function(j, i) {

  # extract posterior summaries of spring length outcomes
  L_Mb_mean = post_summ(post, sub_index("^L_Mb[year,pop]", year = ts_yrs, pop = j))["mean",]
  L_Mb_lwr = post_summ(post, sub_index("^L_Mb[year,pop]", year = ts_yrs, pop = j))["2.5%",]
  L_Mb_upr = post_summ(post, sub_index("^L_Mb[year,pop]", year = ts_yrs, pop = j))["97.5%",]

  # obtain scaled/centered versions
  L_Mb_star_mean = (L_Mb_mean - jags_data$L_Mb_center[j])/jags_data$L_Mb_scale[j]
  L_Mb_star_lwr = (L_Mb_lwr - jags_data$L_Mb_center[j])/jags_data$L_Mb_scale[j]
  L_Mb_star_upr = (L_Mb_upr - jags_data$L_Mb_center[j])/jags_data$L_Mb_scale[j]

  # summarize posterior of migration survival outcomes
  phi_Mb_Ma_mean = array_format(post_summ(post, sub_index("^phi_Mb_Ma[year,LH_type,origin,pop]", pop = j, year = ts_yrs, LH_type = i_spring, origin = o_nor))["mean",])[-1,i_spring,o_nor,j]
  phi_Mb_Ma_lwr = array_format(post_summ(post, sub_index("^phi_Mb_Ma[year,LH_type,origin,pop]", pop = j, year = ts_yrs, LH_type = i_spring, origin = o_nor))["2.5%",])[-1,i_spring,o_nor,j]
  phi_Mb_Ma_upr = array_format(post_summ(post, sub_index("^phi_Mb_Ma[year,LH_type,origin,pop]", pop = j, year = ts_yrs, LH_type = i_spring, origin = o_nor))["97.5%",])[-1,i_spring,o_nor,j]

  # create vectors to predict survival at: for credible regions and mean curve
  L_Mb_seq = seq(min(L_Mb_lwr), max(L_Mb_upr), length = 30)
  L_Mb_star_seq = seq(min(L_Mb_star_lwr), max(L_Mb_star_upr), length = 30)

  # extract posteriors of coefficients
  tau0 = post_subset(post, sub_index("tau0[pop]", pop = j), matrix = TRUE)
  tau1 = post_subset(post, sub_index("tau1[pop]", pop = j), matrix = TRUE)

  # function to create predicted survival curve for one posterior sample
  pred_fn = function(i) {
    pred_phi_Mb_Ma = plogis(tau0[i] + tau1[i] * L_Mb_star_seq)
    names(pred_phi_Mb_Ma) = paste0("pred_phi_Mb_Ma[", 1:30, "]")
    pred_phi_Mb_Ma
  }

  # calculate predicted survivals and add to posterior samples
  pred_phi_Mb_Ma = t(sapply(1:post_dim(post, "saved"), pred_fn))
  post_pred_phi_Mb_Ma = post_convert(cbind(postpack:::id_mat(post), pred_phi_Mb_Ma))

  # extract summarize predicted length
  pred_phi_Mb_Ma_mean = post_summ(post_pred_phi_Mb_Ma, "")["mean",]
  pred_phi_Mb_Ma_lwr = post_summ(post_pred_phi_Mb_Ma, "")["2.5%",]
  pred_phi_Mb_Ma_upr = post_summ(post_pred_phi_Mb_Ma, "")["97.5%",]

  # make a blank plot with the correct dimensions
  par(mgp = c(2,0.35, 0), tcl = -0.15, mar = c(1,1,2,1))
  plot(1,1, type = "n", 
       ylim = range(pred_phi_Mb_Ma_lwr, pred_phi_Mb_Ma_upr, phi_Mb_Ma_lwr, phi_Mb_Ma_upr, na.rm = T),
       xlim = range(L_Mb_seq), 
       main = pops[j]
  )

  # expected relationship
  polygon(c(L_Mb_seq, rev(L_Mb_seq)), c(pred_phi_Mb_Ma_lwr, rev(pred_phi_Mb_Ma_upr)), border = NA, col = alpha("salmon", 0.5))
  lines(pred_phi_Mb_Ma_lwr ~ L_Mb_seq, col = "red", lty = 2)
  lines(pred_phi_Mb_Ma_upr ~ L_Mb_seq, col = "red", lty = 2)
  lines(pred_phi_Mb_Ma_mean ~ L_Mb_seq, col = "red", lwd = 2)

  # realized pairs
  segments(L_Mb_lwr, phi_Mb_Ma_mean, L_Mb_upr, phi_Mb_Ma_mean, col = alpha("red", 0.5))
  segments(L_Mb_mean, phi_Mb_Ma_lwr, L_Mb_mean, phi_Mb_Ma_upr, col = alpha("red", 0.5))
  points(phi_Mb_Ma_mean ~ L_Mb_mean, pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_fn)
mtext(side = 1, outer = T, line = 0.25, "Length at Spring Tagging (mm)")
mtext(side = 2, outer = T, line = 0.25, "Migration to LGR Survival")
```