### 2022-04-28 03-post-process/output-plots-children/04-egg-to-parr-surv.Rmd

# Egg-to-Parr Survival {.tabset .tabset-pills}

## Beverton-Holt

All red components are posterior summaries (medians and 95% credible intervals). The black curve is from an `nls()` fit on reconstructed parr and roughly approximated total egg production -- the grey region is the confidence interval around this fit.

```{r eggs-to-parr, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  # extract posterior samples of the BH productivity and capacity parameters
  bh_params = post_subset(post, sub_index(c("alpha[pop]", "^beta[pop]$"), pop = j), matrix = TRUE)

  # extract posterior summaries of total spawners by brood year
  # Sa_tot = post_summ(post, sub_index("^Sa_tot[year,pop]", pop = j, year = ts_yrs))

  # extract posterior summaries of total egg production by brood year
  E = post_summ(post, sub_index("^E[year,pop]", pop = j, year = ts_yrs))

  # extract posterior summaries of total parr recruits by brood year
  Pb = post_summ(post, sub_index("^Pb[year,pop]", pop = j, year = ts_yrs))

  # predict parr recruits along a sequence of spawner abundances
  # Sa_pred = seq(0, max(Sa_tot[5,]), length = 30)
  E_pred = seq(0, max(E[5,]), length = 30)

  Pb_pred = t(sapply(1:post_dim(post, "saved"), function(i) BH(E_pred, bh_params[i,1], bh_params[i,2])))
  colnames(Pb_pred) = paste0("Pb_pred[", 1:length(E_pred), "]")
  Pb_pred = post_convert(cbind(postpack:::id_mat(post), Pb_pred))
  Pb_pred = post_summ(Pb_pred, "Pb_pred")

  # graphical parameters
  par(mar = c(1.5,1.5,1.5,0.5), tcl = -0.15, mgp = c(2,0.35,0))

  # empty plot
  plot(1, 1, type = "n", xlim = range(0, E[5,]), ylim = range(0, Pb[5,], Pb_pred[5,]),
       xlab = "", ylab = "", xaxt = "n", yaxt = "n", main = pops[j])

  x_ticks = axisTicks(par("usr")[1:2], log = FALSE)
  y_ticks = axisTicks(par("usr")[3:4], log = FALSE)
  axis(side = 1, at = x_ticks, labels = x_ticks/1e6)
  axis(side = 2, at = y_ticks, labels = y_ticks/1e3)

  # draw fitted relationship: Bayesian
  polygon(c(E_pred, rev(E_pred)), c(Pb_pred[4,], rev(Pb_pred[5,])), col = alpha("salmon", 0.5), border = NA)
  lines(Pb_pred[4,] ~ E_pred, col = "red", lty = 2)
  lines(Pb_pred[5,] ~ E_pred, col = "red", lty = 2)
  lines(Pb_pred[3,] ~ E_pred, col = "red", lwd = 2)

  # draw state pairs + uncertainty
  segments(E[4,], Pb[3,], E[5,], Pb[3,], col = alpha("red", 0.5))
  segments(E[3,], Pb[4,], E[3,], Pb[5,], col = alpha("red", 0.5))
  points(Pb[3,] ~ E[3,], pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Total Egg Production (Millions)")
mtext(side = 2, outer = T, line = 0.25, "Parr Recruits (Thousands)")
```

## Total Eggs vs. Total Spawners

This figure illustrates that total spawners is a reasonable index of total egg production, according to the assumptions of our model. The three lines illustrate the relationship if each spawner (regardless of age or sex) had a fixed number of eggs.

```{r eggs-vs-spawners, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  # extract posterior summaries of total spawners by brood year
  Sa_tot = post_summ(post, sub_index("^Sa_tot[year,pop]", pop = j, year = ts_yrs))

  # extract posterior summaries of total egg production by brood year
  E = post_summ(post, sub_index("^E[year,pop]", pop = j, year = ts_yrs))

  # graphical parameters
  par(mar = c(1.5,1.5,1.5,0.5), tcl = -0.15, mgp = c(2,0.35,0))

  # empty plot
  plot(E["mean",] ~ Sa_tot["mean",],
       ylim = c(0, max(E["97.5%",])),
       xlim = c(0, max(Sa_tot["97.5%",])), type = "n",
       ylab = "", xlab = "", las = 1, yaxt = "n")

  # yaxis ticks
  y_ticks = axisTicks(par("usr")[3:4], log = F)
  axis(side = 2, at = y_ticks, labels = y_ticks/1e6, las = 2)

  segments(Sa_tot["2.5%",], E["mean",], Sa_tot["97.5%",], E["mean",], col = alpha("red", 0.5))
  segments(Sa_tot["mean",], E["2.5%",], Sa_tot["mean",], E["97.5%",], col = alpha("red", 0.5))
  points(E["mean",] ~ Sa_tot["mean",], pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)
  abline(c(0, 1000), lty = 1)
  abline(c(0, 1800), lty = 2)
  abline(c(0, 3600), lty = 3)

  if (j == 1) {
    legend("bottomright", title = "If Each Spawner Had:",
           legend = c("1000 Eggs", "1800 Eggs", "3600 Eggs"),
           lty = c(1,2,3), bty = "n", cex = 0.75)
  }
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Total Spawners")
mtext(side = 2, outer = T, line = 0.25, "Total Egg Production (Millions)")
```

## Parr per Egg

This figure shows the density dependence built in to the Beverton-Holt recruitment function: with increasing egg density, survival per egg decreases. The horizontal line/region shows the theoretical maximum ($\alpha_j$) in the absence of density dependence.

```{r parr-per-egg-v-eggs, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  # extract posterior summaries of total egg production by brood year
  E = post_summ(post, sub_index("^E[year,pop]", pop = j, year = ts_yrs))

  # conversion rates
  Pb_per_E = post_summ(post, sub_index("^Pb_per_E[year,pop]", pop = j, year = ts_yrs))

  # max survival
  alpha = post_summ(post, sub_index("^alpha[pop]", pop = j))

  # graphical parameters
  par(mar = c(1.5,1.5,1.5,0.5), tcl = -0.15, mgp = c(2,0.35,0))

  # empty plot
  plot(Pb_per_E["mean",] ~ E["mean",],
       ylim = c(0, 1),
       xlim = c(0, max(E["97.5%",])), type = "n",
       ylab = "", xlab = "", las = 1, xaxt = "n", main = pops[j])

  # x-axis
  x_ticks = axisTicks(par("usr")[1:2], log = F)
  axis(side = 1, at = x_ticks, labels = x_ticks/1e6)

  segments(E["2.5%",], Pb_per_E["mean",], E["97.5%",], Pb_per_E["mean",], col = alpha("red", 0.5))
  segments(E["mean",], Pb_per_E["2.5%",], E["mean",], Pb_per_E["97.5%",], col = alpha("red", 0.5))
  points(Pb_per_E["mean",] ~ E["mean",], pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)

  rect(par("usr")[1], alpha["2.5%",], par("usr")[2], alpha["97.5%",], border = alpha("red", 0.5), col = alpha("salmon", 0.5))
  abline(h = alpha["mean",], lwd = 2, col = "red")
  box()
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Total Egg Production (Millions)")
mtext(side = 2, outer = T, line = 0.25, "Parr per Egg")
```

## Parr per Spawner

This figure is identical to above, but shows parr production per spawner instead of per egg.

```{r parr-per-spawner-v-spawners, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  # extract posterior summaries of total spawners by brood year
  Sa_tot = post_summ(post, sub_index("^Sa_tot[year,pop]", pop = j, year = ts_yrs))

  # conversion rates
  Pb_per_Sa_tot = post_summ(post, sub_index("^Pb_per_Sa_tot[year,pop]", pop = j, year = ts_yrs))

  # graphical parameters
  par(mar = c(1.5,1.5,1.5,0.5), tcl = -0.15, mgp = c(2,0.35,0))

  # empty plot
  plot(Pb_per_Sa_tot["mean",] ~ Sa_tot["mean",],
       ylim = c(0, max(Pb_per_Sa_tot["97.5%",])),
       xlim = c(0, max(Sa_tot["97.5%",])), type = "n",
       ylab = "", xlab = "", las = 1)

  segments(Sa_tot["2.5%",], Pb_per_Sa_tot["mean",], Sa_tot["97.5%",], Pb_per_Sa_tot["mean",], col = alpha("red", 0.5))
  segments(Sa_tot["mean",], Pb_per_Sa_tot["2.5%",], Sa_tot["mean",], Pb_per_Sa_tot["97.5%",], col = alpha("red", 0.5))
  points(Pb_per_Sa_tot["mean",] ~ Sa_tot["mean",], pch = 21, col = "red", bg = alpha("salmon", 0.5), cex = 1.2)
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Total Spawners")
mtext(side = 2, outer = T, line = 0.25, "Parr per Spawner")
```