### 2022-04-28 03-post-process/output-plots-children/09-rho-params.Rmd

# Correlation Parameters

```{r rho_plot, fig.width = 6, fig.height = 7}
# get the names of all rho parameters
# these are elements of a correlation matrix
# but because the rho term is the same for all population pairs, it is sufficient 
# to summarize just one element from each matrix
rho_params = c(
  match_params(post, "rho_.+[2,1]"),
  match_params(post, "rho_.+[2,1,1]"),
  match_params(post, "rho_.+[2,1,2]")
)

# calculate/format posterior summaries
rho_summs = t(post_summ(post, rho_params, probs = c(0.025, 0.25, 0.5, 0.75, 0.975)))
rho_summs = as.data.frame(rho_summs)
rho_summs = cbind(param = rownames(rho_summs), rho_summs)
rownames(rho_summs) = NULL

# make prettier names for each term
rho_names = c(
  "rho_Lphi_E_Pb[2,1]" = "Egg-to-Parr Survival",
  "rho_lL_Pb[2,1]" = "Summer Mean Length",
  "rho_lDelta_L_Pb_Mb[2,1]" = "Summer to Spring 'Growth Factor'",
  "rho_Lpi[2,1]" = "Mig. Strategy Apportionment",
  "rho_Lphi_Ma_O0[2,1]" = "LGR to Ocean Survival",
  "rho_Lphi_O0_O1[2,1]" = "Yr1 Ocean Survival",
  "rho_Lphi_Rb_Ra[2,1]" = "BON to LGR Survival",
  "rho_Lphi_Sb_Sa[2,1]" = "Pre-spawn Survival",
  "rho_Lphi_Pa_Mb[2,1,1]" = "Overwinter Survival (Fall)",
  "rho_Lphi_Pa_Mb[2,1,2]" = "Overwinter Survival (Spring)",
  "rho_Lphi_Mb_Ma[2,1,1]" = "Trib to LGR Survival (NOR)",
  "rho_Lphi_Mb_Ma[2,1,2]" = "Trib to LGR Survival (HOR)",
  "rho_Lpsi_O1[2,1,1]" = "Age-3 Maturation (NOR)",
  "rho_Lpsi_O1[2,1,2]" = "Age-3 Maturation (HOR)",
  "rho_Lpsi_O2[2,1,1]" = "Age-4 Maturation (NOR)",
  "rho_Lpsi_O2[2,1,2]" = "Age-4 Maturation (HOR)"
  )
rho_summs$name = rho_names[rho_summs$param]

# reorder by decreasing mean correlation
rho_summs = rho_summs[order(rho_summs$mean, decreasing = TRUE),]
at_y = nrow(rho_summs):1

# use different colors for correlations that are between origins vs. between populations
col = ifelse(rho_summs$param %in% c("rho_Lphi_Ma_O0[2,1]", "rho_Lphi_Rb_Ra[2,1]"), "red", "black")

# make the plot
par(lend = "square", mar = c(3,10,1,1), tcl = -0.15, mgp = c(1.5,0.35,0))
plot(1,1, type = "n", xlim = c(-1,1), ylim = c(1,nrow(rho_summs)), xlab = "Correlation", ylab = "", yaxt = "n")
abline(v = 0, lty = 2)
segments(rho_summs$`2.5%`, at_y, rho_summs$`97.5%`, col = col)
segments(rho_summs$`25%`, at_y, rho_summs$`75%`, lwd = 4, col = col)
points(y = at_y, x = rho_summs$mean, cex = 1, pch = 3, lwd = 2, col = col)
axis(side = 2, at = at_y, labels = rho_summs$name, las = 2, cex.axis = 0.8)
abline(h = at_y - 0.5, col = "grey")
legend("topleft", legend = c("Across Populations", "Across Origins"), text.col = c("black", "red"), bty = "n", cex = 0.8)
box()
```