### 2023-02-28 03-post-process/output-plots-children/13-post-vs-prior.Rmd
Add posterior vs. prior plots to output-plots
Only for the handful of parameters with potentially informative priors. Consider adding greater flexibility later

---
editor_options: 
  chunk_output_type: console
---

# Prior vs. Posterior {.tabset .tabset-pills}

```{r post-v-prior-fig-sizes}
# set global knitting options
knitr::opts_chunk$set(fig.width = 6, fig.height = 5)
```

```{r post-v-prior-fn}
# argument list:
# post_samps & prior_samps: samples from their respective PDFs
# post_col & prior_col: (non-transparent) colors for each PDF (will be made transparent for bar fills based on the alpha argument)
# main: plot title
# y_max: maximum bin probability height to display; should be [0,1]
# legend_loc: position of the legend (e.g., "topright")
# n_bins: the number of bins
# alpha: transparency setting for bar fills

post_v_prior = function(post_samps, prior_samps, prior_col = "salmon", post_col = "royalblue",
                        main = NULL, y_max = 0.3, xlim = NULL, legend_loc = NA, n_bins = 50, alpha = 0.5) {

  # get the range of all samples
  rng = range(post_samps, prior_samps)

  # if xlim was provided, take it into consideration when creating breaks
  if (is.null(xlim)) {
    breaks = seq(rng[1], rng[2], length = n_bins)
  } else {
    breaks = seq(min(rng[1], xlim[1]), max(rng[2], xlim[2]), length = n_bins)
  }

  # calculate the data for the histogram
  prior_hist = hist(prior_samps, plot = FALSE, breaks = breaks)
  post_hist = hist(post_samps, plot = FALSE, breaks = breaks)

  # convert bin counts to probabilities
  prior_probs = prior_hist$counts/sum(prior_hist$counts)
  post_probs = post_hist$counts/sum(post_hist$counts)

  # graphics parameters
  par(xaxs = "i", yaxs = "i", mgp = c(2,0.2,0), cex.axis = 0.8, tcl = -0.15, mar = c(1.5,1,1.5,0.5), lend = "square", ljoin = "mitre")

  # create a y_max if not supplied
  if (is.null(y_max)) {
    y_max = max(prior_probs, post_probs)
  }

  # create the histogram for prior PDF
  # barplot where height is equal to bin prob
  mp = barplot(
    prior_probs, ylim = c(0,y_max), space = 0, border = prior_col, col = scales::alpha(prior_col, alpha), axes = FALSE
    )

  # create the histogram for posterior PDF
  mp = barplot(
    post_probs, ylim = c(0,y_max), space = 0, border = post_col, col = scales::alpha(post_col, alpha), axes = FALSE, add = TRUE
    )

  # draw the x-axis
  at_x = axisTicks(par("usr")[1:2], log = FALSE)
  axis(side = 1, at = mp[at_x], labels = round(prior_hist$breaks[at_x], 1))

  # draw the y-axis
  axis(side = 2)

  # draw segments to connect axes at origin
  usr = par("usr")
  segments(usr[1], usr[3], usr[2], usr[3], xpd = TRUE)
  segments(usr[1], usr[3], usr[1], usr[4], xpd = TRUE)

  # add a legend in the desired location if not NULL
  if (!is.na(legend_loc)) {
    legend(legend_loc, legend = c("Prior", "Posterior"), bty = "n", pch = 22, pt.cex = 1.75, cex = 0.8, col = c(prior_col, post_col), pt.bg = scales::alpha(c(prior_col, post_col), 0.5))
  }
}
```

```{r}
# wrapper to create plot for all populations
post_v_prior_pops = function(post_param, prior_param, legend_loc = "topright", xlim = NULL, ...) {
  # function to plot one population
  plot_f = function(j) {
    post_v_prior(
      post_samps = post_subset(post, sub_index(post_param, pop = j, ...), matrix = TRUE),
      prior_samps = post_subset(post, prior_param, matrix = TRUE),
      main = pops[j], 
      legend_loc = ifelse(j == 1, legend_loc, NA),
      xlim = xlim
    )
  }

  # apply to each population
  par(mfrow = c(2,2), oma = c(1,1.5,0,0))
  junk = sapply(1:jags_data$nj, plot_f)
  mtext(side = 1, outer = TRUE, line = 0, "Parameter Value")
  mtext(side = 2, outer = TRUE, line = 0.5, "Probability")
}
```

## "Location" Parameters {.tabset .tabset-dropdown}

### Max. Egg-to-Parr Survival

**Prior Used**: `dbeta(`r jags_data$alpha_prior[1]`, `r jags_data$alpha_prior[2]`)`

```{r alpha-post-v-prior}
post_v_prior_pops(post_param = "alpha[pop]", prior_param = "alpha_pr", xlim = c(0,1))
```

### Mean Ocean Survival (NOR) {.tabset .tabset-pills}

#### Yr1

**Prior Used**: `dbeta(`r jags_data$mu_phi_O0_O1[1]`, `r jags_data$mu_phi_O0_O1[2]`)`

```{r mu_phi_O0_O1-post-v-prior}
post_v_prior_pops(post_param = "mu_phi_O0_O1[origin,pop]", prior_param = "mu_phi_O0_O1_pr",
                  origin = o_nor, xlim = c(0,1)
)
```

#### Yr2

**Prior Used**: `dbeta(`r jags_data$mu_phi_O1_O2[1]`, `r jags_data$mu_phi_O1_O2[2]`)`

```{r mu_phi_O1_O2-post-v-prior}
post_v_prior_pops(post_param = "mu_phi_O1_O2[origin,pop]", prior_param = "mu_phi_O1_O2_pr",
                  origin = o_nor, xlim = c(0,1), legend_loc = "topleft"
)
```

#### Yr3

**Prior Used**: `dbeta(`r jags_data$mu_phi_O2_O3[1]`, `r jags_data$mu_phi_O2_O3[2]`)`

```{r mu_phi_O2_O3-post-v-prior}
post_v_prior_pops(post_param = "mu_phi_O2_O3[origin,pop]", prior_param = "mu_phi_O2_O3_pr",
                  origin = o_nor, xlim = c(0,1), legend_loc = "topleft"
)
```

### Mean Maturation {.tabset .tabset-pills}

#### Age-3 {.tabset .tabset-pills}

**Prior Used**: `dbeta(`r jags_data$mu_psi_O1[1]`, `r jags_data$mu_psi_O1[2]`)`

##### NOR

```{r mu_psi_O1_nor-post-v-prior}
post_v_prior_pops(post_param = "mu_psi_O1[origin,pop]", prior_param = "mu_psi_O1_pr",
                  origin = o_nor, xlim = c(0,1)
)
```

##### HOR

```{r mu_psi_O1_hor-post-v-prior}
post_v_prior_pops(post_param = "mu_psi_O1[origin,pop]", prior_param = "mu_psi_O1_pr",
                  origin = o_hor, xlim = c(0,1)
)
```

#### Age-4 {.tabset .tabset-pills}

**Prior Used**: `dbeta(`r jags_data$mu_psi_O2[1]`, `r jags_data$mu_psi_O2[2]`)`

##### NOR

```{r mu_psi_O2_nor-post-v-prior}
post_v_prior_pops(post_param = "mu_psi_O2[origin,pop]", prior_param = "mu_psi_O2_pr",
                  origin = o_nor, xlim = c(0,1), legend_loc = "topleft"
)
```

##### HOR

```{r mu_psi_O2_hor-post-v-prior}
post_v_prior_pops(post_param = "mu_psi_O2[origin,pop]", prior_param = "mu_psi_O2_pr",
                  origin = o_hor, xlim = c(0,1), legend_loc = "topleft"
)
```

## "Scale" Parameters {.tabset .tabset-dropdown}

### SD Logit Egg-to-Parr Survival

**Prior Used**: `dt(0, 1/`r jags_data$sig_Lphi_E_Pb_prior`^2, 1) T(0, `r jags_data$dt_upr`)`

```{r sig_Lphi_E_Pb-post-v-prior}
post_v_prior_pops(post_param = "sig_Lphi_E_Pb[pop]", prior_param = "sig_Lphi_E_Pb_pr",
                  xlim = c(0, jags_data$dt_upr)
)
```

### SD Logit Yr1 Ocean Survival

**Prior Used**: `dt(0, 1/`r jags_data$sig_Lphi_O0_O1_prior`^2, 1) T(0, `r jags_data$dt_upr`)`

```{r sig_Lphi_O0_O1-post-v-prior}
post_v_prior_pops(post_param = "sig_Lphi_O0_O1[pop]", prior_param = "sig_Lphi_O0_O1_pr",
                  xlim = c(0, jags_data$dt_upr)
)
```

### SD Logit Maturation {.tabset .tabset-pills}

#### Age-3 {.tabset .tabset-pills}

**Prior Used**: `dt(0, 1/`r jags_data$sig_Lpsi_O1`^2, 1) T(0, `r jags_data$dt_upr`)`

##### NOR

```{r sig_Lpsi_O1_nor-post-v-prior}
post_v_prior_pops(post_param = "sig_Lpsi_O1[pop,origin]", prior_param = "sig_Lpsi_O1_pr",
                  origin = o_nor, xlim = c(0, jags_data$dt_upr)
)
```

##### HOR

```{r sig_Lpsi_O1_hor-post-v-prior}
post_v_prior_pops(post_param = "sig_Lpsi_O1[pop,origin]", prior_param = "sig_Lpsi_O1_pr",
                  origin = o_hor, xlim = c(0, jags_data$dt_upr)
)
```

#### Age-4 {.tabset .tabset-pills}

**Prior Used**: `dt(0, 1/`r jags_data$sig_Lpsi_O2`^2, 1) T(0, `r jags_data$dt_upr`)`

##### NOR

```{r sig_Lpsi_O2_nor-post-v-prior}
post_v_prior_pops(post_param = "sig_Lpsi_O2[pop,origin]", prior_param = "sig_Lpsi_O2_pr",
                  origin = o_nor, xlim = c(0, jags_data$dt_upr)
)
```

##### HOR

```{r sig_Lpsi_O2_hor-post-v-prior}
post_v_prior_pops(post_param = "sig_Lpsi_O2[pop,origin]", prior_param = "sig_Lpsi_O2_pr",
                  origin = o_hor, xlim = c(0, jags_data$dt_upr)
)
```