### 2023-03-14 03-post-process/output-plots-children/04-egg-to-parr-surv.Rmd

---
editor_options: 
  chunk_output_type: console
---

# Egg-to-Parr Survival {.tabset .tabset-pills}

## Survival Time Series {.tabset .tabset-pills}

These figures show the realized time series of egg-to-parr survival that include all sources of variability in red: that explained by Beverton-Holt dynamics and that not explained by it.

### Realized Alone

Just the realized time series shown.

```{r phi_E_Pb-plots, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  plot_tseries(post_summ(post, sub_index("^phi_E_Pb[year,pop]$", year = ts_yrs, pop = j)),
               yrs = all_yrs[ts_yrs],
               main = pops[j])
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, "Egg-to-Parr Survival")
```

### Realized & BH

The blue shows the time series with variability explained only by Beverton-Holt dynamics.

```{r phi_E_Pb_dot-plots, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  plot_tseries(post_summ(post, sub_index("^phi_E_Pb[year,pop]$", year = ts_yrs, pop = j)),
               yrs = all_yrs[ts_yrs],
               main = pops[j])

  # draw on the posterior time series that includes BH prediction only
  dot_col = "royalblue"
  dot = post_summ(post, sub_index("^phi_E_Pb_dot[year,pop]", pop = j, year = ts_yrs))
  polygon(x = c(all_yrs[ts_yrs], rev(all_yrs[ts_yrs])),
          y = c(dot["2.5%",], rev(dot["97.5%",])),
          col = scales::alpha(dot_col, 0.15), border = NA)
  lines(dot["mean",] ~ all_yrs[ts_yrs], col = dot_col, lwd = 2)
  lines(dot["2.5%",] ~ all_yrs[ts_yrs], col = dot_col, lty = 2)
  lines(dot["97.5%",] ~ all_yrs[ts_yrs], col = dot_col, lty = 2)
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, "Egg-to-Parr Survival")
```

### Realized & [BH + AR(1)]

The blue shows the time series with variability explained by Beverton-Holt dynamics and the carry-over residual.

```{r phi_E_Pb_dot2-plots, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  plot_tseries(post_summ(post, sub_index("^phi_E_Pb[year,pop]$", year = ts_yrs, pop = j)),
               yrs = all_yrs[ts_yrs],
               main = pops[j])

  # draw on the posterior time series that includes BH prediction and the carry over residual from last year
  dot_col = "royalblue"
  dot = post_summ(post, sub_index("^phi_E_Pb_dot2[year,pop]", pop = j, year = ts_yrs))
  polygon(x = c(all_yrs[ts_yrs], rev(all_yrs[ts_yrs])),
          y = c(dot["2.5%",], rev(dot["97.5%",])),
          col = scales::alpha(dot_col, 0.15), border = NA)
  lines(dot["mean",] ~ all_yrs[ts_yrs], col = dot_col, lwd = 2)
  lines(dot["2.5%",] ~ all_yrs[ts_yrs], col = dot_col, lty = 2)
  lines(dot["97.5%",] ~ all_yrs[ts_yrs], col = dot_col, lty = 2)
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, "Egg-to-Parr Survival")
```

## Survival vs. Total Egg Production

This is the scale the Beverton-Holt recruitment function is estimated on.

```{r egg-surv-vs-eggs, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  # extract posterior samples of the BH productivity and capacity parameters
  alpha = post_subset(post, sub_index("alpha[pop]", pop = j), matrix = TRUE)
  beta = post_subset(post, sub_index("beta[pop]", pop = j), matrix = TRUE)

  # extract posterior summaries of total egg production by brood year
  E = post_summ(post, sub_index("^E[year,pop]", pop = j, year = ts_yrs))

  # extract posterior summaries egg-to-parr survival by brood year
  phi_E_Pb = post_summ(post, sub_index("^phi_E_Pb[year,pop]", pop = j, year = ts_yrs))

  # predict egg-to-parr survival along a sequence of egg abundances
  E_pred = seq(0, max(E[5,]), length = 30)
  phi_E_Pb_pred = t(sapply(1:post_dim(post, "saved"), function(i) 1/(1/alpha[i,] + E_pred/beta[i,])))
  colnames(phi_E_Pb_pred) = paste0("phi_E_Pb_pred[", 1:length(E_pred), "]")
  phi_E_Pb_pred = post_convert(cbind(postpack:::id_mat(post), phi_E_Pb_pred))
  phi_E_Pb_pred = post_summ(phi_E_Pb_pred, "Pb_pred")

  # graphical parameters
  par(mar = c(1.5,1.5,1.5,0.5), tcl = -0.15, mgp = c(2,0.35,0))

  # empty plot
  plot(1, 1, type = "n", xlim = range(0, E[5,]), ylim = range(0, phi_E_Pb[5,], phi_E_Pb_pred[5,]),
       xlab = "", ylab = "", xaxt = "n", yaxt = "n", main = pops[j])

  x_ticks = axisTicks(par("usr")[1:2], log = FALSE)
  y_ticks = axisTicks(par("usr")[3:4], log = FALSE)
  axis(side = 1, at = x_ticks, labels = x_ticks/1e6)
  axis(side = 2, at = y_ticks)

  # draw fitted relationship
  polygon(c(E_pred, rev(E_pred)), c(phi_E_Pb_pred[4,], rev(phi_E_Pb_pred[5,])), col = alpha("salmon", 0.5), border = NA)
  lines(phi_E_Pb_pred[4,] ~ E_pred, col = "red", lty = 2)
  lines(phi_E_Pb_pred[5,] ~ E_pred, col = "red", lty = 2)
  lines(phi_E_Pb_pred[3,] ~ E_pred, col = "red", lwd = 2)

  # draw state pairs + uncertainty
  segments(E[4,], phi_E_Pb[3,], E[5,], phi_E_Pb[3,], col = alpha("red", 0.5))
  segments(E[3,], phi_E_Pb[4,], E[3,], phi_E_Pb[5,], col = alpha("red", 0.5))
  points(phi_E_Pb[3,] ~ E[3,], pch = 16, col = alpha("salmon", 0.5), cex = 1.8)
  text(phi_E_Pb[3,] ~ E[3,], labels = substr(all_yrs[ts_yrs], 3, 4), cex = 0.65, col = "black")
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Total Egg Production (Millions)")
mtext(side = 2, outer = T, line = 0.25, "Egg-to-Parr Survival")
```

## Parr vs. Total Egg Production

Parr abundance is the product of egg-to-parr survival and total egg production; in this way we can back-out the standard Beverton-Holt function.

```{r eggs-to-parr, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  # extract posterior samples of the BH productivity and capacity parameters
  bh_params = post_subset(post, sub_index(c("alpha[pop]", "^beta[pop]$"), pop = j), matrix = TRUE)

  # extract posterior summaries of total spawners by brood year
  # Sa_tot = post_summ(post, sub_index("^Sa_tot[year,pop]", pop = j, year = ts_yrs))

  # extract posterior summaries of total egg production by brood year
  E = post_summ(post, sub_index("^E[year,pop]", pop = j, year = ts_yrs))

  # extract posterior summaries of total parr recruits by brood year
  Pb = post_summ(post, sub_index("^Pb[year,pop]", pop = j, year = ts_yrs))

  # predict parr recruits along a sequence of spawner abundances
  # Sa_pred = seq(0, max(Sa_tot[5,]), length = 30)
  E_pred = seq(0, max(E[5,]), length = 30)

  Pb_pred = t(sapply(1:post_dim(post, "saved"), function(i) BH(E_pred, bh_params[i,1], bh_params[i,2])))
  colnames(Pb_pred) = paste0("Pb_pred[", 1:length(E_pred), "]")
  Pb_pred = post_convert(cbind(postpack:::id_mat(post), Pb_pred))
  Pb_pred = post_summ(Pb_pred, "Pb_pred")

  # graphical parameters
  par(mar = c(1.5,1.5,1.5,0.5), tcl = -0.15, mgp = c(2,0.35,0))

  # empty plot
  plot(1, 1, type = "n", xlim = range(0, E[5,]), ylim = range(0, Pb[5,], Pb_pred[5,]),
       xlab = "", ylab = "", xaxt = "n", yaxt = "n", main = pops[j])

  x_ticks = axisTicks(par("usr")[1:2], log = FALSE)
  y_ticks = axisTicks(par("usr")[3:4], log = FALSE)
  axis(side = 1, at = x_ticks, labels = x_ticks/1e6)
  axis(side = 2, at = y_ticks, labels = y_ticks/1e3)

  # draw fitted relationship: Bayesian
  polygon(c(E_pred, rev(E_pred)), c(Pb_pred[4,], rev(Pb_pred[5,])), col = alpha("salmon", 0.5), border = NA)
  lines(Pb_pred[4,] ~ E_pred, col = "red", lty = 2)
  lines(Pb_pred[5,] ~ E_pred, col = "red", lty = 2)
  lines(Pb_pred[3,] ~ E_pred, col = "red", lwd = 2)

  # draw state pairs + uncertainty
  segments(E[4,], Pb[3,], E[5,], Pb[3,], col = alpha("red", 0.5))
  segments(E[3,], Pb[4,], E[3,], Pb[5,], col = alpha("red", 0.5))
  points(Pb[3,] ~ E[3,], pch = 16, col = alpha("salmon", 0.5), cex = 1.8)
  text(Pb[3,] ~ E[3,], labels = substr(all_yrs[ts_yrs], 3, 4), cex = 0.65, col = "black")
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Total Egg Production (Millions)")
mtext(side = 2, outer = T, line = 0.25, "Parr Recruits (Thousands)")
```

## Total Eggs vs. Total Spawners

This figure illustrates that total spawners is a reasonable index of total egg production, according to the assumptions of our model. The three lines illustrate the relationship if each spawner (regardless of age or sex) had a fixed number of eggs.

```{r eggs-vs-spawners, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  # extract posterior summaries of total spawners by brood year
  Sa_tot = post_summ(post, sub_index("^Sa_tot[year,pop]", pop = j, year = ts_yrs))

  # extract posterior summaries of total egg production by brood year
  E = post_summ(post, sub_index("^E[year,pop]", pop = j, year = ts_yrs))

  # graphical parameters
  par(mar = c(1.5,1.5,1.5,0.5), tcl = -0.15, mgp = c(2,0.35,0))

  # empty plot
  plot(E["mean",] ~ Sa_tot["mean",],
       ylim = c(0, max(E["97.5%",])),
       xlim = c(0, max(Sa_tot["97.5%",])), type = "n",
       ylab = "", xlab = "", las = 1, yaxt = "n")

  # yaxis ticks
  y_ticks = axisTicks(par("usr")[3:4], log = F)
  axis(side = 2, at = y_ticks, labels = y_ticks/1e6, las = 2)

  segments(Sa_tot["2.5%",], E["mean",], Sa_tot["97.5%",], E["mean",], col = alpha("red", 0.5))
  segments(Sa_tot["mean",], E["2.5%",], Sa_tot["mean",], E["97.5%",], col = alpha("red", 0.5))
  points(E["mean",] ~ Sa_tot["mean",], pch = 16, col = alpha("salmon", 0.5), cex = 1.8)
  text(E["mean",] ~ Sa_tot["mean",], labels = substr(all_yrs[ts_yrs], 3, 4), cex = 0.65, col = "black")

  abline(c(0, 1000), lty = 1)
  abline(c(0, 1800), lty = 2)
  abline(c(0, 3600), lty = 3)

  if (j == 1) {
    legend("bottomright", title = "If Each Spawner Had:",
           legend = c("1000 Eggs", "1800 Eggs", "3600 Eggs"),
           lty = c(1,2,3), bty = "n", cex = 0.75)
  }
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Total Spawners")
mtext(side = 2, outer = T, line = 0.25, "Total Egg Production (Millions)")
```