### 2023-04-12 03-post-process/output-plots-children/05-length.Rmd
Update file names on output-plots children

---
editor_options: 
  chunk_output_type: console
---

# Parr & Smolt Mean Length {.tabset .tabset-pills}

## Parr Mean Length {.tabset .tabset-pills}

### Time Series {.tabset .tabset-pills}

#### Realized Alone

```{r L_Pb_tseries-1, fig.width = 6, fig.height = 5}
plot_f = function(j) {

  # draw the realized time series
  plot_tseries(est = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j)), 
               main = pops[j], yrs = all_yrs[ts_yrs]
  )
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, "Parr Mean Length (mm)")
```

#### Realized & Predicted

```{r L_Pb_tseries-2, fig.width = 6, fig.height = 5}
plot_f = function(j) {

  # draw the realized time series
  plot_tseries(est = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j)), 
               main = pops[j], yrs = all_yrs[ts_yrs]
  )

  # draw the predicted time series
  add_tseries(est = post_summ(post, sub_index("^L_Pb_dot[year,pop]", pop = j, year = ts_yrs)),
              yrs = all_yrs[ts_yrs])
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, "Parr Mean Length (mm)")
```

### Density-Dependent Relationship

```{r L_Pb-relationship, fig.width = 6, fig.height = 5}
plot_f = function(j) {

  # obtain posterior samples of total egg abundance scaled to WUL
  E = post_subset(post, sub_index("^E[year,pop]", pop = j, year = ts_yrs), matrix = TRUE)
  E_scaled = (E/jags_data$E_scale)/jags_data$wul[j]
  colnames(E_scaled) = gsub("E", "E_scaled", x = colnames(E_scaled))
  post_E = post_convert(cbind(postpack:::id_mat(post), E_scaled))

  # summarize scaled egg abundance
  E_scaled_mean = post_summ(post_E, ".")["mean",]
  E_scaled_lwr = post_summ(post_E, ".")["2.5%",]
  E_scaled_upr = post_summ(post_E, ".")["97.5%",]

  # create a vector of scaled egg abundances to calculate at
  E_scaled_seq = seq(min(E_scaled_lwr), max(E_scaled_upr), length = 30)

  # extract posteriors of coefficients
  omega0 = post_subset(post, sub_index("omega0[pop]", pop = j), matrix = TRUE)
  omega1 = post_subset(post, sub_index("omega1[pop]", pop = j), matrix = TRUE)

  # extract posterior summaries of summer length outcomes
  L_Pb_mean = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["mean",]
  L_Pb_lwr = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["2.5%",]
  L_Pb_upr = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["97.5%",]

  # function to create predicted survival curves for one posterior sample
  pred_fn = function(i) {
    pred_L_Pb = exp(omega0[i] + omega1[i] * log(E_scaled_seq))
    names(pred_L_Pb) = paste0("pred_L_Pb[", 1:30, "1]")
    pred_L_Pb
  }

  # calculate predicted survivals and add to posterior samples
  pred_L_Pb = t(sapply(1:post_dim(post, "saved"), pred_fn))
  post_pred_L_Pb = post_convert(cbind(postpack:::id_mat(post), pred_L_Pb))

  # extract summarize predicted length
  pred_L_Pb_mean = post_summ(post_pred_L_Pb, ".")["mean",]
  pred_L_Pb_lwr = post_summ(post_pred_L_Pb, ".")["2.5%",]
  pred_L_Pb_upr = post_summ(post_pred_L_Pb, ".")["97.5%",]

  # make a blank plot with the correct dimensions
  par(mgp = c(2,0.35, 0), tcl = -0.15, mar = c(1,1,2,1))

  plot(1,1, type = "n", 
       ylim = range(pred_L_Pb_lwr, pred_L_Pb_upr, L_Pb_lwr, L_Pb_upr, na.rm = T),
       xlim = range(0, E_scaled_seq), 
       main = pops[j]
  )

  # expected relationship
  polygon(c(E_scaled_seq, rev(E_scaled_seq)), c(pred_L_Pb_lwr, rev(pred_L_Pb_upr)), border = NA, col = alpha("salmon", 0.5))
  lines(pred_L_Pb_lwr ~ E_scaled_seq, col = "red", lty = 2)
  lines(pred_L_Pb_upr ~ E_scaled_seq, col = "red", lty = 2)
  lines(pred_L_Pb_mean ~ E_scaled_seq, col = "red", lwd = 2)

  # realized pairs
  segments(E_scaled_lwr, L_Pb_mean, E_scaled_upr, L_Pb_mean, col = alpha("red", 0.5))
  segments(E_scaled_mean, L_Pb_lwr, E_scaled_mean, L_Pb_upr, col = alpha("red", 0.5))
  points(L_Pb_mean ~ E_scaled_mean, pch = 16, col = alpha("salmon", 0.5), cex = 1.8)
  text(L_Pb_mean ~ E_scaled_mean, labels = substr(all_yrs[ts_yrs], 3, 4), cex = 0.65, col = "black")
}

xlab = paste0("Eggs/WUL (", prettyNum(jags_data$E_scale, big.mark = ","), "s per km)")
par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, xlab)
mtext(side = 2, outer = T, line = 0.25, TeX("$Parr\\, Mean\\, Length\\,(mm)$"))
```

## Smolt Mean Length {.tabset .tabset-pills}

### Time Series {.tabset .tabset-pills}

#### Realized Alone

```{r L_Mb-tseries-1, fig.width = 6, fig.height = 5}
plot_f = function(j) {

  # draw the realized time series
  plot_tseries(est = post_summ(post, sub_index("^L_Mb[year,pop]", year = ts_yrs, pop = j)), 
               main = pops[j], yrs = all_yrs[ts_yrs]
  )
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, "Smolt Mean Length (mm)")
```

#### Realized & Predicted

```{r get-L_Mb_dot}
# function to get posterior samples of expected smolt length
# not calculated in JAGS model, add to post object here
f = function() {
  # extract posterior samples of parr mean length and parr to smolt growth factor
  L_Pb = post_subset(post, "^L_Pb[", matrix = TRUE)
  Delta_L_Pb_Mb_dot = post_subset(post, "^Delta_L_Pb_Mb_dot[", matrix = TRUE)

  # perform basic operations for one iteration -- for making dimension names later
  L_Pb_1 = array_format(L_Pb[1,])
  Delta_L_Pb_Mb_dot_1 = array_format(Delta_L_Pb_Mb_dot[1,])
  L_Mb_dot_1 = L_Pb_1 * Delta_L_Pb_Mb_dot_1

  # function to perform the calculation for one MCMC sample
  g = function(i) {
    L_Pb_i = array_format(L_Pb[i,])
    Delta_L_Pb_Mb_dot_i = array_format(Delta_L_Pb_Mb_dot[i,])
    L_Mb_dot_i = L_Pb_i * Delta_L_Pb_Mb_dot_i
    as.numeric(L_Mb_dot_i)
  }

  # apply to each posterior sample
  out = t(sapply(1:nrow(L_Pb), g))

  # build element indices: the year
  y_ind = rep(1:nrow(L_Pb_1), ncol(L_Pb_1))

  # build element indices: the population
  j_ind = rep(1:ncol(L_Pb_1), each = nrow(L_Pb_1))

  # make element names
  colnames(out) = paste0("L_Mb_dot[", y_ind, ",", j_ind, "]")

  # drop out those that have NA values (first k_max years)
  na_ind = which(!is.na(out[1,]))
  out[,na_ind]
}

# combine with the rest of the posterior samples
post = post_bind(post, f()); rm(f)
```

```{r L_Mb-tseries-2, fig.width = 6, fig.height = 5}
plot_f = function(j) {

  # draw the realized time series
  plot_tseries(est = post_summ(post, sub_index("^L_Mb[year,pop]", year = ts_yrs, pop = j)), 
               main = pops[j], yrs = all_yrs[ts_yrs]
  )

  # draw the realized time series
  add_tseries(est = post_summ(post, sub_index("^L_Mb_dot[year,pop]", year = ts_yrs, pop = j)),
              yrs = all_yrs[ts_yrs])
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, "Smolt Mean Length (mm)")
```

## Parr $\to$ Smolt "Growth" Factor {.tabset .tabset-pills}

### Time Series {.tabset .tabset-pills}

#### Realized Alone

```{r Delta_L_Pb_Mb-1, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  plot_tseries(post_summ(post, sub_index("^Delta_L_Pb_Mb[year,pop]$", year = ts_yrs, pop = j)),
               main = pops[j], yrs = all_yrs[ts_yrs], ylim = c(1,2))
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, TeX("$Parr \\rightarrow Smolt\\,Growth\\, Factor$"))
```

#### Realized & "Data"

```{r Delta_L_Pb_Mb-2, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  Delta_L_Pb_Mb_obs = cbind(mean = jags_data$L_Mb_obs[observable,j]/jags_data$L_Pb_obs[observable,j], lwr95 = NA, upr95 = NA)
  plot_tseries(post_summ(post, sub_index("^Delta_L_Pb_Mb[year,pop]$", year = ts_yrs, pop = j)),
               Delta_L_Pb_Mb_obs, main = pops[j], yrs = all_yrs[ts_yrs], ylim = c(1,2))
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, TeX("$Parr \\rightarrow Smolt\\,Growth\\, Factor$"))
```

#### Realized & Predicted

```{r Delta_L_Pb_Mb-3, fig.width = 6, fig.height = 5}
plot_f = function(j) {
  # draw the realized time series
  plot_tseries(post_summ(post, sub_index("^Delta_L_Pb_Mb[year,pop]$", year = ts_yrs, pop = j)),
               main = pops[j], yrs = all_yrs[ts_yrs], ylim = c(1,2))

  # draw the predicted time series
  add_tseries(est = post_summ(post, sub_index("^Delta_L_Pb_Mb_dot[year,pop]", pop = j, year = ts_yrs)),
              yrs = all_yrs[ts_yrs])
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, "Brood Year")
mtext(side = 2, outer = T, line = 0.25, TeX("$Parr \\rightarrow Smolt\\,Growth\\, Factor$"))
```

### Relationship

```{r Delta_L_Pb_Mb-relationship, fig.width = 6, fig.height = 5}
plot_f = function(j) {

  # extract posterior summaries of summer length outcomes
  L_Pb_mean = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["mean",]
  L_Pb_lwr = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["2.5%",]
  L_Pb_upr = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))["97.5%",]

  # extract posterior summaries of growth outcomes
  Delta_L_Pb_Mb_mean = post_summ(post, sub_index("^Delta_L_Pb_Mb[year,pop]", year = ts_yrs, pop = j))["mean",]
  Delta_L_Pb_Mb_lwr = post_summ(post, sub_index("^Delta_L_Pb_Mb[year,pop]", year = ts_yrs, pop = j))["2.5%",]
  Delta_L_Pb_Mb_upr = post_summ(post, sub_index("^Delta_L_Pb_Mb[year,pop]", year = ts_yrs, pop = j))["97.5%",]

  # create a vector of scaled summer lengths to calculate at
  L_Pb_seq = seq(min(L_Pb_lwr), max(L_Pb_upr), length = 30)
  L_Pb_star_seq = (L_Pb_seq - jags_data$L_Pb_center[j])/jags_data$L_Pb_scale[j]

  # extract posteriors of coefficients
  theta0 = post_subset(post, sub_index("theta0[pop]", pop = j), matrix = TRUE)
  theta1 = post_subset(post, sub_index("theta1[pop]", pop = j), matrix = TRUE)

  # function to create predicted curves for one posterior sample
  pred_fn = function(i) {
    pred_Delta_L_Pb_Mb = exp(theta0[i] + theta1[i] * L_Pb_star_seq)
    names(pred_Delta_L_Pb_Mb) = paste0("pred_Delta_L_Pb_Mb[", 1:30, "1]")
    pred_Delta_L_Pb_Mb
  }

  # calculate predicted growth values and add to posterior samples
  pred_Delta_L_Pb_Mb = t(sapply(1:post_dim(post, "saved"), pred_fn))
  post_pred_Delta_L_Pb_Mb = post_convert(cbind(postpack:::id_mat(post), pred_Delta_L_Pb_Mb))

  # extract summarize predicted growth
  pred_Delta_L_Pb_Mb_mean = post_summ(post_pred_Delta_L_Pb_Mb, ".")["mean",]
  pred_Delta_L_Pb_Mb_lwr = post_summ(post_pred_Delta_L_Pb_Mb, ".")["2.5%",]
  pred_Delta_L_Pb_Mb_upr = post_summ(post_pred_Delta_L_Pb_Mb, ".")["97.5%",]

  # make a blank plot with the correct dimensions
  par(mgp = c(2,0.35, 0), tcl = -0.15, mar = c(1,1,2,1))
  plot(1,1, type = "n", 
       ylim = range(pred_Delta_L_Pb_Mb_lwr, pred_Delta_L_Pb_Mb_upr, Delta_L_Pb_Mb_lwr, Delta_L_Pb_Mb_upr, na.rm = TRUE),
       xlim = range(L_Pb_seq), 
       main = pops[j]
  )

  # expected relationship
  polygon(c(L_Pb_seq, rev(L_Pb_seq)), c(pred_Delta_L_Pb_Mb_lwr, rev(pred_Delta_L_Pb_Mb_upr)), border = NA, col = alpha("salmon", 0.5))
  lines(pred_Delta_L_Pb_Mb_lwr ~ L_Pb_seq, col = "red", lty = 2)
  lines(pred_Delta_L_Pb_Mb_upr ~ L_Pb_seq, col = "red", lty = 2)
  lines(pred_Delta_L_Pb_Mb_mean ~ L_Pb_seq, col = "red", lwd = 2)

  # realized pairs
  segments(L_Pb_lwr, Delta_L_Pb_Mb_mean, L_Pb_upr, Delta_L_Pb_Mb_mean, col = alpha("red", 0.5))
  segments(L_Pb_mean, Delta_L_Pb_Mb_lwr, L_Pb_mean, Delta_L_Pb_Mb_upr, col = alpha("red", 0.5))
  points(Delta_L_Pb_Mb_mean ~ L_Pb_mean, pch = 16, col = alpha("salmon", 0.5), cex = 1.8)
  text(Delta_L_Pb_Mb_mean ~ L_Pb_mean, labels = substr(all_yrs[ts_yrs], 3, 4), cex = 0.65, col = "black")
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_f)
mtext(side = 1, outer = T, line = 0.25, TeX("$Parr\\, Mean\\, Length\\, (mm)$"))
mtext(side = 2, outer = T, line = 0.25, TeX("$Parr\\, \\to\\, Smolt\\, Growth\\, Factor$"))
```

## Smolt vs. Parr Mean Length

```{r L_Mb-v-L_Pb, fig.width = 6, fig.height = 5}
plot_fn = function(j) {

  # extract posterior summaries of parr mean length outcomes
  L_Pb_summ = post_summ(post, sub_index("^L_Pb[year,pop]", year = ts_yrs, pop = j))
  L_Pb_mean = L_Pb_summ["mean",]
  L_Pb_lwr = L_Pb_summ["2.5%",]
  L_Pb_upr = L_Pb_summ["97.5%",]

  # extract posterior summaries of parr mean length outcomes
  L_Mb_summ = post_summ(post, sub_index("^L_Mb[year,pop]", year = ts_yrs, pop = j))
  L_Mb_mean = L_Mb_summ["mean",]
  L_Mb_lwr = L_Mb_summ["2.5%",]
  L_Mb_upr = L_Mb_summ["97.5%",]

  # obtain scaled/centered versions
  L_Pb_star_mean = (L_Pb_mean - jags_data$L_Pb_center[j])/jags_data$L_Pb_scale[j]
  L_Pb_star_lwr = (L_Pb_lwr - jags_data$L_Pb_center[j])/jags_data$L_Pb_scale[j]
  L_Pb_star_upr = (L_Pb_upr - jags_data$L_Pb_center[j])/jags_data$L_Pb_scale[j]

  # create vectors to predict smolt mean length at: for credible regions and mean curve
  L_Pb_seq = seq(min(L_Pb_lwr), max(L_Pb_upr), length = 30)
  L_Pb_star_seq = seq(min(L_Pb_star_lwr), max(L_Pb_star_upr), length = 30)

  # extract posteriors of coefficients
  theta0 = post_subset(post, sub_index("theta0[pop]", pop = j), matrix = TRUE)
  theta1 = post_subset(post, sub_index("theta1[pop]", pop = j), matrix = TRUE)

  # function to create predicted survival curve for one posterior sample
  pred_fn = function(i) {
    pred_Delta_L_Pb_Mb = exp(theta0[i] + theta1[i] * L_Pb_star_seq)
    pred_L_Mb = L_Pb_seq * pred_Delta_L_Pb_Mb
    names(pred_L_Mb) = paste0("pred_L_Mb[", 1:30, "]")
    pred_L_Mb
  }

  # calculate predicted survivals and add to posterior samples
  pred_L_Mb = t(sapply(1:post_dim(post, "saved"), pred_fn))
  post_pred_L_Mb = post_convert(cbind(postpack:::id_mat(post), pred_L_Mb))

  # extract summarize predicted length
  pred_L_Mb_summ = post_summ(post_pred_L_Mb, ".")
  pred_L_Mb_mean = pred_L_Mb_summ["mean",]
  pred_L_Mb_lwr = pred_L_Mb_summ["2.5%",]
  pred_L_Mb_upr = pred_L_Mb_summ["97.5%",]

  # make a blank plot with the correct dimensions
  par(mgp = c(2,0.35, 0), tcl = -0.15, mar = c(1,1,2,1))
  plot(1,1, type = "n", 
       ylim = range(pred_L_Mb_lwr, pred_L_Mb_upr, L_Mb_lwr, L_Mb_upr, na.rm = T),
       xlim = range(L_Pb_seq), 
       main = pops[j]
  )

  # expected relationship
  polygon(c(L_Pb_seq, rev(L_Pb_seq)), c(pred_L_Mb_lwr, rev(pred_L_Mb_upr)), border = NA, col = alpha("salmon", 0.5))
  lines(pred_L_Mb_lwr ~ L_Pb_seq, col = "red", lty = 2)
  lines(pred_L_Mb_upr ~ L_Pb_seq, col = "red", lty = 2)
  lines(pred_L_Mb_mean ~ L_Pb_seq, col = "red", lwd = 2)

  # realized pairs
  segments(L_Pb_lwr, L_Mb_mean, L_Pb_upr, L_Mb_mean, col = alpha("red", 0.5))
  segments(L_Pb_mean, L_Mb_lwr, L_Pb_mean, L_Mb_upr, col = alpha("red", 0.5))
  points(L_Mb_mean ~ L_Pb_mean, pch = 16, col = alpha("salmon", 0.5), cex = 1.8)
  text(L_Mb_mean ~ L_Pb_mean, labels = substr(all_yrs[ts_yrs], 3, 4), cex = 0.65, col = "black")
}

par(mfrow = c(2,2), oma = c(1.5,1.5,0,0))
junk = sapply(1:jags_data$nj, plot_fn)
mtext(side = 1, outer = T, line = 0.25, "Parr Mean Length (mm)")
mtext(side = 2, outer = T, line = 0.25, "Smolt Mean Length (mm)")
```