### 2023-05-02 03-post-process/output-plots-children/03-fit-to-data.Rmd
Consistency updates: 03-fit-to-data.Rmd

---
editor_options: 
  chunk_output_type: console
---

# Fit to Data {.tabset .tabset-pills}

## Figures {.tabset .tabset-dropdown}

In all figures below, the thick red line is the posterior median value each year, the red region is the equal-tailed 95% credible interval, and blue points are "observed" data (estimated from raw data external to the model).

### Abundance Data {.tabset .tabset-pills}

#### Fall Parr Abundance

```{r fall-trap-N-fit, fig.width = p4_width, fig.height = p4_height}
# function to create the plot for one population
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^Pa[year,LH_type,pop]$", year = ts_yrs, LH_type = i_fall, pop = j)),
               obs = get_obs_ests_log_normal(log(jags_data$Pa_obs[observable,i_fall,j]), jags_data$sig_Pa_obs[observable,i_fall,j]),
               yrs = all_yrs[ts_yrs], label_text = pops[j], y_scale = 1000, pt_cex = pt_cex)
}

# loop over populations
mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Fall Parr Abundance (1000s)")
```

#### Spring Smolt Abundance

```{r spring-trap-N-fit, fig.width = p4_width, fig.height = p4_height}
# function to create the plot for one population
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^Mb[year,LH_type,origin,pop]$", year = ts_yrs, origin = o_nor, LH_type = i_spring, pop = j)), 
               obs = get_obs_ests_log_normal(log(jags_data$Mb_obs[observable,i_spring,o_nor,j]), jags_data$sig_Mb_obs[observable,i_spring,o_nor,j]),
               yrs = all_yrs[ts_yrs], label_text = pops[j], y_scale = 1000, pt_cex = pt_cex)
}

# loop over populations
mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Spring Smolt Abundance (1000s)")
```

#### Adult Return Abundance

```{r adult-N-fit, fig.width = p4_width, fig.height = p4_height}
# function to create the plot for one population
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^Ra_tot[year,pop]$", year = ts_yrs, pop = j)),
               obs = get_obs_ests_log_normal(log(jags_data$Ra_obs[observable,j]), jags_data$sig_Ra_obs[observable,j]),
               yrs = all_yrs[ts_yrs], label_text = pops[j], y_scale = 1000, pt_cex = pt_cex)

}

# loop over populations
mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Return Year", "Adult Return Abundance (1000s)")
```

### Survival Data {.tabset .tabset-dropdown}

#### Summer Parr $\rightarrow$ LGR

```{r summer-surv-fit, fig.width = p4_width, fig.height = p4_height}
# function to create the plot for one population
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^phi_Pb_Ma[year,pop]", year = ts_yrs, pop = j)),
               obs = get_obs_ests_logit_normal(jags_data$Lphi_obs_Pb_Ma[observable,j], jags_data$sig_Lphi_obs_Pb_Ma[observable,j]),
               label_text = pops[j], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
}

# loop over populations
mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Summer Parr \u2192 LGR Survival")
```

#### Fall Parr $\rightarrow$ LGR

```{r fall-surv-fit, fig.width = p4_width, fig.height = p4_height}
# function to create the plot for one population
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^phi_Pa_Ma[year,LH_type,pop]$", year = ts_yrs, LH_type = i_fall, pop = j)),
               obs = get_obs_ests_logit_normal(jags_data$Lphi_obs_Pa_Ma[observable,i_fall,j], jags_data$sig_Lphi_obs_Pa_Ma[observable,i_fall,j]),
               label_text = pops[j], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
}

# loop over populations
mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Fall Parr \u2192 LGR Survival")
```

#### Winter Parr $\rightarrow$ LGR

```{r winter-surv-fit, fig.width = p4_width, fig.height = p4_height}
# function to create the plot for one population
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^phi_Pa_Ma[year,LH_type,pop]$", year = ts_yrs, LH_type = i_spring, pop = j)),
               obs = get_obs_ests_logit_normal(jags_data$Lphi_obs_Pa_Ma[observable,i_spring,j], jags_data$sig_Lphi_obs_Pa_Ma[observable,i_spring,j]),
               label_text = pops[j], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
}

# loop over populations
mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Winter Parr \u2192 LGR Survival")
```

#### Spring Smolt $\rightarrow$ LGR (NOR)

```{r spring-surv-fit-nor, fig.width = p4_width, fig.height = p4_height}
# function to create the plot for one population
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^phi_Mb_Ma[year,LH_type,origin,pop]$", year = ts_yrs, LH_type = i_spring, origin = o_nor, pop = j)),
               obs = get_obs_ests_logit_normal(jags_data$Lphi_obs_Mb_Ma[observable,i_spring,o_nor,j], jags_data$sig_Lphi_obs_Mb_Ma[observable,i_spring,o_nor,j]),
               label_text = pops[j], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
}

# loop over populations
mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Spring Smolt \u2192 LGR Survival (NOR)")
```

#### Spring Smolt $\rightarrow$ LGR (HOR)

```{r spring-surv-fit-hor, fig.width = p4_width, fig.height = p4_height}
# function to create the plot for one population
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^phi_Mb_Ma[year,LH_type,origin,pop]$", year = ts_yrs, LH_type = i_spring, origin = o_hor, pop = j)),
               obs = get_obs_ests_logit_normal(jags_data$Lphi_obs_Mb_Ma[observable,i_spring,o_hor,j], jags_data$sig_Lphi_obs_Mb_Ma[observable,i_spring,o_hor,j]),
               label_text = pops[j], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
}

# loop over populations
mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Spring Smolt \u2192 LGR Survival (HOR)")
```

#### LGR Smolt $\rightarrow$ BON

```{r hydro-down-surv-fit, fig.width = p2_width, fig.height = p2_height}
# function to create the plot for one origin type
plot_f = function(o) {
  plot_tseries(est = post_summ(post, sub_index("^phi_Ma_O0[year,origin]$", year = ts_yrs, origin = o)),
               obs = get_obs_ests_logit_normal(jags_data$Lphi_obs_Ma_O0[observable,o], jags_data$sig_Lphi_obs_Ma_O0[observable,o]),
               label_text = c("NOR", "HOR")[o], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
}

# loop over origins
mypar(mfrow = c(1,2))
junk = sapply(1:jags_data$no, plot_f)
axis_labels("Brood Year", "LGR Smolt \u2192 BON")
```

#### BON Adults $\rightarrow$ LGR

```{r hydro-up-nor-surv-fit, fig.width = p2_width, fig.height = p2_height}
# function to create the plot for one origin type
plot_f = function(o) {
  plot_tseries(est = post_summ(post, sub_index("^phi_Rb_Ra[year,origin]$", year = ts_yrs, origin = o)),
               obs = get_obs_ests_multinomial(cbind(jags_data$x_LGR[observable,o], jags_data$x_BON[observable,o] - jags_data$x_LGR[observable,o]), 1),
               label_text = c("NOR", "HOR")[o], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
}

# loop over origins
mypar(mfrow = c(1,2))
junk = sapply(1:jags_data$no, plot_f)
axis_labels("Brood Year", "BON \u2192 LGR Survival")
```

### Composition Data {.tabset .tabset-pills}

```{r calculate-aggregated-comps-model}
# extract the full posteriors of the two composition sets
p_Ra_full = post_subset(post, "p_Ra", matrix = TRUE)
p_Sa_prime_full = post_subset(post, "p_Sa", matrix = TRUE)

# primary containers: will store final recalculated output
p_Ra_age = p_Sa_prime_age = p_Ra_origin = p_Sa_prime_origin = NULL

# number of posterior samples
n = post_dim(post, "saved")

# loop through posterior samples, calculate various aggregates of composition by type
for (i in 1:n) {

  # format the posterior draw as arrays -- easier to subset
  p_Ra = array_format(p_Ra_full[i,])
  p_Sa_prime = array_format(p_Sa_prime_full[i,])

  ## BY AGE, AGGREGATED ACROSS ORIGINS

  # containers
  p_Ra_age_tmp = p_Sa_prime_age_tmp = array(NA, dim = c(jags_data$ny, jags_data$nk, jags_data$nj))
  inds_1_age = inds_2_age = inds_3_age = p_Ra_age_tmp
  for (j in 1:jags_data$nj) {
    for (k in 1:jags_data$nk) {
      # calculate aggregate proportions for each age
      p_Ra_age_tmp[,k,j] = rowSums(p_Ra[,ko_age[[k]],j])
      p_Sa_prime_age_tmp[,k,j] = rowSums(p_Sa_prime[,ko_age[[k]],j])

      # build the element identifiers - for building names later
      inds_1_age[,k,j] = 1:jags_data$ny
      inds_2_age[,k,j] = k
      inds_3_age[,k,j] = j
    }
  }

  # vectorize, add names, and remove NAs
  p_Ra_age_v = as.numeric(p_Ra_age_tmp)
  p_Sa_prime_age_v = as.numeric(p_Sa_prime_age_tmp)
  names(p_Ra_age_v) = paste0("p_Ra_age[", as.numeric(inds_1_age), ",", as.numeric(inds_2_age), ",", as.numeric(inds_3_age), "]")
  names(p_Sa_prime_age_v) = paste0("p_Sa_prime_age[", as.numeric(inds_1_age), ",", as.numeric(inds_2_age), ",", as.numeric(inds_3_age), "]")
  p_Ra_age_v = p_Ra_age_v[!is.na(p_Ra_age_v)]
  p_Sa_prime_age_v = p_Sa_prime_age_v[!is.na(p_Sa_prime_age_v)]

  ## BY ORIGIN, AGGREGATED ACROSS AGES

  # containers
  p_Ra_origin_tmp = p_Sa_prime_origin_tmp = array(NA, dim = c(jags_data$ny, jags_data$no, jags_data$nj))
  inds_1_origin = inds_2_origin = inds_3_origin = p_Ra_origin_tmp
  for (j in 1:jags_data$nj) {
    for (o in 1:jags_data$no) {
      # calculate aggregate proportions for each origin
      p_Ra_origin_tmp[,o,j] = rowSums(p_Ra[,ko_origin[[o]],j])
      p_Sa_prime_origin_tmp[,o,j] = rowSums(p_Sa_prime[,ko_origin[[o]],j])

      # build the element identifiers - for building names later
      inds_1_origin[,o,j] = 1:jags_data$ny
      inds_2_origin[,o,j] = o
      inds_3_origin[,o,j] = j
    }
  }

  # vectorize, add names, and remove NAs
  p_Ra_origin_v = as.numeric(p_Ra_origin_tmp)
  p_Sa_prime_origin_v = as.numeric(p_Sa_prime_origin_tmp)
  names(p_Ra_origin_v) = paste0("p_Ra_origin[", as.numeric(inds_1_origin), ",", as.numeric(inds_2_origin), ",", as.numeric(inds_3_origin), "]")
  names(p_Sa_prime_origin_v) = paste0("p_Sa_prime_origin[", as.numeric(inds_1_origin), ",", as.numeric(inds_2_origin), ",", as.numeric(inds_3_origin), "]")
  p_Ra_origin_v = p_Ra_origin_v[!is.na(p_Ra_origin_v)]
  p_Sa_prime_origin_v = p_Sa_prime_origin_v[!is.na(p_Sa_prime_origin_v)]

  ## COMBINE NEW CALCULATED QUANTITIES ACROSS POSTERIOR SAMPLES

  p_Ra_age = rbind(p_Ra_age, p_Ra_age_v)
  p_Sa_prime_age = rbind(p_Sa_prime_age, p_Sa_prime_age_v)
  p_Ra_origin = rbind(p_Ra_origin, p_Ra_origin_v)
  p_Sa_prime_origin = rbind(p_Sa_prime_origin, p_Sa_prime_origin_v)

}

# combine new calculated quantities into a big matrix
q_new = cbind(p_Ra_age, p_Sa_prime_age, p_Ra_origin, p_Sa_prime_origin)

# combine with the rest of the posterior samples
post = post_bind(post, q_new)

```

```{r calculate-aggregated-comps-data}

# these are the observed proportions by age (aggregated across origins) and origin (aggregated across ages)

# containers
x_Ra_age_obs = x_Sa_prime_age_obs = array(NA, dim = c(jags_data$ny_obs, jags_data$nk, jags_data$nj))
x_Ra_origin_obs = x_Sa_prime_origin_obs = array(NA, dim = c(jags_data$ny_obs, jags_data$no, jags_data$nj))

for (j in 1:jags_data$nj) {
  # calculate aggregate proportions for each age
  for (k in 1:jags_data$nk) {
    x_Ra_age_obs[,k,j] = rowSums(jags_data$x_Ra[,ko_age[[k]],j])
    x_Sa_prime_age_obs[,k,j] = rowSums(jags_data$x_Sa_prime[,ko_age[[k]],j])
  }

  # calculate aggregate proportions for each origin
  for (o in 1:jags_data$no) {
    x_Ra_origin_obs[,o,j] = rowSums(jags_data$x_Ra[,ko_origin[[o]],j])
    x_Sa_prime_origin_obs[,o,j] = rowSums(jags_data$x_Sa_prime[,ko_origin[[o]],j])
  }
}

dimnames(x_Ra_age_obs)[[1]] = dimnames(x_Ra_origin_obs)[[1]] = dimnames(x_Sa_prime_age_obs)[[1]] = dimnames(x_Sa_prime_origin_obs)[[1]] = dimnames(jags_data$x_Ra)[[1]]

x_Ra_new_obs = list(
  x_Ra_age_obs = x_Ra_age_obs,
  x_Ra_origin_obs = x_Ra_origin_obs
)

x_Sa_prime_new_obs = list(
  x_Sa_prime_age_obs = x_Sa_prime_age_obs,
  x_Sa_prime_origin_obs = x_Sa_prime_origin_obs
)

```

Within a single year, the values across all panels sum to one.

#### Weir Sampling {.tabset .tabset-dropdown}

```{r weir-comp-fit-funcs}
# for multinomial sample size
weir_sample_size = function(j) {
  mypar(mfrow = c(1,1))
  mp = barplot(jags_data$nx_Ra[observable,j], xaxt = "n", col = tran_cols["data"], border = main_cols["data"],
               ylim = make_lim(0, jags_data$nx_Ra[observable,j]),
               xlab = "", ylab = "")
  usr = par("usr")
  at_x = seq(1, length(observable), 4)
  axis(side = 1, at = mp[at_x], labels = substr(names(jags_data$nx_Ra[observable,j]), 3, 4)[at_x])
  segments(usr[1], usr[3], usr[2], usr[3], xpd = TRUE)
  axis_labels("Return Year", "Weir Sample Size")
}

# for completely disaggregated fits
weir_comp_fit_plot = function(j) {

  base_param = sub_index("^p_Ra[year,DIM,pop]$", pop = j, year = ts_yrs)
  labels = paste0("Age-", rep(jags_data$kmin:jags_data$kmax, jags_data$no), " (", rep(c("NOR", "HOR"), each = jags_data$nk), ")")

  # function to draw the time series for one panel: one population/age/origin combo
  plot_f = function(ko) {
    plot_tseries(est = post_summ(post, stringr::str_replace(base_param, "DIM", as.character(ko))),
                 obs = get_obs_ests_multinomial(jags_data$x_Ra[observable,,j], ko),
                 label_text = labels[ko], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
  }

  mypar(mfrow = c(2,3))
  junk = sapply(1:jags_data$nko, plot_f)
  axis_labels("Return Year", "Proportion of Return")
}

# for fits to each age, aggregated across origins
weir_comp_fit_plot_age = function(j) {

  base_param = sub_index("^p_Ra_age[year,DIM,pop]$", pop = j, year = ts_yrs)
  labels = paste0("Age-", jags_data$kmin:jags_data$kmax)

  # function to draw the time series for one panel: one population/age combo
  plot_f = function(k) {
    plot_tseries(est = post_summ(post, stringr::str_replace(base_param, "DIM", as.character(k))),
                 obs = get_obs_ests_multinomial(x_Ra_new_obs$x_Ra_age_obs[observable,,j], k),
                 label_text = labels[k], yrs = all_yrs[ts_yrs], pt_cex = pt_cex, ylim = ylim_01)
  }

  mypar(mfrow = c(1,3))
  junk = sapply(1:jags_data$nk, plot_f)
  axis_labels("Return Year", "Proportion of Return")
}

# for fits to each origin, aggregated across ages
weir_comp_fit_plot_origin = function(j) {

  base_param = sub_index("^p_Ra_origin[year,DIM,pop]$", pop = j, year = ts_yrs)
  labels = c("NOR", "HOR")

  # function to draw the time series for one panel: one population/origin combo
  plot_f = function(o) {
    plot_tseries(est = post_summ(post, stringr::str_replace(base_param, "DIM", as.character(o))),
                 obs = get_obs_ests_multinomial(x_Ra_new_obs$x_Ra_origin_obs[observable,,j], o),
                 label_text = labels[o], yrs = all_yrs[ts_yrs], pt_cex = pt_cex, ylim = ylim_01)
  }

  mypar(mfrow = c(1,2))
  junk = sapply(1:jags_data$no, plot_f)
  axis_labels("Return Year", "Proportion of Return")
}
```

##### By Age/Origin {.tabset .tabset-pills}

This is the resolution the model fits to: each year slice is treated as a 6-element multinomial vector.

###### CAT

```{r cat-weir-comp-fit, fig.width = p6_width, fig.height = p6_height}
weir_comp_fit_plot(j_cat)
```

###### LOS

```{r los-weir-comp-fit, fig.width = p6_width, fig.height = p6_height}
weir_comp_fit_plot(j_los)
```

###### MIN

```{r min-weir-comp-fit, fig.width = p6_width, fig.height = p6_height}
weir_comp_fit_plot(j_min)
```

###### UGR

```{r ugr-weir-comp-fit, fig.width = p6_width, fig.height = p6_height}
weir_comp_fit_plot(j_ugr)
```

##### By Age {.tabset .tabset-pills}

These figures show the agreement of the model with the weir composition data by age, aggregated across origins.

###### CAT

```{r cat-weir-comp-fit-age, fig.width = p3_width, fig.height = p3_height}
weir_comp_fit_plot_age(j_cat)
```

###### LOS

```{r los-weir-comp-fit-age, fig.width = p3_width, fig.height = p3_height}
weir_comp_fit_plot_age(j_los)
```

###### MIN

```{r min-weir-comp-fit-age, fig.width = p3_width, fig.height = p3_height}
weir_comp_fit_plot_age(j_min)
```

###### UGR

```{r ugr-weir-comp-fit-age, fig.width = p3_width, fig.height = p3_height}
weir_comp_fit_plot_age(j_ugr)
```

##### By Origin {.tabset .tabset-pills}

These figures show the agreement of the model with the weir composition data by origin, aggregated across ages.

###### CAT

```{r cat-weir-comp-fit-origin, fig.width = p2_width, fig.height = p2_height}
weir_comp_fit_plot_origin(j_cat)
```

###### LOS

```{r los-weir-comp-fit-origin, fig.width = p2_width, fig.height = p2_height}
weir_comp_fit_plot_origin(j_los)
```

###### MIN

```{r min-weir-comp-fit-origin, fig.width = p2_width, fig.height = p2_height}
weir_comp_fit_plot_origin(j_min)
```

###### UGR

```{r ugr-weir-comp-fit-origin, fig.width = p2_width, fig.height = p2_height}
weir_comp_fit_plot_origin(j_ugr)
```

##### Total Sample Size {.tabset .tabset-pills}

These figures show the number of samples gathered for age/origin composition information during carcass surveys. The model will fit more closely to years with more samples.

###### CAT

```{r cat-weir-ss, fig.width = 5, fig.height = 2.5}
weir_sample_size(j_cat)
```

###### LOS

```{r los-weir-ss, fig.width = 5, fig.height = 2.5}
weir_sample_size(j_los)
```

###### MIN

```{r min-weir-ss, fig.width = 5, fig.height = 2.5}
weir_sample_size(j_min)
```

###### UGR

```{r ugr-weir-ss, fig.width = 5, fig.height = 2.5}
weir_sample_size(j_ugr)
```

#### Carcass Sampling {.tabset .tabset-dropdown}

```{r carc-comp-fit-funcs}
# for multinomial sample size
carc_sample_size = function(j) {
  mypar(mfrow = c(1,1))
  mp = barplot(jags_data$nx_Sa_prime[observable,j], xaxt = "n", col = tran_cols["data"], border = main_cols["data"],
               ylim = make_lim(0, max(jags_data$nx_Sa_prime[observable,j])),
               xlab = "", ylab = "")
  usr = par("usr")
  at_x = seq(1, length(observable), 4)
  axis(side = 1, at = mp[at_x], labels = substr(names(jags_data$nx_Sa_prime[observable,j]), 3, 4)[at_x])
  segments(usr[1], usr[3], usr[2], usr[3], xpd = TRUE)
  axis_labels("Return Year", "Carcass Sample Size")
}

# for completely disaggregated fits
carc_comp_fit_plot = function(j) {

  base_param = sub_index("^p_Sa_prime[year,DIM,pop]$", pop = j, year = ts_yrs)
  labels = paste0("Age-", rep(jags_data$kmin:jags_data$kmax, jags_data$no), " (", rep(c("NOR", "HOR"), each = jags_data$nk), ")")

  # function to draw the time series for one panel: one population/age/origin combo
  plot_f = function(ko) {
    plot_tseries(est = post_summ(post, stringr::str_replace(base_param, "DIM", as.character(ko))),
                 obs = get_obs_ests_multinomial(jags_data$x_Sa_prime[observable,,j], ko),
                 label_text = labels[ko], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
  }

  mypar(mfrow = c(2,3))
  junk = sapply(1:jags_data$nko, plot_f)
  axis_labels("Return Year", "Proportion of Carcasses")
}

# for fits to each age, aggregated across origins
carc_comp_fit_plot_age = function(j) {

  base_param = sub_index("^p_Sa_prime_age[year,DIM,pop]$", pop = j, year = ts_yrs)
  labels = paste0("Age-", jags_data$kmin:jags_data$kmax)

  # function to draw the time series for one panel: one population/age combo
  plot_f = function(k) {
    plot_tseries(est = post_summ(post, stringr::str_replace(base_param, "DIM", as.character(k))),
                 obs = get_obs_ests_multinomial(x_Sa_prime_new_obs$x_Sa_prime_age_obs[observable,,j], k),
                 label_text = labels[k], yrs = all_yrs[ts_yrs], pt_cex = pt_cex, ylim = ylim_01)
  }

  mypar(mfrow = c(1,3))
  junk = sapply(1:jags_data$nk, plot_f)
  axis_labels("Return Year", "Proportion of Carcasses")
}

# for fits to each origin, aggregated across ages
carc_comp_fit_plot_origin = function(j) {

  base_param = sub_index("^p_Sa_prime_origin[year,DIM,pop]$", pop = j, year = ts_yrs)
  labels = c("NOR", "HOR")

  # function to draw the time series for one panel: one population/origin combo
  plot_f = function(o) {
    plot_tseries(est = post_summ(post, stringr::str_replace(base_param, "DIM", as.character(o))),
                 obs = get_obs_ests_multinomial(x_Sa_prime_new_obs$x_Sa_prime_origin_obs[observable,,j], o),
                 label_text = labels[o], yrs = all_yrs[ts_yrs], pt_cex = pt_cex, ylim = ylim_01)
  }

  mypar(mfrow = c(1,2))
  junk = sapply(1:jags_data$no, plot_f)
  axis_labels("Return Year", "Proportion of Carcasses")
}
```

##### By Age/Origin {.tabset .tabset-pills}

This is the resolution the model fits to: each year slice is treated as a 12-element multinomial vector.

###### CAT

```{r cat-carc-comp-fit, fig.width = p6_width, fig.height = p6_height}
carc_comp_fit_plot(j_cat)
```

###### LOS

```{r los-carc-comp-fit, fig.width = p6_width, fig.height = p6_height}
carc_comp_fit_plot(j_los)
```

###### MIN

```{r min-carc-comp-fit, fig.width = p6_width, fig.height = p6_height}
carc_comp_fit_plot(j_min)
```

###### UGR

```{r ugr-carc-comp-fit, fig.width = p6_width, fig.height = p6_height}
carc_comp_fit_plot(j_ugr)
```

##### By Age {.tabset .tabset-pills}

These figures show the agreement of the model with the carcass composition data by age, aggregated across origins.

###### CAT

```{r cat-carc-comp-fit-age, fig.width = p3_width, fig.height = p3_height}
carc_comp_fit_plot_age(j_cat)
```

###### LOS

```{r los-carc-comp-fit-age, fig.width = p3_width, fig.height = p3_height}
carc_comp_fit_plot_age(j_los)
```

###### MIN

```{r min-carc-comp-fit-age, fig.width = p3_width, fig.height = p3_height}
carc_comp_fit_plot_age(j_min)
```

###### UGR

```{r ugr-carc-comp-fit-age, fig.width = p3_width, fig.height = p3_height}
carc_comp_fit_plot_age(j_ugr)
```

##### By Origin {.tabset .tabset-pills}

These figures show the agreement of the model with the carcass composition data by origin, aggregated across ages.

###### CAT

```{r cat-carc-comp-fit-origin, fig.width = p2_width, fig.height = p2_height}
carc_comp_fit_plot_origin(j_cat)
```

###### LOS

```{r los-carc-comp-fit-origin, fig.width = p2_width, fig.height = p2_height}
carc_comp_fit_plot_origin(j_los)
```

###### MIN

```{r min-carc-comp-fit-origin, fig.width = p2_width, fig.height = p2_height}
carc_comp_fit_plot_origin(j_min)
```

###### UGR

```{r ugr-carc-comp-fit-origin, fig.width = p2_width, fig.height = p2_height}
carc_comp_fit_plot_origin(j_ugr)
```

##### Total Sample Size {.tabset .tabset-pills}

These figures show the number of samples gathered for age/origin composition information during carcass surveys. The model will fit more closely to years with more samples.

###### CAT

```{r cat-carc-ss, fig.width = 5, fig.height = 2.5}
carc_sample_size(j_cat)
```

###### LOS

```{r los-carc-ss, fig.width = 5, fig.height = 2.5}
carc_sample_size(j_los)
```

###### MIN

```{r min-carc-ss, fig.width = 5, fig.height = 2.5}
carc_sample_size(j_min)
```

###### UGR

```{r ugr-carc-ss, fig.width = 5, fig.height = 2.5}
carc_sample_size(j_ugr)
```

### Spawn Status Data

```{r spawn-status-fits, fig.width = p6_width, fig.height = p6_height}
plot_f = function(j) {
  # top plot: pre-spawn survival time series
  plot_tseries(est = post_summ(post, sub_index("^phi_Sb_Sa[year,pop]", pop = j, year = ts_yrs)),
               obs = get_obs_ests_multinomial(cbind(jags_data$x_carcass_spawned[,j], jags_data$x_carcass_total[,j] - jags_data$x_carcass_spawned[,j])[observable,], 1),
               label_text = pops[j], ylim = c(0,1), yrs = all_yrs[ts_yrs], pt_cex = pt_cex, label_loc = "bottomleft")

  if (j == 1) mtext(side = 2, "Pre-Spawn Survival", line = 1, cex = 0.9)
  # om = par("mar"); nm = om; nm[1] = 2; par(mar = nm); on.exit(par(mar = om))
  mp = barplot(jags_data$x_carcass_total[observable,j], xaxt = "n", ylim = make_lim(0, jags_data$x_carcass_total),
               col = tran_cols["data"], border = main_cols["data"])
  at_x = axisTicks(par("usr")[1:2], log = FALSE)
  axis(side = 1, at = mp[at_x], labels = substr(names(jags_data$x_carcass_total[observable,j]), 3, 4)[at_x])
  box()

  if (j == 1) mtext(side = 2, "Carcasses Sampled", line = 1, cex = 0.9)
}

mypar(mfrow = c(1,1))
layout(matrix(1:(jags_data$nj * 2), 2, jags_data$nj), heights = c(1, 0.65))
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Return Year")
```

### Mean Length Data {.tabset .tabset-pills}

#### Summer Parr

```{r summer-length-fit, fig.width = p4_width, fig.height = p4_height}
# function to create the plot for one population
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^L_Pb[year,pop]$", year = ts_yrs, pop = j)),
               obs = get_obs_ests_log_normal(log(jags_data$L_Pb_obs[observable,j]), jags_data$sig_L_Pb_obs[observable,j]),
               label_text = pops[j], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
}

# loop over populations
mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Parr Mean Length (mm)")
```

#### Spring Smolt

```{r spring-length-fit, fig.width = p4_width, fig.height = p4_height}
# function to create the plot for one population
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^L_Mb[year,pop]$", year = ts_yrs, pop = j)),
               obs = get_obs_ests_log_normal(log(jags_data$L_Mb_obs[observable,j]), jags_data$sig_L_Mb_obs[observable,j]),
               label_text = pops[j], yrs = all_yrs[ts_yrs], pt_cex = pt_cex)
}

# loop over populations
mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Smolt Mean Length (mm)")
```

```{r pp-check-output, child = "output-plots-children/03a-pp-checks.Rmd", eval = model_info$do_pp_check}
```

```{r waic-output, child = "output-plots-children/03b-waic.Rmd", eval = model_info$do_lppd}
```