### 2023-05-04 03-post-process/output-plots-children/13-post-vs-prior.Rmd

---
editor_options: 
  chunk_output_type: console
---

# Prior vs. Posterior {.tabset .tabset-pills}

```{r post-v-prior-fig-sizes}
# set global knitting options
knitr::opts_chunk$set(fig.width = 6, fig.height = 5)
```

```{r post-v-prior-fn}
post_v_prior = function(post_samps, prior_samps, prior_col = "salmon", post_col = "royalblue",
                        main = NULL, y_max = 0.5, xlim = NULL, legend_loc = NA, label_text = NULL, n_bins = 50, alpha = 0.5) {

  if (!is.null(xlim)) {
    post_samps_keep = post_samps[post_samps >= xlim[1] & post_samps <= xlim[2]]
    prior_samps_keep = prior_samps[prior_samps >= xlim[1] & prior_samps <= xlim[2]]
  } else {
    post_samps_keep = post_samps
    prior_samps_keep = prior_samps
  }
  p_post_keep = length(post_samps_keep)/length(post_samps)
  p_prior_keep = length(prior_samps_keep)/length(prior_samps)

  # get the range of all samples
  rng = range(post_samps_keep, prior_samps_keep)
  if (!is.null(xlim)) rng[2] = max(xlim[2], rng[2])

  # set bins
  breaks = seq(rng[1], rng[2], length = n_bins)

  # calculate the data for the histogram
  prior_hist = hist(prior_samps_keep, plot = FALSE, breaks = breaks)
  post_hist = hist(post_samps_keep, plot = FALSE, breaks = breaks)

  # convert bin counts to probabilities
  # account for fraction not shown
  prior_probs = prior_hist$counts/sum(prior_hist$counts) * p_prior_keep
  post_probs = post_hist$counts/sum(post_hist$counts) * p_post_keep

  # graphics parameters
  par(xaxs = "i", yaxs = "i")

  # create a y_max if not supplied
  if (is.null(y_max)) {
    y_max = max(prior_probs, post_probs)
  }

  # create the histogram for prior PDF
  # barplot where height is equal to bin prob
  mp = barplot(
    prior_probs, ylim = c(0,y_max), space = 0, border = prior_col,
    col = scales::alpha(prior_col, alpha), axes = FALSE
  )

  # create the histogram for posterior PDF
  mp = barplot(
    post_probs, ylim = c(0,y_max), space = 0, border = post_col, col = scales::alpha(post_col, alpha), axes = FALSE, add = TRUE
  )

  panel_label(label_text)

  # draw the x-axis
  at_x = axisTicks(par("usr")[1:2], log = FALSE)
  axis(side = 1, at = mp[at_x], labels = round(prior_hist$breaks[at_x], 1))

  # draw the y-axis
  axis(side = 2)

  # draw segments to connect axes at origin
  usr = par("usr"); xdiff = diff(usr[1:2]); ydiff = diff(usr[3:4])
  segments(usr[1], usr[3], usr[2], usr[3], xpd = TRUE, col = par("col.axis"))
  segments(usr[1], usr[3], usr[1], usr[4], xpd = TRUE, col = par("col.axis"))

  # add a legend in the desired location if not NULL
  if (!is.na(legend_loc)) {
    legend(legend_loc, legend = c("Prior", "Posterior"), bty = "n", pch = 22, pt.cex = 1.75, cex = 0.8, col = c(prior_col, post_col), pt.bg = scales::alpha(c(prior_col, post_col), 0.5))
  }

  # if some portion of one of the dists is not shown, include a message
  if (p_post_keep < 1 | p_prior_keep < 1) {
    nice_round = function(x, digits = 2) {
      x_round = round(x, digits = digits)
      if (x_round == 0 & x > 0) x_out  = "<0.01" else x_out = x_round
      x_out
    }

    legend(x = usr[1] - xdiff * 0, y = usr[4] + ydiff * 0.025, xpd = TRUE,
           legend = c(nice_round(1 - p_prior_keep), nice_round(1 - p_post_keep)),
           text.col = c(prior_col, post_col), bty = "n", title.col = "black",
           title = "P Not Shown", cex = 0.8
    )
  }
}
```

```{r post-v-prior-wrappers}
# wrapper to create plot for all populations
post_v_prior_pops = function(post_param, prior_param, legend_loc = "topright", xlim = NULL, is_rho = FALSE, y_max = 0.5, ...) {

  if (is_rho) {
    n = length(vcov_labels)
    titles = vcov_labels
    mfrow = c(3,2)
    looping = vcov_inds
    prior_param = paste0(prior_param, "pop")
    post_param = paste0(post_param, "pop")
  } else {
    n = jags_data$nj
    titles = pops
    mfrow = c(2,2)
    looping = 1:jags_data$nj
  }

  # function to plot one population/pair
  plot_f = function(j) {
    post_v_prior(
      post_samps = post_subset(post, sub_index(post_param, pop = j, ...), matrix = TRUE),
      prior_samps = post_subset(post, sub_index(prior_param, pop = j, ...), matrix = TRUE),
      label_text = titles[j], 
      legend_loc = ifelse(j == looping[1], legend_loc, NA),
      y_max = y_max,
      xlim = xlim
    )
  }

  # apply to each population/pair
  mypar(mfrow = mfrow)
  junk = sapply(1:n, function(j) plot_f(looping[j]))
  axis_labels("Parameter Value", "Probability")
}

# wrapper to create plot for all origins
post_v_prior_origins = function(post_param, prior_param, legend_loc = "topright", xlim = NULL, ...) {
  # function to plot one population
  plot_f = function(o) {
    post_v_prior(
      post_samps = post_subset(post, sub_index(post_param, origin = o, ...), matrix = TRUE),
      prior_samps = post_subset(post, sub_index(prior_param, origin = o, ...), matrix = TRUE),
      label_text = c("NOR", "HOR")[o], 
      legend_loc = ifelse(o == 1, legend_loc, NA),
      xlim = xlim
    )
  }

  # apply to each origin
  mypar(mfrow = c(1,2))
  junk = sapply(1:jags_data$no, plot_f)
  axis_labels("Parameter Value", "Probability")
}
```

## "Location" Parameters {.tabset .tabset-dropdown}

### Max. Egg-to-Parr Survival

**Prior Used**: `dbeta(`r jags_data$alpha_prior[1]`, `r jags_data$alpha_prior[2]`)`

```{r alpha-post-v-prior}
post_v_prior_pops(post_param = "alpha[pop]", prior_param = "alpha_pr", xlim = c(0,1))
```

### Mean Ocean Survival (NOR) {.tabset .tabset-pills}

#### Yr1

**Prior Used**: `dbeta(`r jags_data$mu_phi_O0_O1[1]`, `r jags_data$mu_phi_O0_O1[2]`)`

```{r mu_phi_O0_O1-post-v-prior}
post_v_prior_pops(post_param = "mu_phi_O0_O1[origin,pop]", prior_param = "mu_phi_O0_O1_pr",
                  origin = o_nor, xlim = c(0,1), y_max = 0.75
)
```

#### Yr2

**Prior Used**: `dbeta(`r jags_data$mu_phi_O1_O2[1]`, `r jags_data$mu_phi_O1_O2[2]`)`

```{r mu_phi_O1_O2-post-v-prior}
post_v_prior_pops(post_param = "mu_phi_O1_O2[origin,pop]", prior_param = "mu_phi_O1_O2_pr",
                  origin = o_nor, xlim = c(0,1), legend_loc = "topleft"
)
```

#### Yr3

**Prior Used**: `dbeta(`r jags_data$mu_phi_O2_O3[1]`, `r jags_data$mu_phi_O2_O3[2]`)`

```{r mu_phi_O2_O3-post-v-prior}
post_v_prior_pops(post_param = "mu_phi_O2_O3[origin,pop]", prior_param = "mu_phi_O2_O3_pr",
                  origin = o_nor, xlim = c(0,1), legend_loc = "topleft"
)
```

### Mean Maturation {.tabset .tabset-pills}

#### Age-3 {.tabset .tabset-pills}

**Prior Used**: `dbeta(`r jags_data$mu_psi_O1[1]`, `r jags_data$mu_psi_O1[2]`)`

##### NOR

```{r mu_psi_O1_nor-post-v-prior}
post_v_prior_pops(post_param = "mu_psi_O1[origin,pop]", prior_param = "mu_psi_O1_pr",
                  origin = o_nor, xlim = c(0,1), y_max = 0.75
)
```

##### HOR

```{r mu_psi_O1_hor-post-v-prior}
post_v_prior_pops(post_param = "mu_psi_O1[origin,pop]", prior_param = "mu_psi_O1_pr",
                  origin = o_hor, xlim = c(0,1), y_max = 0.75
)
```

#### Age-4 {.tabset .tabset-pills}

**Prior Used**: `dbeta(`r jags_data$mu_psi_O2[1]`, `r jags_data$mu_psi_O2[2]`)`

##### NOR

```{r mu_psi_O2_nor-post-v-prior}
post_v_prior_pops(post_param = "mu_psi_O2[origin,pop]", prior_param = "mu_psi_O2_pr",
                  origin = o_nor, xlim = c(0,1), legend_loc = "topleft"
)
```

##### HOR

```{r mu_psi_O2_hor-post-v-prior}
post_v_prior_pops(post_param = "mu_psi_O2[origin,pop]", prior_param = "mu_psi_O2_pr",
                  origin = o_hor, xlim = c(0,1), legend_loc = "topleft"
)
```

## "Scale" Parameters {.tabset .tabset-dropdown}

### Egg-to-Parr Survival

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_E_Pb_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lphi_E_Pb_prior[2]`)`

```{r sig_Lphi_E_Pb-post-v-prior}
post_v_prior_pops(post_param = "sig_Lphi_E_Pb[pop]", prior_param = "sig_Lphi_E_Pb_pr[pop]",
                  xlim = c(0, 2)
)
```

### Parr Mean Length

**Process Noise Distribution**: Multivariate log-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_lL_Pb_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_lL_Pb_prior[2]`)`

```{r sig_lL_Pb-post-v-prior}
post_v_prior_pops(post_param = "sig_lL_Pb[pop]", prior_param = "sig_lL_Pb_pr[pop]",
                  xlim = c(0, 1), y_max = 0.75
)
```

### Proportion Fall Migrants

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lpi_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lpi_prior[2]`)`

```{r sig_Lpi-post-v-prior}
post_v_prior_pops(post_param = "sig_Lpi[pop]", prior_param = "sig_Lpi_pr[pop]",
                  xlim = c(0, 2)
)
```

### Overwinter Survival

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_Pa_Mb_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lphi_Pa_Mb_prior[2]`)`

```{r sig_Lphi_Pa_Mb-post-v-prior}
post_v_prior_pops(post_param = "sig_Lphi_Pa_Mb[pop]", prior_param = "sig_Lphi_Pa_Mb_pr[pop]",
                  xlim = c(0, 2)
)
```

### "Growth" Factor

**Process Noise Distribution**: Multivariate log-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_lDelta_L_Pb_Mb_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_lDelta_L_Pb_Mb_prior[2]`)`

```{r sig_lDelta_L_Pb_Mb-post-v-prior}
post_v_prior_pops(post_param = "sig_lDelta_L_Pb_Mb[pop]", prior_param = "sig_lDelta_L_Pb_Mb_pr[pop]",
                  xlim = c(0,1), y_max = 0.75
)
```

### Trib to LGR Survival

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_Mb_Ma_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lphi_Mb_Ma_prior[2]`)`

```{r sig_Lphi_Mb_Ma-post-v-prior}
post_v_prior_pops(post_param = "sig_Lphi_Mb_Ma[pop]", prior_param = "sig_Lphi_Mb_Ma_pr[pop]",
                  xlim = c(0, 2)
)
```

### LGR to Ocean Survival

**Process Noise Distribution**: Bivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_Ma_O0_prior[1]`,`r jags_data$no`),`r jags_data$Tau_Lphi_Ma_O0_prior[2]`)`

```{r sig_Lphi_Ma_O0-post-v-prior, fig.width = 6, fig.height = 3}
post_v_prior_origins(post_param = "sig_Lphi_Ma_O0[origin]", prior_param = "sig_Lphi_Ma_O0_pr[origin]",
                  xlim = c(0, 2)
)
```

### Yr 1 Ocean Survival

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_O0_O1_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lphi_O0_O1_prior[2]`)`

```{r sig_Lphi_O0_O1-post-v-prior}
post_v_prior_pops(post_param = "sig_Lphi_O0_O1[pop]", prior_param = "sig_Lphi_O0_O1_pr[pop]",
                  xlim = c(0, 2)
)
```

### Maturity {.tabset .tabset-pills}

#### Pr(Age-3)

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lpsi_O1_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lpsi_O1_prior[2]`)`

```{r sig_Lpsi-O1-post-v-prior}
post_v_prior_pops(post_param = "sig_Lpsi_O1[pop]", prior_param = "sig_Lpsi_O1_pr[pop]",
                  xlim = c(0, 2)
)
```

#### Pr(Age-4)

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lpsi_O2_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lpsi_O2_prior[2]`)`

```{r sig_Lpsi_O2-post-v-prior}
post_v_prior_pops(post_param = "sig_Lpsi_O2[pop]", prior_param = "sig_Lpsi_O2_pr[pop]",
                  xlim = c(0, 2)
)
```

### BON to LGR Survival

**Process Noise Distribution**: Bivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_Rb_Ra_prior[1]`,`r jags_data$no`),`r jags_data$Tau_Lphi_Rb_Ra_prior[2]`)`

```{r sig_Lphi_Rb_Ra-post-v-prior, fig.width = 6, fig.height = 3}
post_v_prior_origins(post_param = "sig_Lphi_Rb_Ra[origin]", prior_param = "sig_Lphi_Rb_Ra_pr[origin]",
                  xlim = c(0, 2)
)
```

### Pre-spawn Survival

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_Sb_Sa_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lphi_Sb_Sa_prior[2]`)`

```{r sig_Lphi_Sb_Sa-post-v-prior}
post_v_prior_pops(post_param = "sig_Lphi_Sb_Sa[pop]", prior_param = "sig_Lphi_Sb_Sa_pr[pop]",
                  xlim = c(0, 2)
)
```

## Correlation Parameters {.tabset .tabset-dropdown}

```{r}
# set these figure options for all plots in this section
y_max = 0.15
xlim = c(-1,1)
knitr::opts_chunk$set(fig.width = 5, fig.height = 6)
```

### Egg-to-Parr Survival

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_E_Pb_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lphi_E_Pb_prior[2]`)`

```{r rho_Lphi_E_Pb-post-v-prior}
post_v_prior_pops(post_param = "rho_Lphi_E_Pb", prior_param = "rho_Lphi_E_Pb_pr",
                  xlim = xlim, y_max = y_max, is_rho = TRUE
)
```

### Parr Mean Length

**Process Noise Distribution**: Multivariate log-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_lL_Pb_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_lL_Pb_prior[2]`)`

```{r rho_lL_Pb-post-v-prior}
post_v_prior_pops(post_param = "rho_lL_Pb", prior_param = "rho_lL_Pb_pr",
                  xlim = xlim, y_max = y_max, is_rho = TRUE
)
```

### Proportion Fall Migrants

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lpi_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lpi_prior[2]`)`

```{r rho_Lpi-post-v-prior}
post_v_prior_pops(post_param = "rho_Lpi", prior_param = "rho_Lpi_pr",
                  xlim = xlim, y_max = y_max, is_rho = TRUE
)
```

### Overwinter Survival

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_Pa_Mb_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lphi_Pa_Mb_prior[2]`)`

```{r rho_Lphi_Pa_Mb-post-v-prior}
post_v_prior_pops(post_param = "rho_Lphi_Pa_Mb", prior_param = "rho_Lphi_Pa_Mb_pr",
                  xlim = xlim, y_max = y_max, is_rho = TRUE
)
```

### "Growth" Factor

**Process Noise Distribution**: Multivariate log-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_lDelta_L_Pb_Mb_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_lDelta_L_Pb_Mb_prior[2]`)`

```{r rho_lDelta_L_Pb_Mb-post-v-prior}
post_v_prior_pops(post_param = "rho_lDelta_L_Pb_Mb", prior_param = "rho_lDelta_L_Pb_Mb_pr",
                  xlim = xlim, y_max = y_max, is_rho = TRUE
)
```

### Trib to LGR Survival

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_Mb_Ma_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lphi_Mb_Ma_prior[2]`)`

```{r rho_Lphi_Mb_Ma-post-v-prior}
post_v_prior_pops(post_param = "rho_Lphi_Mb_Ma", prior_param = "rho_Lphi_Mb_Ma_pr",
                  xlim = xlim, y_max = y_max, is_rho = TRUE
)
```

### LGR to Ocean Survival

**Process Noise Distribution**: Bivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_Ma_O0_prior[1]`,`r jags_data$no`),`r jags_data$Tau_Lphi_Ma_O0_prior[2]`)`

```{r rho_Lphi_Ma_O0-post-v-prior, fig.width = 3, fig.height = 3}
mypar(mfrow = c(1,1))
post_v_prior(post_samps = post_subset(post, "rho_Lphi_Ma_O0[2,1]", matrix = TRUE),
             prior_samps = post_subset(post, "rho_Lphi_Ma_O0_pr[2,1]", matrix = TRUE),
             xlim = xlim, y_max = y_max)
axis_labels("Parameter Value", "Probability")
```

### Yr 1 Ocean Survival

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_O0_O1_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lphi_O0_O1_prior[2]`)`

```{r rho_Lphi_O0_O1-post-v-prior}
post_v_prior_pops(post_param = "rho_Lphi_O0_O1", prior_param = "rho_Lphi_O0_O1_pr",
                  xlim = xlim, y_max = y_max, is_rho = TRUE
)
```

### Maturity {.tabset .tabset-pills}

#### Pr(Age-3)

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lpsi_O1_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lpsi_O1_prior[2]`)`

```{r rho_Lpsi-O1-post-v-prior}
post_v_prior_pops(post_param = "rho_Lpsi_O1", prior_param = "rho_Lpsi_O1_pr",
                  xlim = xlim, y_max = y_max, is_rho = TRUE
)
```

#### Pr(Age-4)

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lpsi_O2_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lpsi_O2_prior[2]`)`

```{r rho_Lpsi_O2-post-v-prior}
post_v_prior_pops(post_param = "rho_Lpsi_O2", prior_param = "rho_Lpsi_O2_pr",
                  xlim = xlim, y_max = y_max, is_rho = TRUE
)
```

### BON to LGR Survival

**Process Noise Distribution**: Bivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_Rb_Ra_prior[1]`,`r jags_data$no`),`r jags_data$Tau_Lphi_Rb_Ra_prior[2]`)`

```{r rho_Lphi_Rb_Ra-post-v-prior, fig.width = 3, fig.height = 3}
par(oma = c(1,1.75,0,0))
post_v_prior(post_samps = post_subset(post, "rho_Lphi_Rb_Ra[2,1]", matrix = TRUE),
             prior_samps = post_subset(post, "rho_Lphi_Rb_Ra_pr[2,1]", matrix = TRUE),
             xlim = xlim, y_max = y_max)
mtext(side = 1, outer = TRUE, line = 0, "Parameter Value")
mtext(side = 2, outer = TRUE, line = 0.5, "Probability")
```

### Pre-spawn Survival

**Process Noise Distribution**: Multivariate logit-normal

**Prior Used**: `dscaled.wishart(rep(`r jags_data$Tau_Lphi_Sb_Sa_prior[1]`,`r jags_data$nj`),`r jags_data$Tau_Lphi_Sb_Sa_prior[2]`)`

```{r rho_Lphi_Sb_Sa-post-v-prior}
post_v_prior_pops(post_param = "rho_Lphi_Sb_Sa", prior_param = "rho_Lphi_Sb_Sa_pr",
                  xlim = xlim, y_max = y_max, is_rho = TRUE
)
```