### 2023-05-04 03-post-process/output-plots-children/10-rho-params.Rmd
Consistency updates: 10-rho-params.Rmd

---
editor_options: 
  chunk_output_type: console
---

# Correlation Parameters {.tabset .tabset-pills}

## Summary Across Processes

```{r summarize-rho-means}
# get posterior summaries of the mean correlation across all pop pairs
# for each process separately
rho_params1 = match_params(post, "rho", type = "base_only")
rho_params2 = match_params(post, "rho_.+_pr", type = "base_only")
rho_params = rho_params1[!(rho_params1 %in% rho_params2)]

f = function(p) {
  x = rowMeans(post_subset(post, paste0("^", p, "["), matrix = TRUE))
  c(mean = mean(x), sd = sd(x), quantile(x, c(0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975)))
}
rho_mean_out = t(sapply(rho_params, f))
rho_mean_out = rho_mean_out[order(rho_mean_out[,"mean"]),]

# assign names to the different rho parameters
rho_labels = c(
  "rho_Lphi_E_Pb" = "Egg-to-Parr Survival",
  "rho_Lphi_O0_O1" = "Yr1 Ocean Survival",
  "rho_Lpsi_O1" = "Pr(Mature Age-3)",
  "rho_Lpsi_O2" = "Pr(Mature Age-4)",
  "rho_lDelta_L_Pb_Mb" = "'Growth' Factor",
  "rho_Lphi_Mb_Ma" = "Trib. to LGR Survival",
  "rho_lL_Pb" = "Mean Parr Length",
  "rho_Lphi_Pa_Mb" = "Overwinter Survival",
  "rho_Lphi_Sb_Sa" = "Pre-spawn Survival",
  "rho_Lphi_Rb_Ra" = "BON -> LGR Survival",
  "rho_Lpi" = "Pr(Fall Migrant)",
  "rho_Lphi_Ma_O0" = "LGR -> Ocean Survival"
)
```

Posterior summaries of the correlation matrices for all process model components that used a multivariate (logit- or log-) normal distribution model inter-annual variability.
All processes assumed covariance among population pairs except "LGR to Ocean" and "BON to LGR", which had equal values for all populations but covaried across origin types.

Notation:

* Main: Mean correlation of all population pairs ("+" = posterior mean, thick lines = 50% CRI, thin lines = 95% CRI)
* Dots: Posterior mean correlation of individual population pairs
<!-- * Grey: Prior distribution (95%, 80%, and 50% CRIs in increasing darkness) -->

```{r rho-mean-plot, fig.width = 5, fig.height = 5, out.width = "50%"}
# plot to show summary of correlation parameters across processes
# mean across pairs shown as blue
# individual population pairs shown as red

# set up empty plotting region
mypar(mar = c(2,8,1,1), yaxs = "i", mfrow = c(1,1), oma = c(0,0,0,0), mgp = c(2,0.25,0))
mp = barplot(rho_mean_out[,"mean"], 
             horiz = TRUE, xlim = c(-1,1),
             names.arg = rho_labels[rownames(rho_mean_out)],
             las = 1, col = "white", border = "white")
usr = par("usr")

box_upr = mp - rbind(diff(mp)/2, NA)
box_lwr = mp + rbind(diff(mp)/2, NA)
box_upr[nrow(box_upr),] = box_upr[nrow(box_upr)-1,] + diff(box_upr[1:2])
box_lwr[nrow(box_lwr),] = box_lwr[nrow(box_upr)-1,] + diff(box_lwr[1:2])

# draw on the prior distribution for rho
draw_rho_prior = function(param, box_lwr, box_upr) {
  rho_pr = post_summ(post, paste0(param, "_pr[2,1]"), probs = c(0.025, 0.1, 0.25, 0.75, 0.9, 0.975))
  rect(rho_pr["2.5%",], box_lwr, rho_pr["97.5%",], box_upr, col = scales::alpha("grey25", 0.15), border = NA)
  rect(rho_pr["10%",], box_lwr, rho_pr["90%",], box_upr, col = scales::alpha("grey25", 0.15), border = NA)
  rect(rho_pr["25%",], box_lwr, rho_pr["75%",], box_upr, col = scales::alpha("grey25", 0.15), border = NA)
  abline(v = 0, lty = 2)
}

pop_rhos = function(param, at_mp) {

  # if the process is not one of the hydropower migration survival ones, do this
  # those processes are correlated across origins, because all populations share the same values each year
  if (!(param %in% c("rho_Lphi_Ma_O0", "rho_Lphi_Rb_Ra"))) {
    x = post_summ(post, paste0(param, vcov_inds), probs = c(0.1, 0.25, 0.75, 0.9))
    y = at_mp + runif(6, -0.4, 0.4)
    points(y ~ x["mean",], pch = 21, col = main_cols["model"], bg = tran_cols["model"])
  } 
}

# junk = sapply(1:nrow(mp), function(i) draw_rho_prior(rownames(rho_mean_out)[i], box_lwr[i,], box_upr[i,]))
segments(rho_mean_out[,"25%"], mp, rho_mean_out[,"75%"], mp, lwd = 6, col = main_cols["model"])
segments(rho_mean_out[,"2.5%"], mp, rho_mean_out[,"97.5%"], mp, col = main_cols["model"])
points(mp ~ rho_mean_out[,"mean"], pch = 3, cex = 1, col = main_cols["model"], lwd = 3)
junk = sapply(1:nrow(mp), function(i) pop_rhos(rownames(rho_mean_out)[i], mp[i,]))

abline(h = mp + rbind(diff(mp)/2, NA), xpd = FALSE, col = par("col.axis"), lty = 2)
abline(v = 0, lty = 2, col = par("col.axis"))
par(tcl = -7)
axis(side = 2, at = mp + rbind(diff(mp)/2, NA), col = par("col.axis"), line = NA, labels = FALSE, lty = 2)
box(col = par("col.axis"))
```

## Process-Specific Summaries {.tabset .tabset-dropdown}

```{r rho_mat_plot-fns}

# one correlagram: for one summary statistic
rho_mat_plot = function(param, q = "mean", main) {
  rho_ests = post_summ(post, param, probs = c(0.025, 0.975))
  rho_mat = array_format(rho_ests[q,])
  rownames(rho_mat) = colnames(rho_mat) = pops
  col = colorRampPalette(c("salmon", "white", "royalblue"))(20)
  corrplot::corrplot(rho_mat, method = "color", type = "lower", tl.srt = 45, col = col, addCoef.col = "black", diag = FALSE, cl.pos = "n")
  mtext(line = -2.5, text = main, side = 3, font = 2, cex = 1, col = "grey50")
}

# a set of three: 80% quantiles & mean
rho_mat_plots = function(param, is_prior = FALSE) {
  par(mfrow = c(1,3), mar = c(1,1,1,1), cex = 1)
  dist = ifelse(is_prior, "Prior", "Posterior")
  rho_mat_plot(param = param, q = "2.5%", main = paste0(dist, " 2.5%"))
  rho_mat_plot(param = param, q = "mean", main = paste0(dist, " Mean"))
  rho_mat_plot(param = param, q = "97.5%", main = paste0(dist, " 97.5%"))
}

rho_mat_fig_h = 4
rho_mat_fig_w = 8
```

### Prior

```{r rho_prior-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_Lphi_E_Pb_pr[", is_prior = TRUE)
```

### Egg-to-Parr Survival

```{r rho_Lphi_E_Pb-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_Lphi_E_Pb[")
```

### Mean Parr Length

```{r rho_lL_Pb-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_lL_Pb[")
```

### Proportion Fall Migrants

```{r rho_Lpi-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_Lpi[")
```

### Overwinter Survival

```{r rho_Lphi_Pa_Mb-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_Lphi_Pa_Mb[")
```

### 'Growth' Factor

```{r rho_lDelta_L_Pb_Mb-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_lDelta_L_Pb_Mb[")
```

### Trib. to LGR Survival

```{r rho_Lphi_Mb_Ma-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_Lphi_Mb_Ma[")
```

### LGR to Ocean Survival

```{r rho_Lphi_Mb_O0-plot}
est = post_summ(post, "rho_Lphi_Ma_O0[2,1]")
kable(t(est), digits = 3, row.names = FALSE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("condensed"))
```

### Yr1 Ocean Survival

```{r rho_Lphi_O0_O1-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_Lphi_O0_O1[")
```

### Maturity {.tabset .tabset-pills}

#### Pr(Age-3)

```{r rho_Lpsi_O1-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_Lpsi_O1[")
```

#### Pr(Age-4)

```{r rho_Lpsi_O2-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_Lpsi_O2[")
```

### BON to LGR Survival

```{r rho_Lphi_Rb_Ra-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
est = post_summ(post, "rho_Lphi_Rb_Ra[2,1]")
kable(t(est), digits = 3, row.names = FALSE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("condensed"))
```

### Pre-Spawn Survival

```{r rho_Lphi_Sb_Sa-plot, fig.height = rho_mat_fig_h, fig.width = rho_mat_fig_w}
rho_mat_plots("rho_Lphi_Sb_Sa[")
```