### 2023-06-02 03-post-process/output-plots-children/08-misc-processes.Rmd
Rename file: misc-processes

---
editor_options: 
  chunk_output_type: console
---

# Miscellaneous Processes {.tabset .tabset-pills}

## Migratory Strategy Apportionment

```{r pi-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^pi[year,LH_type,pop]$", year = ts_yrs, LH_type = i_fall, pop = j)),
               label_text = pops[j], yrs = all_yrs[ts_yrs])
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Proportion of Parr with Fall Migration Strategy")
```

## Ocean Survival {.tabset .tabset-pills}

### Yr1 {.tabset .tabset-pills}

#### NOR {.tabset .tabset-pills}

##### Realized Alone

The `r colorize("realized time series", main_cols["model"])`.

```{r phi-O0-O1-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^phi_O0_O1[year,origin,pop]", year = ts_yrs, origin = o_nor, pop = j)
  plot_tseries(est = post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = c(0, 0.4))
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Yr1 Ocean Survival (NOR)")
```

##### Realized & [Expected + AR(1)]

The `r colorize("realized time series", main_cols["model"])` and the `r colorize("expected time series including the carry-over residual", main_cols["model2"])`.

```{r phi-O0-O1-plots2, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^phi_O0_O1[year,origin,pop]", year = ts_yrs, origin = o_nor, pop = j)
  plot_tseries(post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = c(0, 0.4))

  keep = sub_index("^phi_O0_O1_dot2[year,origin,pop]", year = ts_yrs, origin = o_nor, pop = j)
  add_tseries(post_summ(post, keep), yrs = all_yrs[ts_yrs])
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Yr1 Ocean Survival (NOR)")
```

#### HOR {.tabset .tabset-pills}

##### Realized Alone

The `r colorize("realized time series", main_cols["model"])`.

```{r phi-O0-O1-hor-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^phi_O0_O1[year,origin,pop]", year = ts_yrs, origin = o_hor, pop = j)
  plot_tseries(est = post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = c(0, 0.4))
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Yr1 Ocean Survival (HOR)")
```

### Yr2 {.tabset .tabset-pills}

#### NOR

```{r phi-O1-O2-nor-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^phi_O1_O2[year,origin,pop]", year = ts_yrs, origin = o_nor, pop = j)
  plot_tseries(est = post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = ylim_01)
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Yr2 Ocean Survival (NOR)")
```

#### HOR

```{r phi-O1-O2-hor-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^phi_O1_O2[year,origin,pop]", year = ts_yrs, origin = o_hor, pop = j)
  plot_tseries(est = post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = ylim_01)
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Yr2 Ocean Survival (HOR)")
```

### Yr3 {.tabset .tabset-pills}

#### NOR

```{r phi-O2-O3-nor-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^phi_O2_O3[year,origin,pop]", year = ts_yrs, origin = o_nor, pop = j)
  plot_tseries(est = post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = ylim_01)
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Yr3 Ocean Survival (NOR)")
```

#### HOR

```{r phi-O2-O3-hor-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^phi_O2_O3[year,origin,pop]", year = ts_yrs, origin = o_hor, pop = j)
  plot_tseries(est = post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = ylim_01)
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Yr3 Ocean Survival (HOR)")
```

## Maturity {.tabset .tabset-pills}

### Pr(Age-3) {.tabset .tabset-pills}

#### NOR

```{r psi-O1-nor-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^psi_O1[year,origin,pop]", year = ts_yrs, origin = o_nor, pop = j)
  plot_tseries(est = post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = c(0, 0.4))
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Pr(Mature Age-3|NOR)")
```

#### HOR

```{r psi-O1-hor-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^psi_O1[year,origin,pop]", year = ts_yrs, origin = o_hor, pop = j)
  plot_tseries(est = post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = c(0, 0.4))
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Pr(Mature Age-3|HOR)")
```

### Pr(Age-4) {.tabset .tabset-pills}

#### NOR

```{r psi-O2-nor-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^psi_O2[year,origin,pop]", year = ts_yrs, origin = o_nor, pop = j)
  plot_tseries(est = post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = ylim_01)
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Pr(Mature Age-4|NOR)")
```

#### HOR

```{r psi-O2-hor-plots, fig.width = p4_width, fig.height = p4_height}
plot_f = function(j) {
  keep = sub_index("^psi_O2[year,origin,pop]", year = ts_yrs, origin = o_hor, pop = j)
  plot_tseries(est = post_summ(post, keep), yrs = all_yrs[ts_yrs], label_text = pops[j], ylim = ylim_01)
}

mypar()
junk = sapply(1:jags_data$nj, plot_f)
axis_labels("Brood Year", "Pr(Mature Age-4|HOR)")
```

## Straying Dynamics {.tabset .tabset-pills}

### Number of Strays

```{r n-stray, fig.width = p4_width, fig.height = p4_height}
plot_fn = function(j) {
  plot_tseries(est = post_summ(post, sub_index("^G[year,origin,pop]", pop = j, year = ts_yrs, origin = o_hor)), yrs = all_yrs[ts_yrs], label_text = pops[j])
}

mypar()
junk = sapply(1:jags_data$nj, plot_fn)
axis_labels("Return Year", "Number of Strays Entering Population")
```

### Stray Composition

```{r stray-comp, fig.width = p4_width, fig.height = p4_height}
plot_fn = function(j) {
  x = post_summ(post, sub_index("p_G[age,origin,pop]", age = ".", origin = o_hor, pop = j))

  mp = barplot(x["mean",], ylim = ylim_01, xaxt = "n", col = tran_cols["model"], border = main_cols["model"])
  segments(mp, x["2.5%",], mp, x["97.5%",], col = alpha("red", 0.5))
  axis(side = 1, at = mp, labels = paste0("Age", jags_data$kmin:jags_data$kmax), las = 1)
  segments(par("usr")[1], par("usr")[3], par("usr")[2], par("usr")[3], xpd = TRUE, col = par("col.axis"))
  segments(par("usr")[1], par("usr")[3], par("usr")[1], par("usr")[4], xpd = TRUE, col = par("col.axis"))
  panel_label(pops[j])
}

mypar()
junk = sapply(1:jags_data$nj, plot_fn)
axis_labels("Age of Return", "Proportion of Strays by Age")
```

## Correction for Carcass Age Bias

 {.tabset .tabset-pills}

Exploratory analyses indicated that older fish (age-4 and age-5) are over represented in carcass surveys relative to age-3 fish (based on comparisons between weir and carcass composition data.
This could cause model fit inconsistencies, and because Minam River does not have a weir, using only the carcass data to inform its age composition could give biased estimates.
We thus devised a correction factor that allowed the observation model to fit to biased carcass counts based on a corrected latent value.

```{r add-p-Sa}
# extract posterior samples of spawners[year,age,origin,pop]
Sa_post = post_subset(post, "^Sa[", matrix = TRUE)

# extract posterior samples of corrected spawner composition [year,age-origin,pop]
p_Sa_prime_post = post_subset(post, "^p_Sa_prime[", matrix = TRUE)

# function to get p_Sa (uncorrected spawner composition) for one posterior sample
g = function(i) {
  # cat("\r", i)
  Sa_i = array_format(Sa_post[i,])
  Sa_i = abind::abind(Sa_i[,,1,], Sa_i[,,2,], along = 2)
  p_Sa_i = apply(Sa_i, 3, function(x) apply(x, 1, function(y) y/sum(y)), simplify = FALSE)
  p_Sa_i = do.call(abind::abind, c(lapply(p_Sa_i, t), along = 3))
  na_ind = which(!is.na(p_Sa_i))
  p_Sa_i = p_Sa_i[na_ind]
  names(p_Sa_i) = stringr::str_remove(colnames(p_Sa_prime_post), "_prime")
  p_Sa_i
}

# apply the function
p_Sa_post = sapply(1:nrow(Sa_post), g)
post = post_bind(post, t(p_Sa_post))
```

```{r p-Sa-prime-plots-fn}
# function to create the plot for one population
plot_f = function(j, ko) {
  ko_names = paste0("Age-", rep(3:5, 2))
  ko_names = paste(ko_names, rep(c("(NOR)", "(HOR)"), each = 3))
  # realized time series
  plot_tseries(est = post_summ(post, sub_index("^p_Sa[year,age,pop]$", year = ts_yrs, pop = j, age = ko)),
               yrs = all_yrs[ts_yrs], label_text = ko_names[ko])

  # time series that includes BH prediction only
  add_tseries(post_summ(post, sub_index("^p_Sa_prime[year,age,pop]", year = ts_yrs, pop = j, age = ko)),
              yrs = all_yrs[ts_yrs])
}
```

Spawner percent composition displayed as `r colorize("latent (actual) time series", main_cols["model"])`  and the `r colorize("observed (i.e., biased in sampling, but accounted for in model) time series", main_cols["model2"])`.
Put another way, `r colorize("perceived", main_cols["model2"])` and `r colorize("reality", main_cols["model"])`.

## CAT

```{r p-Sa-prime-plots-CAT, fig.width = p6_width, fig.height = p6_height}
# loop over age/origin combos for CAT
mypar(mfrow = c(2,3))
junk = sapply(1:jags_data$nko, function(ko) plot_f(j = j_cat, ko = ko))
axis_labels("Brood Year", "% Composition")
```

## LOS

```{r p-Sa-prime-plots-LOS, fig.width = p6_width, fig.height = p6_height}
# loop over age/origin combos for CAT
mypar(mfrow = c(2,3))
junk = sapply(1:jags_data$nko, function(ko) plot_f(j = j_los, ko = ko))
axis_labels("Brood Year", "% Composition")
```

## MIN

```{r p-Sa-prime-plots-MIN, fig.width = p6_width, fig.height = p6_height}
# loop over age/origin combos for CAT
mypar(mfrow = c(2,3))
junk = sapply(1:jags_data$nko, function(ko) plot_f(j = j_min, ko = ko))
axis_labels("Brood Year", "% Composition")
```

## UGR

```{r p-Sa-prime-plots-UGR, fig.width = p6_width, fig.height = p6_height}
# loop over age/origin combos for CAT
mypar(mfrow = c(2,3))
junk = sapply(1:jags_data$nko, function(ko) plot_f(j = j_ugr, ko = ko))
axis_labels("Brood Year", "% Composition")
```