### 2023-06-28 03-post-process/ms-content/ms-content.Rmd
Initialize ms-content.Rmd

---
output: 
  pdf_document:
    keep_tex: true
params:
  scenario:
    label: "Scenario"
    value: "base_vlong"
    input: text
  thin_percent:
    label: "Proportion of Posterior Samples to Retain?"
    value: 1
    input: numeric
  include_sim_years:
    label: "Include Simulated Years on Plots?"
    value: FALSE
header-includes:
  - \usepackage{sansmathfonts}
  - \usepackage{helvet}
  - \renewcommand{\rmdefault}{\sfdefault}
  - \usepackage{float}
  - \usepackage[labelfont={bf,sc},labelsep=period]{caption}
  - \captionsetup{width=\textwidth}
editor_options: 
  chunk_output_type: console
---

\rule{\textwidth}{1pt}

\vspace{-6pt}

::: {.Huge data-latex=""}
**Potential Manuscript Content**
:::

\vspace{-6pt}

::: {.Large data-latex=""}
**GR-sslcm**
:::

\vspace{-6pt}

\rule{\textwidth}{1pt}

\listoftables

\listoffigures

\pagenumbering{gobble}
\newpage
\pagenumbering{arabic}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center", fig.pos = "H")
knitr::opts_knit$set(root.dir = "C:/Users/bstaton/Desktop/Staton/1_critfc/analyses/GR-sslcm/03-post-process")

# set up a chunk hook to calculate how long chunks took to run
# modified from https://bookdown.org/yihui/rmarkdown-cookbook/time-chunk.html#time-chunk

start_times = list()  # store the time for each chunk
end_times = list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  function(before, options) {
    if (before) {
      start_times[[options$label]] <<- lubridate::now()
    } else {
      end_times[[options$label]] <<- lubridate::now()
    }
  }
}))
```

```{r load-post, time_it = TRUE}
source("C:/Users/bstaton/Desktop/source_rmd_chunk.R")
rmd_file = "C:/Users/bstaton/Desktop/Staton/1_critfc/analyses/GR-sslcm/03-post-process/output-plots.Rmd"
source_rmd_chunk(rmd_file, "setup")
source_rmd_chunk(rmd_file, "dimension-ids")
knitr::opts_chunk$set(out.width = "100%", dev = "cairo_pdf")
```

```{r load-packages}
source("../00-packages.R")
```

```{r plot-settings}
solid_col = "grey30"
tranp_col = alpha(solid_col, 0.25)
pt_cex = 1.3
```