### 2023-06-28 03-post-process/ms-content/ms-content.Rmd
Create capacity relationship figure

---
output: 
  pdf_document:
    keep_tex: true
params:
  scenario:
    label: "Scenario"
    value: "base_vlong"
    input: text
  thin_percent:
    label: "Proportion of Posterior Samples to Retain?"
    value: 1
    input: numeric
  include_sim_years:
    label: "Include Simulated Years on Plots?"
    value: FALSE
header-includes:
  - \usepackage{sansmathfonts}
  - \usepackage{helvet}
  - \renewcommand{\rmdefault}{\sfdefault}
  - \usepackage{float}
  - \usepackage[labelfont={bf,sc},labelsep=period]{caption}
  - \captionsetup{width=\textwidth}
editor_options: 
  chunk_output_type: console
---

\rule{\textwidth}{1pt}

\vspace{-6pt}

::: {.Huge data-latex=""}
**Potential Manuscript Content**
:::

\vspace{-6pt}

::: {.Large data-latex=""}
**GR-sslcm**
:::

\vspace{-6pt}

\rule{\textwidth}{1pt}

\listoftables

\listoffigures

\pagenumbering{gobble}
\newpage
\pagenumbering{arabic}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center", fig.pos = "H")
knitr::opts_knit$set(root.dir = "C:/Users/bstaton/Desktop/Staton/1_critfc/analyses/GR-sslcm/03-post-process")

# set up a chunk hook to calculate how long chunks took to run
# modified from https://bookdown.org/yihui/rmarkdown-cookbook/time-chunk.html#time-chunk

start_times = list()  # store the time for each chunk
end_times = list()  # store the time for each chunk
knitr::knit_hooks$set(time_it = local({
  function(before, options) {
    if (before) {
      start_times[[options$label]] <<- lubridate::now()
    } else {
      end_times[[options$label]] <<- lubridate::now()
    }
  }
}))
```

```{r load-post, time_it = TRUE}
source("C:/Users/bstaton/Desktop/source_rmd_chunk.R")
rmd_file = "C:/Users/bstaton/Desktop/Staton/1_critfc/analyses/GR-sslcm/03-post-process/output-plots.Rmd"
source_rmd_chunk(rmd_file, "setup")
source_rmd_chunk(rmd_file, "dimension-ids")
knitr::opts_chunk$set(out.width = "100%", dev = "cairo_pdf")
```

```{r load-packages}
source("../00-packages.R")
```

```{r plot-settings}
solid_col = "grey30"
tranp_col = alpha(solid_col, 0.25)
pt_cex = 1.3
```

```{r captions}
beta_caption = "
  Estimated relationship between parr capacity and weighted usable habitat length.
  Darker grey/thicker error bars represent 80% credible regions and lighter grey/thinner error bars represent 95% credible regions.
"

```

```{r beta-fig, fig.width = 3.4, fig.height = 3.4, fig.cap = msdown::prep_caption(beta_caption), time_it = TRUE}
# extract posterior summaries of the capacity parameters
beta_ests = post_summ(post, "^beta[", probs = c(0.025, 0.1, 0.5, 0.9, 0.975))

# extract posterior samples of the slope parameter
beta_per = post_subset(post, "lambda", TRUE)

# create a sequence of WUL values to predict capacity at
wul_seq = seq(0, max(jags_data$wul) * 1.1, length = 30)

# obtain the regression line for each posterior sample
pred_beta = t(sapply(beta_per, function(slope) wul_seq * slope))

# summarize posterior of regression line
pred_beta_summ = apply(pred_beta, 2, function(x) c(mean = mean(x), quantile(x, c(0.025, 0.1, 0.5, 0.9, 0.975))))

# empty plot with correct labeling/dimensions
mypar(mfrow = c(1,1), col.axis = "black")
plot(1,1, type = "n", xlim = range(wul_seq),
     # ylim = c(0, max(beta_ests["97.5%",], pred_beta_summ["97.5%",])),
     ylim = c(0, 3000) * 1000,
     xlab = "",
     ylab = "", yaxt = "n")

# draw prettier y-axis
at_y = c(0, 1000, 2000, 3000) * 1000
axis(side = 2, at = at_y, labels = at_y/1000)
usr = par("usr"); xdiff = diff(usr[1:2]); ydiff = diff(usr[3:4])

# draw the regression uncertainty
polygon(x = c(wul_seq, rev(wul_seq)), y = c(pred_beta_summ["2.5%",], rev(pred_beta_summ["97.5%",])), border = NA, col = tranp_col)
polygon(x = c(wul_seq, rev(wul_seq)), y = c(pred_beta_summ["10%",], rev(pred_beta_summ["90%",])), border = NA, col = tranp_col)

# draw the posterior mean regression line
lines(pred_beta_summ["mean",] ~ wul_seq, lwd = 2, lty = 2, col = solid_col)

# draw the population-specific capacity estimates w/error bars and labels
segments(jags_data$wul, beta_ests["2.5%",], jags_data$wul, beta_ests["97.5%",], col = solid_col)
segments(jags_data$wul, beta_ests["10%",], jags_data$wul, beta_ests["90%",], col = solid_col, lwd = 3)
points(beta_ests["mean",] ~ jags_data$wul, cex = 1.5, pch = 21, col = solid_col, bg = solid_col)

# draw the population labels
xoff = 0.015
yoff = 0.03
text(x = jags_data$wul      + xdiff * c(1.5,-1,1,1) * xoff,
     y = beta_ests["mean",] + ydiff * c(1,-1,1,1) * yoff,
     labels = c("CAT", "LOS", "MIN", "UGR"), pos = c(2,4,2,2), font = 1, cex = 0.8)

# add axis labels
axis_labels("Weighted Usable Habitat Length (km)", "Parr Capacity (Thousands)")
```