### 2023-08-17 03-post-process/output-plots-children/09-equiv-rates.Rmd
Improvements to equivalency rates Rmd section

---
editor_options: 
  chunk_output_type: console
---

# Composite Rates {.tabset .tabset-dropdown}

The values reported in this section are not directly estimated quantities (i.e., do not have a prior distribution assigned), but rather composite rates that stem from combining multiple life stages together.
In all cases, the rates were calculated by dividing the number of fish produced by the process by the number of fish that initiated the process.

These rates are calculated through only brood year `r max(all_yrs[sar_yrs])`, since this is the last brood with of all adult ages (3, 4, and 5) accounted for in the model.

## Smolt $\to$ Adult Rates {.tabset .tabset-pills}

### BON $\to$ BON

```{r SAR-bon-to-bon, fig.width = p4_width, fig.height = p4_height}
plot_fn = function(j) {
  # which years had zero hatchery smolt released
  zero_hor_smolt = post_summ(post, sub_index("Mb[year,LH_type,origin,pop]", year = sar_yrs, LH_type = i_spring, origin = o_hor, pop = j))["mean",] == 0

  # summarize rates
  phi_nor = post_summ(post, sub_index("phi_O0_Rb_BON[year,origin,pop]", year = sar_yrs, origin = o_nor, pop = j))
  phi_hor = post_summ(post, sub_index("phi_O0_Rb_BON[year,origin,pop]", year = sar_yrs, origin = o_hor, pop = j))

  # get a max value for ylim
  ymax = max(post_summ(post, sub_index("phi_O0_Rb_BON[year,origin,pop]", year = sar_yrs, origin = o_nor, pop = "."))["97.5%",])

  # make hatchery values NA in year with no smolt
  phi_hor[,zero_hor_smolt] = NA

  # plot natural origin time series
  plot_tseries(est = phi_nor, yrs = all_yrs[sar_yrs], label_text = pops[j], ylim = c(0, ymax))

  # add hatchery origin time series
  add_tseries(est = phi_hor, yrs = all_yrs[sar_yrs])

  if (j == 2) {
    legend("topright", legend = origins, title = "Origin",
           pch = 22, col = main_cols[c("model", "model2")], pt.cex = 2, pt.bg = tran_cols[c("model", "model2")], bty = "n", cex = 0.8)
  }
}
mypar()
junk = sapply(1:jags_data$nj, plot_fn)
axis_labels("Brood Year", "Smolt \u2192 Adult Survival")
```

### LGR $\to$ LGR

```{r SAR-LGR-to-LGR, fig.width = p4_width, fig.height = p4_height}
plot_fn = function(j) {
  # which years had zero hatchery smolt released
  zero_hor_smolt = post_summ(post, sub_index("Mb[year,LH_type,origin,pop]", year = sar_yrs, LH_type = i_spring, origin = o_hor, pop = j))["mean",] == 0

  # summarize rates
  Ra_per_Ma_nor = post_summ(post, sub_index("Ra_per_Ma[year,origin,pop]", year = sar_yrs, origin = o_nor, pop = j))
  Ra_per_Ma_hor = post_summ(post, sub_index("Ra_per_Ma[year,origin,pop]", year = sar_yrs, origin = o_hor, pop = j))

  # make hatchery values NA in year with no smolt
  Ra_per_Ma_hor[,zero_hor_smolt] = NA

  # plot natural origin time series
  plot_tseries(est = Ra_per_Ma_nor, yrs = all_yrs[sar_yrs], label_text = pops[j], ylim = c(0, 0.1))

  # add hatchery origin time series
  add_tseries(est = Ra_per_Ma_hor, yrs = all_yrs[sar_yrs])

  if (j == 2) {
    legend("topright", legend = origins, title = "Origin",
           pch = 22, col = main_cols[c("model", "model2")], pt.cex = 2, pt.bg = tran_cols[c("model", "model2")], bty = "n", cex = 0.8)
  }
}
mypar()
junk = sapply(1:jags_data$nj, plot_fn)
axis_labels("Brood Year", "Smolt \u2192 Adult Survival Rate")
```

## Parr per Spawner

"Spawner" means total adults reaching the spawning grounds that survive pre-spawn mortality.

```{r parr-per-spawner, fig.width = p4_width, fig.height = p4_height}
plot_fn = function(j) {
  plot_tseries(est = post_summ(post, sub_index("Pb_per_Sa_tot[year,pop]", pop = j, year = ts_yrs)), yrs = all_yrs[ts_yrs], label_text = pops[j])
}

mypar()
junk = sapply(1:jags_data$nj, plot_fn)
axis_labels("Brood Year", "Parr per Spawner")
```

## Smolt per Spawner

"Smolt" means total NOR smolt immediately prior to migration out of basin.

```{r smolt-per-spawner, fig.width = p4_width, fig.height = p4_height}
plot_fn = function(j) {
  plot_tseries(est = post_summ(post, sub_index("Mb_per_Sa_tot[year,pop]", pop = j, year = ts_yrs)), yrs = all_yrs[ts_yrs], label_text = pops[j])
}

mypar()
junk = sapply(1:jags_data$nj, plot_fn)
axis_labels("Brood Year", "Smolt per Spawner")
```

## Spawners per Spawner

This is the sum of all spawners that were produced by the spawners in a given brood year (i.e., accounts for adults from a cohort returning across multiple ages/years). "Brood year" refers to the year the returning cohort was spawned.

```{r spawners-per-spawner, fig.width = p4_width, fig.height = p4_height}
plot_fn = function(j) {
  plot_tseries(est = post_summ(post, sub_index("Sa_tot_per_Sa_tot[year,pop]", pop = j, year = sar_yrs)), yrs = all_yrs[sar_yrs], label_text = pops[j])
  abline(h = 1, lty = 2)
}
mypar()
junk = sapply(1:jags_data$nj, plot_fn)
axis_labels("Brood Year", "Spawners per Spawner")
```

## Spawners per Adult Return {.tabset .tabset-pills}

Unlike the previous figures, this figure shows the proportion of the adults that returned to the Columbia River each year (prior to any mortality on upstream migration) that ultimately spawned successfully.

### Time Series {.tabset .tabset-pills}

```{r spawners-per-return-fn}
# function to calculate the survival term for a given origin, population, and combo of ages
f = function(o, j, include_age3) {
  # spawning adult abundance
  Sa = post_subset(post, sub_index("^Sa[year,age,origin,pop]", year = ".+", age = ".+", origin = o, pop = j), matrix = TRUE)

  # adults at estuary
  Rb = post_subset(post, sub_index("^Rb[year,age,origin,pop]", year = 2:33, age = ".+", origin = o, pop = j), matrix = TRUE)

  # function to calculate for a given MCMC iteration
  g = function(i, include_age3) {
    # put the values from this draw into the format used in the model
    Sa_i = array_format(Sa[i,])[,,o,j]
    Rb_i = array_format(Rb[i,])[,,o,j]

    # choose which columns to include in sum over ages
    if (include_age3) keep_cols = 1:3 else keep_cols = 2:3

    # calculate the composite survival
    phi_Rb_Sa_i = rowSums(Sa_i[,keep_cols], na.rm = TRUE)/rowSums(Rb_i[,keep_cols], na.rm = TRUE)

    # Inf comes from dividing by zero (i.e., the stray years)
    # NaN comes from a dividing zero by zero (i.e., the first year)
    phi_Rb_Sa_i[phi_Rb_Sa_i == Inf | phi_Rb_Sa_i == "NaN"] = NA

    # in some rare cases, the max() constraint produces fish
    phi_Rb_Sa_i[phi_Rb_Sa_i > 1] = 0.99
    as.numeric(phi_Rb_Sa_i)
  }

  # apply to each posterior sample
  out = t(sapply(1:nrow(Sa), g, include_age3 = include_age3))

  # set index names
  y_ind = 1:ncol(out)
  base_name = ifelse(include_age3, "phi_Rb_Sa", "phi_Rb_Sa_45")
  colnames(out) = sub_index(paste0(base_name, "[year,origin,pop]"), year = y_ind, origin = o, pop = j)

  # remove NA values
  na_ind = which(!is.na(out[1,]))
  out[,na_ind]
}

# apply the function to all populations and ages
out = do.call(cbind, lapply(1:jags_data$nj, function(j) {
  cbind(f(o = o_nor, j = j, include_age3 = TRUE),
        f(o = o_hor, j = j, include_age3 = TRUE),
        f(o = o_nor, j = j, include_age3 = FALSE),
        f(o = o_hor, j = j, include_age3 = FALSE)
  )
}))

# combine with other posterior samples
post = post_bind(post, out)

# function to create the plot for one population
plot_f = function(j, include_age3) {

  param = paste0(ifelse(include_age3, "phi_Rb_Sa", "phi_Rb_Sa_45"), "[year,origin,pop]")
  est_nor = post_summ(post, sub_index(param, year = ".+", origin = o_nor, pop = j))

  if (j != j_min) {
    est_hor = post_summ(post, sub_index(param, year = ".+", origin = o_hor, pop = j))
    hor_yrs = stringr::str_extract(colnames(est_hor), "\\[[0-9]+") |> stringr::str_remove("\\[") |> as.numeric()
    missing_hor_yrs = ts_yrs[!(ts_yrs %in% hor_yrs)]

    if (length(missing_hor_yrs) > 0) {
      est_empty = replicate(n = length(missing_hor_yrs), expr = rep(NA, 5))
      rownames(est_empty) = rownames(est_hor)
      est_hor = cbind(est_empty, est_hor)
    }
  }

  plot_tseries(est = est_nor, label_text = pops[j], yrs = all_yrs[ts_yrs], ylim = c(0,1))
  if (j != j_min) add_tseries(est = est_hor, yrs = all_yrs[ts_yrs])

  if (j == 1) {
    legend("topright", legend = origins, title = "Origin",
           pch = 22, col = main_cols[c("model", "model2")], pt.cex = 2,
           pt.bg = tran_cols[c("model", "model2")], bty = "n", cex = 0.8)
  }
}
```

#### All Ages {-}

```{r spawners-per-return-all-ages, fig.width = p4_width, fig.height = p4_height}

# loop over populations
# includes all ages
mypar()
junk = sapply(1:jags_data$nj, plot_f, include_age3 = TRUE)
axis_labels("Return Year", "Spawners per Returning Adult (All Ages)")
```

### Ages 4&5 {-}

```{r spawners-per-return-old-ages, fig.width = p4_width, fig.height = p4_height}

# loop over populations
# includes only ages 4 and 5
mypar()
junk = sapply(1:jags_data$nj, plot_f, include_age3 = FALSE)
axis_labels("Brood Year", "Spawners per Returning Adult (Ages 4&5)")
```

### Tables {.tabset .tabset-pills}

These tables summarize the time series shown.
The summary statistic across years was calculated for each posterior sample and the posterior means are reported.

```{r spawners-per-return-table-fns}
# function reporting across year summary stat
# for a given origin, population, and age grouping
f = function(o, j, include_age3, FUN, ...) {
  if (o == o_hor & j == j_min) {
    return(NA)
  } else {
    # create the parameter name to subset
    param = paste0(ifelse(include_age3, "phi_Rb_Sa", "phi_Rb_Sa_45"), "[.+,origin,pop]")

    # extract posterior samples for this origin and population
    samps = post_subset(post, sub_index(param, origin = o, pop = j), matrix = TRUE)

    # obtain posterior samples of the inter-annual summary statistic returned by FUN
    stat = apply(samps, 1, FUN, ...)

    # return the posterior mean of the summary statistic returned by FUN
    mean(stat)
  }
}

# wrapper for f()
# performs task for all populations, origins, and age groupings
g = function(FUN, digits = 2, ...) {

  # summarize the posteriors
  x = rbind(
    sapply(1:jags_data$nj, function(j) f(o = o_nor, j = j, include_age3 = TRUE, FUN = FUN, ...)),
    sapply(1:jags_data$nj, function(j) f(o = o_hor, j = j, include_age3 = TRUE, FUN = FUN, ...)),
    sapply(1:jags_data$nj, function(j) f(o = o_nor, j = j, include_age3 = FALSE, FUN = FUN, ...)),
    sapply(1:jags_data$nj, function(j) f(o = o_hor, j = j, include_age3 = FALSE, FUN = FUN, ...))
  ) |> as.data.frame()
  colnames(x) = pops

  # add identifiers
  tab = cbind(
    age_type = rep(c("All Ages", "Ages 4&5"), each = 2),
    origin_type = rep(c("NOR", "HOR"), 2),
    round(x, digits = digits)
  )

  # make NA values empty
  tab[is.na(tab)] = " "

  # return the table, ready to be passed to kable
  return(tab)
}

kable_f = function(tab) {
  tab |>
    kbl(format = "html", col.names = c(" ", "Origin", pops)) |>
    kable_styling(full_width = FALSE, bootstrap_options = c("condensed", "striped")) |> 
    column_spec(1:2, bold = TRUE) |>
    collapse_rows(1)
}
```

#### Inter-Annual Mean {-}

```{r, spawners-per-return-table-mean}
# posterior means of the inter-annual mean survival rate
g(FUN = mean) |>
  kable_f()
```

#### Inter-Annual q10 {-}

```{r, spawners-per-return-table-q10}
# posterior means of the survival value that 10% of years fall below
g(FUN = quantile, probs = 0.1) |>
  kable_f()
```

#### Inter-Annual q90 {-}

```{r, spawners-per-return-table-mean-q90}
# posterior means of the survival value that 90% of years fall below
g(FUN = quantile, probs = 0.9) |>
  kable_f()
```