### 2023-12-06 03-post-process/output-plots-children/01-mcmc-dims-runtime.Rmd

---
editor_options: 
  chunk_output_type: console
---

# MCMC Dimensions & Run Time

```{r mcmc-dims-and-runtime}

## MCMC dimensions ##

# query the dimensions of the posterior samples: the full samples even if thin_percent not equal to 1
dims = post_dim(model_info$post)

# build a data frame for printing this information
df_dims = data.frame(
  Quantity = c("Burn-in + Adapt Samples", "Post Burn-in Samples", "Thinning Interval", "Chains", "Saved Samples", "Monitored Quantities", "% of Saved Samples used Here", "Chains Ran in Parallel?"),
  Value = c(unname(dims), NA, NA)
)
df_dims$Value = prettyNum(df_dims$Value, big.mark = ",", scientific = FALSE)
df_dims$Value[df_dims$Quantity == "Chains Ran in Parallel?"] = ifelse(model_info$jags_dims$parallel, "Yes", "No")
df_dims$Value[df_dims$Quantity == "% of Saved Samples used Here"] = paste0(params$thin_percent * 100, "%")

## MCMC Runtime ##

# Calculate hours per 10,000 iterations
hrs_per_ten_thousand_iters = with(model_info$jags_dims, {
  # calculation of total iters differs if parallel or not
  if (parallel) {
    total_iters = n_adapt + n_post + n_burn
  } else {
    total_iters = (n_adapt + n_post + n_burn) * n_chains
  }
  ten_thousands_of_total_iters = total_iters/10000

  # calculate hours elapsed
  time_diff = as.POSIXct.default(model_info$jags_time["stoptime"]) - as.POSIXct.default(model_info$jags_time["starttime"])
  time_number = as.numeric(time_diff)
  time_units = attributes(time_diff)$units
  hour_convert = switch(time_units,
                        "secs" = 3600,
                        "mins" = 60,
                        "hours" = 1,
                        "days" = 1/24)
  hours_diff = time_number/hour_convert

  # calculate/round rate
  round(hours_diff/ten_thousands_of_total_iters, 2)
})

# combine relevant info into a data frame
df_time = data.frame(
  Quantity = c("Time Started", "Time Ended", "Time Elapsed", "Hours Per 10,000 Iterations"),
  Value = c(model_info$jags_time, hrs_per_ten_thousand_iters)
)

# combine the dimensions and run time dfs into one
df = rbind(df_dims, df_time)
groups = c(rep(1, nrow(df_dims)), rep(2, nrow(df_time)))
rownames(df) = NULL

# render the table
kable(df, format.args = list(big.mark = ","), caption = "MCMC sampling dimensions and run time.", col.names = c(" ", " "), escape = FALSE) |>
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "condensed")) |>
  group_rows("Dimensions", start_row = min(which(groups == 1)), end_row = max(which(groups == 1))) |>
  group_rows("Per Chain", start_row = 1, end_row = 3, escape = FALSE) |>
  group_rows("Overall", start_row = 4, end_row = max(which(groups == 1))) |>
  group_rows("Run Time", start_row = min(which(groups == 2)), end_row = max(which(groups == 2)))
```